### 2023-09-23 ukb_mods.Rmd
basic setup

---
title: "UKB Models"
output: 
  html_document:
    code_folding: hide
    toc: true
    toc_float: true
date: '2023-09-23'
---

## Setup

```{r setup, include=FALSE, warning = FALSE, message = FALSE}
knitr::opts_chunk$set(echo = TRUE)
if(!require('pacman')) {
  install.packages('pacman')
}
pacman::p_load(gamlss, gamlss.cens, gamlss.dist, gamlss.mx, gamlss.tr, ggplot2, tidyverse, ggpubr, skimr, gamlss.data, data.table, nlme, plyr, shinySIR, patchwork, ggstatsplot, qqplotr, arrangements, purrr, parallel, rlang)
```

## Load Data

## EDA