### 2023-10-11 ukb_curation-eda.Rmd
fixed summary parsing, visualizing model outputs

---
title: "UKB Models"
output: 
  html_document:
    code_folding: hide
    toc: true
    toc_float: true
date: '2023-09-23'
---

## Setup

```{r setup, include=FALSE, warning = FALSE, message = FALSE}
knitr::opts_chunk$set(echo = TRUE)
if(!require('pacman')) {
  install.packages('pacman')
}
pacman::p_load(gamlss, gamlss.cens, gamlss.dist, gamlss.mx, gamlss.tr, ggplot2, tidyverse, ggpubr, skimr, gamlss.data, data.table, nlme, devtools, slider)

# #pull plotting functions from github
# plot.func.url <- "https://raw.githubusercontent.com/BGDlab/GAMLSS/main/plotting_functions.R?token=GHSAT0AAAAAACGAYP4HZWCTHACGUM6KG2PSZIQUPBA"
# Sys.setenv(GITHUB_PAT = "") #if needed to access private BGDlab GAMLSS repo
# source_url(plot.func.url) #sometimes doesn't work....?

source("/Users/megardn/Desktop/BGD_Repos/GAMLSS/plotting_functions.R")
```

## Load Data

Load UKB Data from Jakob

```{r}
cn_data <- fread(file="data/lifespan_dataset_dkatlas_CNonly.csv")
#table(cn_data$study)
```

## Data Org

Filter

```{r}
ukb_df <- cn_data %>%
  dplyr::filter(study == "UKB") %>%
  mutate(age_yrs = round(age_days/365.25),
         log_age = log(age_days, base=10),
         TBV=(sGMV+GMV+WMV+Ventricles),
         fs_version = as.factor(fs_version),
         sex = as.factor(sex),
         sex_int = as.numeric(case_when(sex == "Male" ~ 0,
                                        sex == "Female" ~ 1)),
         sex.age = sex_int * age_days) %>%
  na.omit() #1 subj has 1 missing val

skim(ukb_df)
n_unique(ukb_df$participant) == nrow(ukb_df) #no duplicate subj
```

Write lists of phenotypes

```{r}
col_list <- colnames(ukb_df)
pheno_list <- col_list[grepl("^lh_|^rh_|^GMV$|^WMV$|^sGMV$|^Ventricles$", col_list)]
length(pheno_list) #total number of models to run

#split by measure type (Vol, SA, CT)
vol_list_global <- c("GMV", "sGMV", "WMV", "Ventricles")
writeLines(vol_list_global, con = "pheno_lists/Vol_list_global.txt")
saveRDS(vol_list_global, file="R_scripts/vol_list_global.rds")

vol_list_regions <- col_list[grepl("^lh_Vol|^rh_Vol", col_list)]
writeLines(vol_list_regions, con = "pheno_lists/Vol_list_regions.txt")
saveRDS(vol_list_regions, file="R_scripts/Vol_list_regions.rds")

sa_list <- col_list[grepl("^lh_SA|^rh_SA", col_list)]
writeLines(sa_list, con = "pheno_lists/SA_list.txt")
saveRDS(sa_list, file="R_scripts/SA_list.rds")

ct_list <- col_list[grepl("^lh_CT|^rh_CT", col_list)]
writeLines(ct_list, con = "pheno_lists/CT_list.txt")
saveRDS(ct_list, file="R_scripts/CT_list.rds")
```

Add brainwide phenotypes!
Summing across regions for volumes (rather than summing GMV, sGMV, etc) for consistency with SA and CT

```{r}
vol_list_regions <- col_list[grepl("^lh_Vol|^rh_Vol", col_list)]

ukb_df_totals <- ukb_df %>%
  mutate(Vol_total=rowSums(select(., .dots=all_of(c(vol_list_regions)))),
         SA_total=rowSums(select(., .dots=all_of(c(sa_list)))),
         CT_total=rowSums(select(., .dots=all_of(c(ct_list)))))

#write data
write_csv(ukb_df_totals, file="data/ukb_CN_data.csv")
```

```{r}
ggplot(data=ukb_df, aes(x=age_days)) +
  geom_histogram(aes(fill=sex), alpha=0.6, position="identity", binwidth = 365)
ggplot(data=ukb_df, aes(x=log_age)) +
  geom_histogram(aes(fill=sex), alpha=0.6, position="identity")

ggplot(data=ukb_df, aes(x=log_age, y=GMV)) +
  geom_point(aes(color=sex), alpha=0.6)
ggplot(data=ukb_df, aes(x=age_days, y=GMV)) +
  geom_point(aes(color=sex), alpha=0.6)
```

### Condense age range

Random male subjects are younger than females - going to drop because having issues estimating total brain metrics in young women where there's no data

```{r}
ukb_df_totals %>%
  dplyr::filter(age_yrs <= 49) %>%
ggplot(data=, aes(x=age_days)) +
  geom_histogram(aes(fill=sex), alpha=0.6, position="identity", binwidth = 365)

ukb_df_totals %>%
  dplyr::filter(age_yrs >= 79) %>%
ggplot(data=, aes(x=age_days)) +
  geom_histogram(aes(fill=sex), alpha=0.6, position="identity", binwidth = 365)
```

Restricting to ages 50-80, since that's when there's roughly ~100 ppts per sex

```{r}
ukb_df_age.filt <- ukb_df_totals %>%
  dplyr::filter(age_yrs >= 50 & age_yrs <= 80)
#write data
write_csv(ukb_df_age.filt, file="data/ukb_CN_data_agefilt.csv")
```

## EDA

Going to stick with age_days since distribution is fairly normal, not lifespan data

```{r}
ggplot(data=ukb_df_totals, aes(x=age_days, y=Vol_total)) +
  geom_point(aes(color=sex), alpha=0.6, position="identity")

ggplot(data=ukb_df_totals, aes(x=age_days, y=TBV)) +
  geom_point(aes(color=sex), alpha=0.6, position="identity") #TBV includes ventricles, WMV, subcortical GMV

ggplot(data=ukb_df_totals, aes(x=age_days, y=CT_total)) +
  geom_point(aes(color=sex), alpha=0.6, position="identity")

ggplot(data=ukb_df_totals, aes(x=age_days, y=SA_total)) +
  geom_point(aes(color=sex), alpha=0.6, position="identity")
```

Look at variances
```{r}
plot.hist.by.sex <- function(data, pheno_list) {

  # Create an empty list to store the ggplot objects
  plots <- list()

  # Loop through the list of variables
  for (pheno in pheno_list) {
    # Check if the variable exists in the data frame
    if (!pheno %in% names(data)) {
      warning(paste("Variable '", pheno, "' not found in the data frame. Skipping..."))
      next
    }

    # Create a ggplot histogram for the variable
    plot <- ggplot(data, aes(x = .data[[pheno]], fill= sex, color=sex)) +
      geom_histogram(alpha=0.5, position="identity") +
      labs(title = paste("Histogram of", pheno), x = pheno, y = "Frequency")

    # Add the plot to the list
    plots[[pheno]] <- plot
  }

  # Return the list of ggplot objects
  return(plots)
}
```

center
```{r}
ukb_centered <- ukb_df_totals %>%
  group_by(sex) %>%
  mutate_if(is.numeric, funs(scale(., center=TRUE, scale=FALSE))) %>%
  ungroup()
```

```{r}
# Generate histograms (mean-centered w/in sex)
global_plots <- plot.hist.by.sex(ukb_centered, vol_list_global)

# Print or display the histograms
for (pheno in vol_list_global) {
  print(global_plots[[pheno]])
}
```

```{r}
# Generate histograms (mean-centered w/in sex)
region_vol_plots <- plot.hist.by.sex(ukb_centered, vol_list_regions)

# Print or display the histograms
for (pheno in vol_list_regions) {
  print(region_vol_plots[[pheno]])
}
```

```{r}
# Generate histograms (mean-centered w/in sex)
region_sa_plots <- plot.hist.by.sex(ukb_centered, sa_list)

# Print or display the histograms
for (pheno in sa_list) {
  print(region_sa_plots[[pheno]])
}
```

```{r}
# Generate histograms (mean-centered w/in sex)
region_ct_plots <- plot.hist.by.sex(ukb_centered, ct_list)

# Print or display the histograms
for (pheno in ct_list) {
  print(region_ct_plots[[pheno]])
}
```