### 2023-10-11 combat_sandbox.Rmd
working through combat for regional vols - some warninigs/errs

---
title: "Combat Sandbox"
output: html_document
date: '2023-10-06'
---

## Setup

```{r setup, include=FALSE}
knitr::opts_chunk$set(echo = TRUE, warning = FALSE, message = FALSE)
if(!require('pacman')) {
  install.packages('pacman')
}
pacman::p_load(devtools, data.table, readr, ggplot2, tidyverse, mgcv)
#install.packages("devtools")
devtools::install_github("andy1764/ComBatFamily", build_vignettes = FALSE) #per discussion with Andrew, sent vignettes separately
library(ComBatFamily)
```

#Prep data

Using additional instructions from J Fortin's [initial repo](https://github.com/Jfortin1/neuroCombat_Rpackage/tree/2ccab2b42c59163717f012c70a54d20e8757b1c7).

Just using regional volume measures for now.

```{r}
ukb_df_sites <- read_csv(file="data/ukb_CN_data_simsites.csv")
vol_list_regions <- readRDS(file="R_scripts/Vol_list_regions.rds")
```

```{r}
#regional volume data to harmonize
rvol.df <- ukb_df_sites %>%
  dplyr::select(all_of(vol_list_regions))

#fake batches
batch <- as.factor(ukb_df_sites$sim.site)

#covariates
covar.mat <- ukb_df_sites %>%
  dplyr::select(c(age_days, sex)) %>%
  mutate(age_days = as.numeric(age_days)) %>%
  as.matrix()
#weirdly converts age_days to a character, idk if that will be a problem

nrow(rvol.df) == length(batch)
nrow(rvol.df) == nrow(covar.mat)
```

## ComBat

simple combat with linear model for covariates
```{r}
cf.lm <- comfam(rvol.df, batch, covar.mat, lm, formula = y ~ age_days + sex)
```

Getting warning - Warning: prediction from a rank-deficient fit may be misleading

combat with age as smooth
```{r}
cf.gam <- comfam(rvol.df, batch, covar.mat, gam, formula = y ~ s(age_days) + sex)
```