### 2023-10-12 combat_sandbox.Rmd
playing around with different combats' variance outputs, plots

---
title: "Combat Sandbox"
output: html_document
date: '2023-10-06'
---

Playing around with different combat configs using the UKB data, which has been assigned to 3 simulated study sites (see `proof-of-concept_mods.Rmd`)

## Setup

```{r setup, include=FALSE}
knitr::opts_chunk$set(echo = TRUE, warning = FALSE, message = FALSE)
if(!require('pacman')) {
  install.packages('pacman')
}
pacman::p_load(devtools, data.table, readr, ggplot2, tidyverse, mgcv, gamlss)
#install.packages("devtools")
devtools::install_github("andy1764/ComBatFamily", build_vignettes = FALSE) #per discussion with Andrew, sent vignettes separately
library(ComBatFamily)
```

#Prep data

Using additional instructions from J Fortin's [initial repo](https://github.com/Jfortin1/neuroCombat_Rpackage/tree/2ccab2b42c59163717f012c70a54d20e8757b1c7).

Just using regional volume measures for now.

```{r}
ukb_df_sites <- read_csv(file="data/ukb_CN_data_simsites.csv")
vol_list_regions <- readRDS(file="R_scripts/Vol_list_regions.rds")
```

```{r}
#regional volume data to harmonize
rvol.df <- ukb_df_sites %>%
  dplyr::select(all_of(vol_list_regions))

#fake batches
batch <- as.factor(ukb_df_sites$sim.site)

#covariates - using model.matrix to dummycode sex as numeric
covar.mat <- model.matrix( ~ age_days + sex, data=ukb_df_sites)

nrow(rvol.df) == length(batch)
nrow(rvol.df) == nrow(covar.mat)
```

## ComBat

simple combat with no covariate preservation
```{r}
cf <- comfam(rvol.df, batch)
cf$dat.combat
cf$dat.combat[, "sim.site"] <- batch
```

simple combat with linear model for covariates
```{r}
cf.lm <- comfam(rvol.df, batch, covar.mat, lm, formula = y ~ age_days + sexMale)
cf.lm$dat.combat
```
plot
```{r}
dim(cf.lm$dat.combat)[1] == length(batch)
#add back site
cf.lm$dat.combat[, "sim.site"] <- batch

ggplot() + 
 #geom_histogram(aes(x=cf.lm$dat.combat$GMV, y=after_stat(density), fill=cf.lm$dat.combat$sim.site, color=cf.lm$dat.combat$sim.site), alpha = .2)+
 geom_density(aes(x=cf.lm$dat.combat$lh_Vol_superiorparietal, fill=cf.lm$dat.combat$sim.site, color=cf.lm$dat.combat$sim.site), alpha=.4) +
  geom_vline(aes(xintercept=cf.lm_summary$sd_below.lh_v_sp, color=cf.lm_summary$sim.site)) +
  geom_vline(aes(xintercept=cf.lm_summary$sd_above.lh_v_sp, color=cf.lm_summary$sim.site))
```

Try setting site A (which has even sex distribution) as ref

```{r}
cf.lm2 <- comfam(rvol.df, batch, covar.mat, lm, formula = y ~ age_days + sexMale, ref.batch = "Site_A")
cf.lm2$dat.combat
cf.lm2$dat.combat[, "sim.site"] <- batch
```

combat gam with age as smooth
```{r}
cf.gam <- comfam(rvol.df, batch, covar.mat, gam, formula = y ~ s(age_days) + sexMale)
cf.gam$dat.combat[, "sim.site"] <- batch
```

```{r}
cf.gam2 <- comfam(rvol.df, batch, covar.mat, gam, formula = y ~ s(age_days) + sexMale, ref.batch = "Site_A")
cf.gam2$dat.combat[, "sim.site"] <- batch
```

combat gamlss!

```{r}
cf.gamlss <- comfam(rvol.df, batch, covar.mat, gamlss, y ~ age_days + sexMale,
              sigma.formula = ~ age_days + sexMale, 
              control = gamlss.control(trace = FALSE))
cf.gamlss$dat.combat[, "sim.site"] <- batch
```

```{r}
cf.gamlss2 <- comfam(rvol.df, batch, covar.mat, gamlss, y ~ age_days + sexMale,
              sigma.formula = ~ age_days + sexMale, 
              control = gamlss.control(trace = FALSE),
              ref.batch = "Site_A")
cf.gamlss2$dat.combat[, "sim.site"] <- batch
```

## Compare Variances

Randomly picked `lh_Vol_superiorparietal` to compare

Get variances
```{r}
#raw data
sum_df <- ukb_df_sites %>%
  group_by(sim.site) %>%
  summarise(sd_below.lh_v_sp = mean(lh_Vol_superiorparietal) - sd(lh_Vol_superiorparietal), sd_above.lh_v_sp = mean(lh_Vol_superiorparietal) + sd(lh_Vol_superiorparietal),
            var.lh_v_sp = var(lh_Vol_superiorparietal))
#no covars
cf_summary <- cf$dat.combat %>%
  group_by(sim.site) %>%
  summarise(sd_below.lh_v_sp = mean(lh_Vol_superiorparietal) - sd(lh_Vol_superiorparietal), sd_above.lh_v_sp = mean(lh_Vol_superiorparietal) + sd(lh_Vol_superiorparietal), var.lh_v_sp = var(lh_Vol_superiorparietal))

#simple lm
cf.lm_summary <- cf.lm$dat.combat %>%
  group_by(sim.site) %>%
  summarise(sd_below.lh_v_sp = mean(lh_Vol_superiorparietal) - sd(lh_Vol_superiorparietal), sd_above.lh_v_sp = mean(lh_Vol_superiorparietal) + sd(lh_Vol_superiorparietal), var.lh_v_sp = var(lh_Vol_superiorparietal))

#simple lm with ref site
cf.lm2_summary <- cf.lm2$dat.combat %>%
  group_by(sim.site) %>%
  summarise(sd_below.lh_v_sp = mean(lh_Vol_superiorparietal) - sd(lh_Vol_superiorparietal), sd_above.lh_v_sp = mean(lh_Vol_superiorparietal) + sd(lh_Vol_superiorparietal), var.lh_v_sp = var(lh_Vol_superiorparietal))

#simple gam
cf.gam_summary <- cf.gam$dat.combat %>%
  group_by(sim.site) %>%
  summarise(sd_below.lh_v_sp = mean(lh_Vol_superiorparietal) - sd(lh_Vol_superiorparietal), sd_above.lh_v_sp = mean(lh_Vol_superiorparietal) + sd(lh_Vol_superiorparietal), var.lh_v_sp = var(lh_Vol_superiorparietal))

#simple gam with ref site
cf.gam2_summary <- cf.gam2$dat.combat %>%
  group_by(sim.site) %>%
  summarise(sd_below.lh_v_sp = mean(lh_Vol_superiorparietal) - sd(lh_Vol_superiorparietal), sd_above.lh_v_sp = mean(lh_Vol_superiorparietal) + sd(lh_Vol_superiorparietal), var.lh_v_sp = var(lh_Vol_superiorparietal))

#simple gam
cf.gamlss_summary <- cf.gamlss$dat.combat %>%
  group_by(sim.site) %>%
  summarise(sd_below.lh_v_sp = mean(lh_Vol_superiorparietal) - sd(lh_Vol_superiorparietal), sd_above.lh_v_sp = mean(lh_Vol_superiorparietal) + sd(lh_Vol_superiorparietal), var.lh_v_sp = var(lh_Vol_superiorparietal))

#simple gam with ref site
cf.gamlss2_summary <- cf.gamlss2$dat.combat %>%
  group_by(sim.site) %>%
  summarise(sd_below.lh_v_sp = mean(lh_Vol_superiorparietal) - sd(lh_Vol_superiorparietal), sd_above.lh_v_sp = mean(lh_Vol_superiorparietal) + sd(lh_Vol_superiorparietal), var.lh_v_sp = var(lh_Vol_superiorparietal))
```
compile

```{r}
var_comp_df <- data.frame(sim.site = sim.site.list,
                          raw_data = sum_df$var.lh_v_sp,
                          cf = cf_summary$var.lh_v_sp,
                          cf.lm = cf.lm_summary$var.lh_v_sp,
                          cf.lm2 = cf.lm2_summary$var.lh_v_sp,
                          cf.gam = cf.gam_summary$var.lh_v_sp,
                          cf.gam2 = cf.gam2_summary$var.lh_v_sp,
                          cf.gamlss = cf.gamlss_summary$var.lh_v_sp,
                          cf.gamlss2 = cf.gamlss2_summary$var.lh_v_sp)

var_comp_long <- var_comp_df %>%
  pivot_longer(!sim.site, names_to="type", values_to = "variance")

ggplot(var_comp_long) +
  geom_col(aes(x=type, y=variance, fill=sim.site, color=sim.site), position = "dodge")

ggplot(var_comp_long) +
  geom_col(aes(x=type, y=variance), position = "dodge") +
  theme(axis.text.x = element_text(angle = 90, vjust = 0.5, hjust=1)) +
  facet_wrap(~sim.site)
```