### 2023-10-30 proof-of-concept_mods.Rmd
looking at effects on variance

---
title: "proof-of-concept_mods"
output: html_document
date: '2023-10-06'
---

## QUESTIONS
- effect size for poly(age_days, 3) in gam/gamlss
- next steps abstract v paper
- abs values of centile differences
- z scores instead of centiles

## Setup

```{r setup, include=FALSE, warning = FALSE, message = FALSE}
knitr::opts_chunk$set(echo = FALSE, warning = FALSE)
if(!require('pacman')) {
  install.packages('pacman')
}
pacman::p_load(gamlss, gamlss.cens, gamlss.dist, gamlss.mx, gamlss.tr, ggplot2, tidyverse, ggpubr, skimr, gamlss.data, data.table, nlme, devtools, slider, tidymodels, broom.mixed, ggseg, cowplot, lsr)

source("./R_scripts/gamlss_helper_funs.R")
```

# Analysis Outline
## Load Data

```{r}
ukb_df_totals <- read_csv(file="data/ukb_CN_data.csv")
ukb_df_age.filt <- read_csv(file="data/ukb_to_model/ukb_CN_data_simsites.csv")

vol_list_global <- readRDS(file="R_scripts/vol_list_global.rds")
vol_list_regions <- readRDS(file="R_scripts/Vol_list_regions.rds")
sa_list <- readRDS(file="R_scripts/SA_list.rds")
ct_list <- readRDS(file="R_scripts/CT_list.rds")
pheno_list <- readRDS(file="R_scripts/pheno_list.rds")
```

## Basic Models
Designate *basic* model formula

```{r}
basic.m1 <- gamlss(formula = GMV ~ poly(age_days, 3) + sexMale,
                       sigma.formula = ~ poly(age_days, 3) + sexMale,
                       nu.formula = ~ 1,
                       family = BCCG, data=ukb_df_age.filt, trace = FALSE)
summary(basic.m1)
```

plot

```{r}
#plot_singlestudy_centiles(basic.m1, "GMV", ukb_df_age.filt, "sexMale")
```

# Review Results

scp'd csvs from cubic (`scp -r 'gardnerm@cubic-login.uphs.upenn.edu:/cbica/home/gardnerm/combat-biovar/ukb_basic/*.csv' ./ukb_basic`)

```{r}
drop1.df <- fread("ukb_basic/drop1_tests.csv", stringsAsFactors = TRUE) %>%
  dplyr::select(-V1)
sex.df <- fread("ukb_basic/sex_summary.csv", stringsAsFactors = TRUE) %>%
  dplyr::select(-V1)
sum.df <- fread("ukb_basic/gamlss_summary.csv", stringsAsFactors = TRUE)%>%
  dplyr::select(-V1)
```

```{r}
#add more info from parcellations
dk.parc <- dk$data %>%
  as.data.frame() %>%
  na.omit() %>%
  dplyr::select(c(hemi, region, label)) %>%
  distinct()

sex.df <- left_join(sex.df, dk.parc, by="label")
sum.df <- left_join(sum.df, dk.parc, by="label")
```

### Raw Data Results

```{r, fig.width=11}
sex.df %>%
  dplyr::filter(sig.sum_bf.corr == TRUE & parameter == "sigma" & dataset ==	"ukb-CN-data-simsites") %>%
  ggplot() +
  geom_col(aes(x=reorder(pheno,t_stat), y=t_stat, fill=pheno_cat)) + 
  theme(axis.text.x = element_text(angle = 90, vjust = 0.5, hjust=1)) +
  ggtitle("Significant effects of male sex on sigma across phenotype-specific brain charts")

sex.df %>%
  dplyr::filter(sig.sum_bf.corr == TRUE & parameter == "sigma" & dataset == "ukb-CN-data-simsites") %>%
  ggplot() +
  geom_col(aes(x=label, y=t_stat, fill=pheno_cat)) + 
  theme(axis.text.x = element_text(angle = 90, vjust = 0.5, hjust=1)) +
  ggtitle("Significant effects of male sex on sigma across phenotype-specific brain charts")

sex.df %>%
  dplyr::filter(sig.sum_bf.corr == TRUE & parameter == "sigma" & dataset == "ukb-CN-data-simsites") %>%
  ggplot() +
  geom_col(aes(x=region, y=t_stat, fill=hemi)) + 
  facet_wrap(~pheno_cat) +
  theme(axis.text.x = element_text(angle = 90, vjust = 0.5, hjust=1)) +
  ggtitle("Significant effects of male sex on sigma across phenotype-specific brain charts")
```

Plot on brains!

```{r}
sex.df %>%
  dplyr::filter(sig.sum_bf.corr == TRUE & parameter == "sigma" & dataset == "ukb-CN-data-simsites") %>%
  group_by(pheno_cat) %>%
  ggplot() +
  geom_brain(atlas = dk, 
             position = position_brain(hemi ~ side),
             aes(fill = t_stat)) +
  scale_fill_gradient2() +
  facet_wrap(~pheno_cat) +
  theme_void() +
  theme(axis.title = element_blank(), axis.text = element_blank(), axis.ticks = element_blank()) +
  ggtitle("t-statistic for effect of male sex on sigma")
```

### Combat Effects

#### Site Effects
Loading combatted data (scp'd from cubic: `scp -r 'gardnerm@cubic-login.uphs.upenn.edu:/cbica/home/gardnerm/combat-biovar/data/ukb_to_model/*.csv' ./data/ukb_to_model`) to see how values change with different combat methods
```{r}
data.csvs <- list.files(path = "data/ukb_to_model", pattern = ".csv", full.names = TRUE)
combined_data <- data.frame()

for (file in data.csvs) {
  # Read each CSV file
  data <- fread(file)

  # Add a "Source_File" column with the file name
  data <- data %>%
    mutate(Source_File = as.factor(basename(file)))

  # Bind the data to the combined dataframe
  combined_data <- bind_rows(combined_data, data, .id = "File_ID")
}

table(combined_data$Source_File)
```

```{r}
var.df <- combined_data %>%
  group_by(Source_File, sim.site) %>%
  summarize(across(all_of(pheno_list), var))

var.long <- pivot_longer(var.df, cols=c(-Source_File, -sim.site), names_to = "feature", values_to = "variance")
```

```{r}
var.long %>%
  dplyr::filter(!(feature %in% vol_list_global)) %>%
  ggplot() +
  geom_density(aes(x=variance, fill=Source_File), alpha = 0.4, position="identity") +
  facet_wrap(~sim.site)

var.long %>%
  dplyr::filter(!(feature %in% vol_list_global) & (Source_File == "cf.gam_data.csv" |Source_File == "ukb_CN_data_simsites.csv" |  Source_File == "cf.gamlss_data.csv" )) %>%
  ggplot() +
  geom_density(aes(x=variance, fill=Source_File), alpha = 0.5, position="identity") +
  facet_wrap(~sim.site)

#Source_File == "cf.lm_data.csv" | 
```

```{r}
var.by.sex <- combined_data %>%
  group_by(Source_File, sex) %>%
  summarize(across(all_of(pheno_list), var))

var.by.sex.long <- pivot_longer(var.by.sex, cols=c(-Source_File, -sex), names_to = "feature", values_to = "variance")
```

```{r}
var.by.sex.long %>%
  dplyr::filter(!(feature %in% vol_list_global) & (Source_File == "cf.gam_data.csv" | Source_File == "cf.gamlss_data.csv" | Source_File == "ukb_CN_data_simsites.csv")) %>%
  ggplot() +
  geom_density(aes(x=variance, fill=sex), alpha = 0.5, position="identity") +
  facet_wrap(~Source_File)
```

Simple t test for sex effects across combat types
```{r}
sex_ttest_df <- data.frame("Source_File" = levels(combined_data$Source_File))

for(pheno in pheno_list){
  # New column name
  new_col_name <- paste0(pheno, ".p")

  #run t-tests
  out <- lapply(split(combined_data, factor(combined_data$Source_File)), function(x)t.test(x[[pheno]] ~ x[["sex"]], paired=FALSE)$p.value)
  stopifnot(names(out) == levels(combined_data$Source_File)) #make sure results are in correct order

  #append to dataframe
  sex_ttest_df[[new_col_name]] <- unlist(unname(out))
}
```

```{r}
sex_ttest_long <- sex_ttest_df %>% pivot_longer(!Source_File, values_to="p.value", names_to="pheno") %>%
  mutate(pheno = sub(".p$", "", pheno),
         sig_bfcor = case_when(pheno %in% vol_list_global & p.value < (0.05/length(vol_list_global)) ~ TRUE,
                            pheno %in% vol_list_global & p.value >= (0.05/length(vol_list_global)) ~ FALSE,
                            !(pheno %in% vol_list_global) & p.value < (0.05/length(ct_list)) ~ TRUE,
                            !(pheno %in% vol_list_global) & p.value >= (0.05/length(ct_list)) ~ FALSE,
                            TRUE ~ NA))
sex_ttest_long %>%
  group_by(Source_File) %>%
  summarize(n_sig_regions = sum(sig_bfcor),
            n_regions = n())
```

Basically same results gam/gamlss

#### Mu

Does the number of regions w significant age effects on mu change depending on combat version?

```{r}
drop1.df %>%
  dplyr::filter(Term=="poly(age_days, 3)" & Moment == "mu") %>%
  group_by(Dataset) %>%
  summarize(n_sig_regions = sum(sig.drop_bf.corr),
            n_regions = n())
```

No.

Does the number of regions w significant sex effects on mu change depending on combat version?

```{r}
drop1.df %>%
  dplyr::filter(Term=="sexMale" & Moment == "mu") %>%
  group_by(Dataset) %>%
  summarize(n_sig_regions = sum(sig.drop_bf.corr),
            n_regions = n())

sum.df %>%
  dplyr::filter(term=="sexMale" & parameter == "mu") %>%
  group_by(dataset) %>%
  summarize(n_sig_regions = sum(sig.sum_bf.corr),
            n_regions = n())
```

Slightly. Only combat w/o covariates or lm combat without a ref site are noticeably off. Consistent across drop1 and summary outputs

```{r}
sex.df %>%
  dplyr::filter(sig.sum_bf.corr == TRUE & parameter == "mu" & pheno_cat != "Global Volume") %>%
  group_by(dataset, pheno_cat) %>%
  ggplot() +
  geom_brain(atlas = dk, 
             position = position_brain(hemi ~ side),
             aes(fill = t_stat)) +
  scale_fill_gradient2() +
  facet_grid(dataset~pheno_cat) + 
  theme_void() +
  theme(axis.title = element_blank(), axis.text = element_blank(), axis.ticks = element_blank()) +
  ggtitle("t-statistic for effect of male sex on mu")
```

Trying with ultra-strict threshold to see if results change:

```{r}
drop1.df <- drop1.df %>%
  mutate(sig.sum.strict = case_when(drop1.pval < (0.05/length(pheno_list)) ~ TRUE,
                            drop1.pval >= (0.05/length(pheno_list)) ~ FALSE,
                            TRUE ~ NA))

drop1.df %>%
  dplyr::filter(Term=="poly(age_days, 3)" & Moment == "mu") %>%
  group_by(Dataset) %>%
  summarize(n_sig_regions = sum(sig.sum.strict),
            n_regions = n())
```

No difference

#### Sigma

```{r}
sex.df %>%
  dplyr::filter(sig.sum_bf.corr == TRUE & parameter == "sigma" & pheno_cat != "Global Volume") %>%
  ggplot() +
  geom_col(aes(x=label, y=t_stat, fill=dataset)) +
  facet_wrap(~pheno_cat, scales = "free_x") + 
  theme(axis.text.x = element_text(angle = 90, vjust = 0.5, hjust=1)) +
  #theme_void() +
  #theme(axis.title = element_blank(), axis.text = element_blank(), axis.ticks = element_blank()) +
  ggtitle("t-statistic for effect of male sex on sigma")

sex.df %>%
  dplyr::filter(sig.sum_bf.corr == TRUE & parameter == "sigma" & pheno_cat != "Global Volume") %>%
  dplyr::filter(dataset == "cf.lm-data" | dataset == "cf.gamlss-data" | dataset == "ukb-CN-data-simsites") %>%
  ggplot() +
  geom_col(aes(x=label, y=t_stat, fill=dataset), position = "dodge") +
  facet_wrap(~pheno_cat, scales = "free_x") + 
  theme(axis.text.x = element_text(angle = 90, vjust = 0.5, hjust=1)) +
  #theme_void() +
  #theme(axis.title = element_blank(), axis.text = element_blank(), axis.ticks = element_blank()) +
  ggtitle("t-statistic for effect of male sex on sigma")

sex.df %>%
  dplyr::filter(sig.sum_bf.corr == TRUE & parameter == "sigma" & pheno_cat != "Global Volume") %>%
  group_by(dataset, pheno_cat) %>%
  ggplot() +
  geom_brain(atlas = dk, 
             position = position_brain(hemi ~ side),
             aes(fill = t_stat)) +
  scale_fill_gradient2() +
  facet_grid(dataset~pheno_cat) + 
  theme_void() +
  theme(axis.title = element_blank(), axis.text = element_blank(), axis.ticks = element_blank()) +
  ggtitle("t-statistic for effect of male sex on sigma")
```
Check number of regions w significant sex effects in sigma
```{r}
sex.sigma.counts <- sex.df %>%
  dplyr::filter(term=="sexMale" & parameter == "sigma") %>%
  group_by(dataset, pheno_cat) %>%
  summarize(n_sig_regions = sum(sig.drop_bf.corr),
            n_regions = n())
sex.sigma.counts
```

More variation
```{r}
ggplot(sex.sigma.counts) +
  geom_col(aes(x=dataset, y=n_sig_regions, fill=dataset)) +
  facet_wrap(~pheno_cat) +
  theme(axis.text.x = element_text(angle = 90, vjust = 0.5, hjust=1))
```

Convert sigma to variance by correcting for log-link function (sex's effect on variance = exp(sigma) - exp(log(sigma) - beta sex)) -> exp(beta_sex)

```{r}
sigma.df <- sum.df %>%
  dplyr::filter(parameter == "sigma") %>%
  mutate(var_est = exp(estimate))

skim(sigma.df$var_est)

sigma.df %>%
  dplyr::filter(sig.sum_bf.corr == TRUE & term =="sexMale") %>%
  group_by(dataset, pheno_cat) %>%
  ggplot() +
  geom_histogram(aes(x=var_est, fill = dataset)) +
  facet_wrap(~pheno_cat, scales= "free")

sigma.df %>%
  dplyr::filter(sig.sum_bf.corr == TRUE & pheno_cat != "Global Volume" & term =="sexMale") %>%
  group_by(dataset, pheno_cat) %>%
  ggplot() +
  geom_brain(atlas = dk, 
             position = position_brain(hemi ~ side),
             aes(fill = var_est)) +
  #scale_fill_gradient2() +
  facet_grid(dataset~pheno_cat) + 
  theme_void() +
  theme(axis.title = element_blank(), axis.text = element_blank(), axis.ticks = element_blank()) +
  ggtitle("Beta weight for effect of male sex on variance")
```

#### Centiles

Do centile scores change based on combat version?
```{r}
pred.csvs <- list.files(path = "ukb_basic", pattern = "_predictions.csv", full.names = TRUE)
combined_preds <- data.frame()

for (file in pred.csvs) {
  # Read each CSV file
  data <- fread(file)

  # Add a "Source_File" column with the file name
  data <- data %>%
    mutate(Source_File = as.factor(basename(file))) %>%
    mutate(dataset = sub("_data_predictions.csv", "", Source_File))

  # Bind the data to the combined dataframe
  combined_preds <- bind_rows(combined_preds, data, .id = "File_ID")
}

table(combined_preds$dataset)
names(combined_preds)
```
calc how much centiles change
```{r}
cent_diffs <- combined_preds %>%
  mutate(dataset = factor(dataset, levels = unique(dataset), ordered = FALSE)) %>%
  mutate(dataset = relevel(dataset, ref= "ukb_CN")) %>%
  arrange(dataset) %>%
  group_by(participant) %>%
  mutate(across(all_of(pheno_list), ~ . - first(.), .names = "diff_{.col}")) %>%
  ungroup()

diff_vol_glob <- sapply(vol_list_global, function(x) paste0("diff_", x))

mean.cent.diffs <- cent_diffs %>%
  group_by(dataset) %>%
  summarize(across(starts_with("diff_"), mean))

max.cent.diffs <- cent_diffs %>%
  group_by(dataset) %>%
  summarize(across(starts_with("diff_"), max))
```

Plot average differences in centiles
```{r}
mean.diffs.long <- mean.cent.diffs %>%
  pivot_longer(cols=starts_with("diff_"), values_to="centile_diff", names_to = "pheno", names_prefix = "diff_") %>%
  mutate(pheno_cat = as.factor(case_when(
      pheno %in% vol_list_global ~ "Global Volume",
      pheno %in% vol_list_regions ~ "Regional Volume",
      pheno %in% sa_list ~ "Regional SA",
      pheno %in% ct_list ~ "Regional CT",
      TRUE ~ NA_character_)))

ggplot(mean.diffs.long) +
  geom_density(aes(x=centile_diff, fill=dataset), alpha=0.5) +
  facet_grid(dataset~pheno_cat)

mean.diffs.long %>%
  dplyr::filter(dataset== "cf.gam" | dataset == "cf.gam_refA" | dataset == "cf.gamlss" | dataset == "cf.gamlss_refA") %>%
  ggplot() +
  geom_density(aes(x=centile_diff, fill=dataset), alpha=0.5) +
  scale_x_continuous(labels = label_comma())

mean.diffs.long %>%
  dplyr::filter(dataset != "ukb_CN") %>%
  mutate(centile_diff = centile_diff*100) %>% #for visualization purposes
  ggplot() +
  geom_histogram(aes(x=centile_diff, fill=dataset), alpha=0.5) +
  theme(axis.text.x = element_text(angle = 90, vjust = 0.5, hjust=1)) +
  scale_x_continuous(labels = label_comma()) + facet_wrap(~dataset, scales = "free_y")
```

Look at max difference per phenotype
```{r}
max.diffs.long <- max.cent.diffs %>%
  pivot_longer(cols=starts_with("diff_"), values_to="max_centile_diff", names_to = "pheno", names_prefix = "diff_") %>%
  mutate(pheno_cat = as.factor(case_when(
      pheno %in% vol_list_global ~ "Global Volume",
      pheno %in% vol_list_regions ~ "Regional Volume",
      pheno %in% sa_list ~ "Regional SA",
      pheno %in% ct_list ~ "Regional CT",
      TRUE ~ NA_character_)))

max.diffs.long %>%
  #dplyr::filter(dataset== "cf.lm_refA" | dataset == "cf.gam_refA" | dataset == "cf.gamlss" | dataset == "cf.gamlss_refA") %>%
  ggplot() +
  geom_density(aes(x=max_centile_diff, fill=dataset), alpha=0.5) +
  scale_x_continuous(labels = label_comma()) + facet_wrap(~pheno_cat)

max.diffs.long %>%
  dplyr::filter(dataset != "ukb_CN") %>%
  ggplot() +
  geom_density(aes(x=max_centile_diff, fill=dataset), alpha=0.5) +
  scale_x_continuous(labels = label_comma()) + facet_wrap(~dataset, scales = "free_y")
```

Test for differences in distribution
```{r}
summary(aov(centile_diff ~ dataset, data=mean.diffs.long))

attach(mean.diffs.long)
pairwise.t.test(x=centile_diff, g=dataset, p.adj = "bonf")
detach()
```

Seems like gamlss methods are significantly better than the other combat methods and not significantly different from raw centile scores (bonferroni-adjusted)

Do the different combat methods introduce any kind of bias in centile scores?

```{r}
mean.cent.diffs.by.sex <- cent_diffs %>%
  group_by(dataset, sex) %>%
  summarize(across(starts_with("diff_"), mean)) %>%
  ungroup() %>%
  pivot_longer(cols=starts_with("diff_"), values_to="centile_diff", names_to = "pheno", names_prefix = "diff_") %>%
  mutate(pheno_cat = as.factor(case_when(
      pheno %in% vol_list_global ~ "Global Volume",
      pheno %in% vol_list_regions ~ "Regional Volume",
      pheno %in% sa_list ~ "Regional SA",
      pheno %in% ct_list ~ "Regional CT",
      TRUE ~ NA_character_)))

mean.cent.diffs.by.sex %>%
  dplyr::filter(dataset != "ukb_CN") %>%
  mutate(centile_diff = centile_diff *100) %>%
ggplot() +
  geom_density(aes(x=centile_diff, fill=sex), alpha=0.5) +
  facet_wrap(~dataset, scales = "free_y")

max.cent.diffs.by.sex <- cent_diffs %>%
  group_by(dataset, sex) %>%
  summarize(across(starts_with("diff_"), max)) %>%
  ungroup() %>%
  pivot_longer(cols=starts_with("diff_"), values_to="max_centile_diff", names_to = "pheno", names_prefix = "diff_") %>%
  mutate(pheno_cat = as.factor(case_when(
      pheno %in% vol_list_global ~ "Global Volume",
      pheno %in% vol_list_regions ~ "Regional Volume",
      pheno %in% sa_list ~ "Regional SA",
      pheno %in% ct_list ~ "Regional CT",
      TRUE ~ NA_character_)))

max.cent.diffs.by.sex %>%
  dplyr::filter(dataset != "ukb_CN") %>%
  mutate(max_centile_diff = max_centile_diff *100) %>%
ggplot() +
  geom_density(aes(x=max_centile_diff, fill=sex), alpha=0.5) +
  facet_wrap(~dataset, scales = "free_y")
```

Other methods are fairly consistently underestimating male centiles and overestimating female centiles.

#### GAMs
scp'd csvs from cubic (`scp -r 'gardnerm@cubic-login.uphs.upenn.edu:/cbica/home/gardnerm/combat-biovar/ukb_basic/*.csv' ./ukb_basic`)

```{r}
anova.gam <- fread("ukb_basic/anova_gam_tests.csv", stringsAsFactors = TRUE) %>%
  dplyr::select(-V1)
gam.sex.df <- fread("ukb_basic/sex_summary_gam.csv", stringsAsFactors = TRUE) %>%
  dplyr::select(-V1)
gam.sum.df <- fread("ukb_basic/gam_summary.csv", stringsAsFactors = TRUE)%>%
  dplyr::select(-V1)
```

Does the number of regions w significant age effects on mean change depending on combat version?

```{r}
anova.gam %>%
  dplyr::filter(term=="poly(age_days, 3)") %>%
  group_by(dataset) %>%
  summarize(n_sig_regions = sum(sig.drop_bf.corr),
            n_regions = n())
```

No.

```{r, fig.height=10, fig.width=15}
anova.gam %>%
  dplyr::filter(term=="poly(age_days, 3)") %>%
  rename(f_val = "F") %>%
  ggplot() +
  theme(axis.text.x = element_text(angle = 90, vjust = 0.5, hjust=1)) +
  geom_col(aes(x=phenotype, y=f_val, fill=dataset), position = "dodge")
```

```{r}
anova.gam_wide.f <- anova.gam %>%
  dplyr::filter(term=="poly(age_days, 3)") %>%
  rename(f_val = "F") %>%
  pivot_wider(id_cols= phenotype, names_from=dataset, values_from=f_val)

anova.gam.f_diff <- anova.gam_wide.f %>%
  rename_all(~ gsub("-", "_", .)) %>%
  mutate(across(starts_with("cf"), ~ . - ukb_CN_data_simsites)) %>%
  dplyr::select(!ukb_CN_data_simsites)%>%
  pivot_longer(!phenotype, names_to = "dataset", values_to = "f_diff") %>%
  mutate(dataset = as.factor(dataset))

summary(aov(f_diff ~ dataset, data=anova.gam.f_diff))

attach(anova.gam.f_diff)
pairwise.t.test(x=f_diff, g=dataset, p.adj = "bonf")
detach()

ggplot(anova.gam.f_diff) + 
  geom_density(aes(x=f_diff, fill=dataset), alpha=0.5)

anova.gam.f_diff %>%
  dplyr::filter(dataset == "cf.gamlss_data" | dataset == "cf.gam_data" | dataset == "cf.gam_refA_data" | dataset == "cf.gamlss_refA_data") %>%
ggplot() + 
  geom_density(aes(x=f_diff, fill=dataset), alpha=0.5)
```

So the effect sizes (if this F is the correct F to use as effect size) are closer to the "ground truth" for gamlss than non-gamlss combat methods. However, similar to other results, only combat w/o covariates and combat with linear covariates are significantly worse.

Does the number of regions w significant sex effects on mean change depending on combat version?

```{r}
anova.gam %>%
  dplyr::filter(term=="sexMale") %>%
  group_by(dataset) %>%
  summarize(n_sig_regions = sum(sig.drop_bf.corr),
            n_regions = n())
```

Not much, only for combat w/out covariates or linear covariates

```{r, fig.heigt=15}
gam.sum.df %>%
  dplyr::filter(sig.sum_bf.corr == TRUE & term == "sexMale") %>%
  dplyr::filter(dataset == "cf.lm-data" | dataset == "cf.gamlss-data" | dataset == "ukb-CN-data-simsites"| dataset == "cf.lm-refA-data") %>%
  ggplot() +
  geom_col(aes(x=label, y=t_stat, fill=dataset), position="dodge") +
  facet_wrap(~pheno_cat, scales = "free_x") + 
  #theme_cowplot()+
  theme(axis.text.x = element_text(angle = 90, vjust = 0.5, hjust=1)) +
  #theme_void() +
  #theme(axis.title = element_blank(), axis.text = element_blank(), axis.ticks = element_blank()) +
  ggtitle("t-statistic for effect of male sex on mean")

gam.sum.df %>%
  dplyr::filter(sig.sum_bf.corr == TRUE & term == "sexMale") %>%
  ggplot() +
    geom_density(aes(x=t_stat, fill=dataset), alpha=0.4) +
  facet_wrap(~dataset)
```