### 2023-11-02 proof-of-concept_mods.Rmd
plots

---
title: "proof-of-concept_mods"
output: html_document
date: '2023-10-06'
---

## QUESTIONS
- effect size for poly(age_days, 3) in gam/gamlss
- z scores instead of centiles!!!!

## Setup

```{r setup, include=FALSE, warning = FALSE, message = FALSE}
knitr::opts_chunk$set(echo = FALSE, warning = FALSE)
if(!require('pacman')) {
  install.packages('pacman')
}
pacman::p_load(gamlss, gamlss.cens, gamlss.dist, gamlss.mx, gamlss.tr, ggplot2, tidyverse, ggpubr, skimr, gamlss.data, data.table, nlme, devtools, slider, tidymodels, broom.mixed, ggseg, cowplot, lsr, paletteer)

source("./R_scripts/gamlss_helper_funs.R")
```

# Analysis Outline
## Load Data

```{r}
ukb_df_totals <- read_csv(file="data/ukb_CN_data.csv")
ukb_df_age.filt <- read_csv(file="data/ukb_to_model/ukb_CN_data_simsites.csv")

vol_list_global <- readRDS(file="R_scripts/vol_list_global.rds")
vol_list_regions <- readRDS(file="R_scripts/Vol_list_regions.rds")
sa_list <- readRDS(file="R_scripts/SA_list.rds")
ct_list <- readRDS(file="R_scripts/CT_list.rds")
pheno_list <- readRDS(file="R_scripts/pheno_list.rds")
```

## Basic Models
Designate *basic* model formula

```{r}
basic.m1 <- gamlss(formula = GMV ~ poly(age_days, 3) + sexMale,
                       sigma.formula = ~ poly(age_days, 3) + sexMale,
                       nu.formula = ~ 1,
                       family = BCCG, data=ukb_df_age.filt, trace = FALSE)
summary(basic.m1)
```

plot

```{r}
#plot_singlestudy_centiles(basic.m1, "GMV", ukb_df_age.filt, "sexMale")
```

# Review Results

scp'd csvs from cubic (`scp -r 'gardnerm@cubic-login.uphs.upenn.edu:/cbica/home/gardnerm/combat-biovar/ukb_basic/*.csv' ./ukb_basic`)

```{r}
drop1.df <- fread("ukb_basic/drop1_tests.csv", stringsAsFactors = TRUE) %>%
  dplyr::select(-V1)
sex.df <- fread("ukb_basic/sex_summary.csv", stringsAsFactors = TRUE) %>%
  dplyr::select(-V1)
sum.df <- fread("ukb_basic/gamlss_summary.csv", stringsAsFactors = TRUE) %>%
  dplyr::select(-V1) %>%
  mutate(dataset = gsub("-data", "", dataset)) %>%
  mutate(dataset = factor(dataset, levels = unique(dataset), ordered = FALSE)) %>%
  mutate(dataset = relevel(dataset, ref= "ukb-CN-simsites"))
```

```{r}
#add more info from parcellations
dk.parc <- dk$data %>%
  as.data.frame() %>%
  na.omit() %>%
  dplyr::select(c(hemi, region, label)) %>%
  distinct()

sex.df <- left_join(sex.df, dk.parc, by="label")
sum.df <- left_join(sum.df, dk.parc, by="label")
```

### Raw Data Results

```{r, fig.width=11}
sex.df %>%
  dplyr::filter(sig.sum_bf.corr == TRUE & parameter == "sigma" & dataset ==	"ukb-CN-data-simsites") %>%
  ggplot() +
  geom_col(aes(x=reorder(pheno,t_stat), y=t_stat, fill=pheno_cat)) + 
  theme(axis.text.x = element_text(angle = 90, vjust = 0.5, hjust=1)) +
  ggtitle("Significant effects of male sex on sigma across phenotype-specific brain charts")

sex.df %>%
  dplyr::filter(sig.sum_bf.corr == TRUE & parameter == "sigma" & dataset == "ukb-CN-data-simsites") %>%
  ggplot() +
  geom_col(aes(x=label, y=t_stat, fill=pheno_cat)) + 
  theme(axis.text.x = element_text(angle = 90, vjust = 0.5, hjust=1)) +
  ggtitle("Significant effects of male sex on sigma across phenotype-specific brain charts")

sex.df %>%
  dplyr::filter(sig.sum_bf.corr == TRUE & parameter == "sigma" & dataset == "ukb-CN-data-simsites") %>%
  ggplot() +
  geom_col(aes(x=region, y=t_stat, fill=hemi)) + 
  facet_wrap(~pheno_cat) +
  theme(axis.text.x = element_text(angle = 90, vjust = 0.5, hjust=1)) +
  ggtitle("Significant effects of male sex on sigma across phenotype-specific brain charts")
```

Plot on brains!

```{r}
sex.df %>%
  dplyr::filter(sig.sum_bf.corr == TRUE & parameter == "sigma" & dataset == "ukb-CN-data-simsites") %>%
  group_by(pheno_cat) %>%
  ggplot() +
  geom_brain(atlas = dk, 
             position = position_brain(hemi ~ side),
             aes(fill = t_stat)) +
  scale_fill_gradient2() +
  facet_wrap(~pheno_cat) +
  theme_void() +
  theme(axis.title = element_blank(), axis.text = element_blank(), axis.ticks = element_blank()) +
  ggtitle("t-statistic for effect of male sex on sigma")
```

### Combat Effects

#### Site Effects
Loading combatted data (scp'd from cubic: `scp -r 'gardnerm@cubic-login.uphs.upenn.edu:/cbica/home/gardnerm/combat-biovar/data/ukb_to_model/*.csv' ./data/ukb_to_model`) to see how values change with different combat methods
```{r}
data.csvs <- list.files(path = "data/ukb_to_model", pattern = ".csv", full.names = TRUE)
combined_data <- data.frame()

for (file in data.csvs) {
  # Read each CSV file
  data <- fread(file)

  # Add a "Source_File" column with the file name
  data <- data %>%
    mutate(Source_File = as.factor(basename(file)))

  # Bind the data to the combined dataframe
  combined_data <- bind_rows(combined_data, data, .id = "File_ID")
}

table(combined_data$Source_File)
```

```{r}
var.df <- combined_data %>%
  group_by(Source_File, sim.site) %>%
  summarize(across(all_of(pheno_list), var))

var.long <- pivot_longer(var.df, cols=c(-Source_File, -sim.site), names_to = "feature", values_to = "variance")
```

```{r}
var.long %>%
  dplyr::filter(!(feature %in% vol_list_global)) %>%
  ggplot() +
  geom_density(aes(x=variance, fill=Source_File), alpha = 0.4, position="identity") +
  facet_wrap(~sim.site)

var.long %>%
  dplyr::filter(!(feature %in% vol_list_global) & (Source_File == "cf.gam_data.csv" |Source_File == "ukb_CN_data_simsites.csv" |  Source_File == "cf.gamlss_data.csv" )) %>%
  ggplot() +
  geom_density(aes(x=variance, fill=Source_File), alpha = 0.5, position="identity") +
  facet_wrap(~sim.site)

#Source_File == "cf.lm_data.csv" | 
```

```{r}
var.by.sex <- combined_data %>%
  group_by(Source_File, sex) %>%
  summarize(across(all_of(pheno_list), var))

var.by.sex.long <- pivot_longer(var.by.sex, cols=c(-Source_File, -sex), names_to = "feature", values_to = "variance")
```

```{r}
var.by.sex.long %>%
  dplyr::filter(!(feature %in% vol_list_global) & (Source_File == "cf.gam_data.csv" | Source_File == "cf.gamlss_data.csv" | Source_File == "ukb_CN_data_simsites.csv")) %>%
  ggplot() +
  geom_density(aes(x=variance, fill=sex), alpha = 0.5, position="identity") +
  facet_wrap(~Source_File)
```

Simple t test for sex effects across combat types
```{r}
sex_ttest_df <- data.frame("Source_File" = levels(combined_data$Source_File))

for(pheno in pheno_list){
  # New column name
  new_col_name <- paste0(pheno, ".p")

  #run t-tests
  out <- lapply(split(combined_data, factor(combined_data$Source_File)), function(x)t.test(x[[pheno]] ~ x[["sex"]], paired=FALSE)$p.value)
  stopifnot(names(out) == levels(combined_data$Source_File)) #make sure results are in correct order

  #append to dataframe
  sex_ttest_df[[new_col_name]] <- unlist(unname(out))
}
```

```{r}
sex_ttest_long <- sex_ttest_df %>% pivot_longer(!Source_File, values_to="p.value", names_to="pheno") %>%
  mutate(pheno = sub(".p$", "", pheno),
         sig_bfcor = case_when(pheno %in% vol_list_global & p.value < (0.05/length(vol_list_global)) ~ TRUE,
                            pheno %in% vol_list_global & p.value >= (0.05/length(vol_list_global)) ~ FALSE,
                            !(pheno %in% vol_list_global) & p.value < (0.05/length(ct_list)) ~ TRUE,
                            !(pheno %in% vol_list_global) & p.value >= (0.05/length(ct_list)) ~ FALSE,
                            TRUE ~ NA))
sex_ttest_long %>%
  group_by(Source_File) %>%
  summarize(n_sig_regions = sum(sig_bfcor),
            n_regions = n())
```

Basically same results gam/gamlss

#### Mu

Does the number of regions w significant age effects on mu change depending on combat version?

```{r}
drop1.df %>%
  dplyr::filter(Term=="poly(age_days, 3)" & Moment == "mu") %>%
  group_by(Dataset) %>%
  summarize(n_sig_regions = sum(sig.drop_bf.corr),
            n_regions = n())
```

No.

Does the number of regions w significant sex effects on mu change depending on combat version?

```{r}
drop1.df %>%
  dplyr::filter(Term=="sexMale" & Moment == "mu") %>%
  group_by(Dataset) %>%
  summarize(n_sig_regions = sum(sig.drop_bf.corr),
            n_regions = n())

sum.df %>%
  dplyr::filter(term=="sexMale" & parameter == "mu") %>%
  group_by(dataset) %>%
  summarize(n_sig_regions = sum(sig.sum_bf.corr),
            n_regions = n())
```

Slightly. Only combat w/o covariates or lm combat without a ref site are noticeably off. Consistent across drop1 and summary outputs

```{r}
sex.df %>%
  dplyr::filter(sig.sum_bf.corr == TRUE & parameter == "mu" & pheno_cat != "Global Volume") %>%
  group_by(dataset, pheno_cat) %>%
  ggplot() +
  geom_brain(atlas = dk, 
             position = position_brain(hemi ~ side),
             aes(fill = t_stat)) +
  scale_fill_gradient2() +
  facet_grid(dataset~pheno_cat) + 
  theme_void() +
  theme(axis.title = element_blank(), axis.text = element_blank(), axis.ticks = element_blank()) +
  ggtitle("t-statistic for effect of male sex on mu")
```

Trying with ultra-strict threshold to see if results change:

```{r}
drop1.df <- drop1.df %>%
  mutate(sig.sum.strict = case_when(drop1.pval < (0.05/length(pheno_list)) ~ TRUE,
                            drop1.pval >= (0.05/length(pheno_list)) ~ FALSE,
                            TRUE ~ NA))

drop1.df %>%
  dplyr::filter(Term=="poly(age_days, 3)" & Moment == "mu") %>%
  group_by(Dataset) %>%
  summarize(n_sig_regions = sum(sig.sum.strict),
            n_regions = n())
```

No difference

#### Sigma

```{r}
sex.df %>%
  dplyr::filter(sig.sum_bf.corr == TRUE & parameter == "sigma" & pheno_cat != "Global Volume") %>%
  ggplot() +
  geom_col(aes(x=label, y=t_stat, fill=dataset)) +
  facet_wrap(~pheno_cat, scales = "free_x") + 
  theme(axis.text.x = element_text(angle = 90, vjust = 0.5, hjust=1)) +
  #theme_void() +
  #theme(axis.title = element_blank(), axis.text = element_blank(), axis.ticks = element_blank()) +
  ggtitle("t-statistic for effect of male sex on sigma")

sex.df %>%
  dplyr::filter(sig.sum_bf.corr == TRUE & parameter == "sigma" & pheno_cat != "Global Volume") %>%
  dplyr::filter(dataset == "cf.lm-data" | dataset == "cf.gamlss-data" | dataset == "ukb-CN-data-simsites") %>%
  ggplot() +
  geom_col(aes(x=label, y=t_stat, fill=dataset), position = "dodge") +
  facet_wrap(~pheno_cat, scales = "free_x") + 
  theme(axis.text.x = element_text(angle = 90, vjust = 0.5, hjust=1)) +
  #theme_void() +
  #theme(axis.title = element_blank(), axis.text = element_blank(), axis.ticks = element_blank()) +
  ggtitle("t-statistic for effect of male sex on sigma")

sex.df %>%
  dplyr::filter(sig.sum_bf.corr == TRUE & parameter == "sigma" & pheno_cat != "Global Volume") %>%
  group_by(dataset, pheno_cat) %>%
  ggplot() +
  geom_brain(atlas = dk, 
             position = position_brain(hemi ~ side),
             aes(fill = t_stat)) +
  scale_fill_gradient2() +
  facet_grid(dataset~pheno_cat) + 
  theme_void() +
  theme(axis.title = element_blank(), axis.text = element_blank(), axis.ticks = element_blank()) +
  ggtitle("t-statistic for effect of male sex on sigma")
```
Check number of regions w significant sex effects in sigma
```{r}
sex.sigma.counts <- sex.df %>%
  dplyr::filter(term=="sexMale" & parameter == "sigma") %>%
  group_by(dataset, pheno_cat) %>%
  summarize(n_sig_regions = sum(sig.drop_bf.corr),
            n_regions = n())
sex.sigma.counts
```

More variation
```{r}
ggplot(sex.sigma.counts) +
  geom_col(aes(x=dataset, y=n_sig_regions, fill=dataset)) +
  facet_wrap(~pheno_cat) +
  theme(axis.text.x = element_text(angle = 90, vjust = 0.5, hjust=1))
```

Convert sigma to variance by correcting for log-link function (sex's effect on variance = exp(sigma) - exp(log(sigma) - beta sex)) -> exp(beta_sex).

```{r}
sigma.df <- sum.df %>%
  dplyr::filter(parameter == "sigma") %>%
  mutate(var_est = exp(estimate),
         )

skim(sigma.df$var_est)

sigma.df %>%
  dplyr::filter(sig.sum_bf.corr == TRUE & term =="sexMale") %>%
  group_by(dataset, pheno_cat) %>%
  ggplot() +
  geom_histogram(aes(x=var_est, fill = dataset), color="black", alpha=0.5, position = "identity") +
  theme_linedraw() +
  facet_wrap(~pheno_cat, scales= "free")
```

```{r, fig.height=15, fig.width=12}
beta.plot <- sigma.df %>%
  dplyr::filter(sig.sum_bf.corr == TRUE & pheno_cat != "Global Volume" & term =="sexMale") %>%
  group_by(dataset, pheno_cat) %>%
  ggplot() +
  geom_brain(atlas = dk, 
             position = position_brain(hemi ~ side),
             aes(fill = var_est)) +
  #scale_fill_gradient2() +
  facet_grid(dataset~pheno_cat) + 
  theme_bw(base_size=20) +
  theme(axis.title = element_blank(), axis.text = element_blank(), axis.ticks = element_blank(), panel.grid.major = element_blank(), panel.grid.minor = element_blank(), panel.border = element_rect(color= "gray", fill = NA)) +
  paletteer::scale_fill_paletteer_c("grDevices::Oslo") +
  ggtitle("Beta weight for effect of male sex on variance")

ggsave("male_variance_beta.jpeg", beta.plot, width=12, height=15)
```

```{r, fig.height=5, fig.widt=5}
sigma.df %>%
  dplyr::filter(pheno_cat == "Global Volume" & term =="sexMale") %>%
  mutate(var_est.sig = case_when(sig.sum_bf.corr == TRUE ~ var_est,
                                 sig.sum_bf.corr == FALSE ~ 0)) %>%
  #group_by(dataset) %>%
  ggplot() +
  geom_col(aes(x=dataset, y = var_est.sig, fill=dataset), position="dodge") +
  #scale_fill_gradient2() +
  facet_grid(~pheno) + 
  theme_bw(base_size=12) +
  theme(axis.text.x = element_text(angle = 90, vjust = 0.5, hjust=1)) +
  theme(legend.position="none") +
  #theme(axis.title = element_blank(), axis.text = element_blank(), axis.ticks = element_blank(), panel.grid.major = element_blank(), panel.grid.minor = element_blank(), panel.border = element_rect(color= "gray", fill = NA)) +
  ggtitle("Beta weight for effect of male sex on global volumes' variance")

sigma.df.glob <- sigma.df %>%
  dplyr::filter(pheno_cat == "Global Volume" & term =="sexMale")
table(sigma.df.glob$dataset, sigma.df.glob$sig.sum_bf.corr)
```

#### Centiles

Do centile scores change based on combat version?
```{r}
pred.csvs <- list.files(path = "ukb_basic", pattern = "_predictions.csv", full.names = TRUE)
combined_preds <- data.frame()

for (file in pred.csvs) {
  # Read each CSV file
  data <- fread(file)

  # Add a "Source_File" column with the file name
  data <- data %>%
    mutate(Source_File = as.factor(basename(file))) %>%
    mutate(dataset = sub("_data_predictions.csv", "", Source_File))

  # Bind the data to the combined dataframe
  combined_preds <- bind_rows(combined_preds, data, .id = "File_ID")
}

table(combined_preds$dataset)
#skim(combined_preds)
```
calc how much centiles change
```{r}
cent_diffs <- combined_preds %>%
  mutate(dataset = factor(dataset, levels = unique(dataset), ordered = FALSE)) %>%
  mutate(dataset = relevel(dataset, ref= "ukb_CN")) %>%
  arrange(dataset) %>%
  group_by(participant) %>%
  mutate(across(all_of(pheno_list), ~ abs(. - first(.)), .names = "abs.diff_{.col}")) %>%
  ungroup()

diff_vol_glob <- sapply(vol_list_global, function(x) paste0("abs.diff_", x))

mean.cent.diffs <- cent_diffs %>%
  group_by(dataset) %>%
  summarize(across(starts_with("abs.diff_"), mean)) %>%
  ungroup()

max.cent.diffs <- cent_diffs %>%
  group_by(dataset) %>%
  summarize(across(starts_with("abs.diff_"), max)) %>%
  ungroup()
#names(max.cent.diffs)
```

Plot average differences in centiles
```{r}
mean.diffs.long <- mean.cent.diffs %>%
  pivot_longer(cols=starts_with("abs.diff_"), values_to="centile_abs.diff", names_to = "pheno", names_prefix = "abs.diff_") %>%
  mutate(pheno_cat = as.factor(case_when(
      pheno %in% vol_list_global ~ "Global Volume",
      pheno %in% vol_list_regions ~ "Regional Volume",
      pheno %in% sa_list ~ "Regional SA",
      pheno %in% ct_list ~ "Regional CT",
      TRUE ~ NA_character_)),
      label = sub("_[^_]*_", "_", pheno))
mean.diffs.long <- left_join(mean.diffs.long, dk.parc, by="label")
```

```{r}
mean.diffs.long %>%
  dplyr::filter(dataset != "ukb_CN") %>%
ggplot() +
  geom_density(aes(x=centile_abs.diff, fill=dataset), alpha=0.5) +
  facet_grid(dataset~pheno_cat)

mean.diffs.long %>%
  dplyr::filter(dataset== "cf.gam" | dataset == "cf.gam_refA" | dataset == "cf.gamlss" | dataset == "cf.gamlss_refA") %>%
  ggplot() +
  geom_density(aes(x=centile_abs.diff, fill=dataset), alpha=0.5) +
  scale_x_continuous(labels = label_comma())

mean.diffs.long %>%
  dplyr::filter(dataset != "ukb_CN") %>%
  mutate(centile_abs.diff = centile_abs.diff*100) %>% #for visualization purposes
  ggplot() +
  geom_histogram(aes(x=centile_abs.diff, fill=dataset), color="black", binwidth=0.1, alpha=0.5, position="identity") +
  theme_linedraw() +
  #theme(axis.text.x = element_text(angle = 90, vjust = 0.5, hjust=1)) +
  scale_x_continuous(labels = label_comma()) +
  ggtitle("Mean absolute differences in centile scores across ComBat \nprocessing pipelines for each phenotype")

cent.diffs.plot <- mean.diffs.long %>%
  dplyr::filter(dataset != "ukb_CN") %>%
  mutate(centile_abs.diff = centile_abs.diff*100) %>% #for visuals
  ggplot() +
  geom_histogram(aes(x=centile_abs.diff, fill=dataset), color="black", alpha=0.6, position="identity") +
  theme_linedraw(base_size = 12) +
  scale_x_continuous(labels = label_comma()) + facet_wrap(~pheno_cat) +
  paletteer::scale_fill_paletteer_d("Polychrome::light") +
  ggtitle("Mean absolute differences in centile scores across \nComBat processing pipelines for each phenotype")
ggsave("mean_centile_diffs.jpeg", cent.diffs.plot)
```

```{r, fig.height=15, fig.width=12}
mean.diffs.long %>%
  dplyr::filter(pheno_cat != "Global Volume" & dataset != "ukb_CN") %>%
  group_by(dataset, pheno_cat) %>%
  ggplot() +
  geom_brain(atlas = dk, 
             position = position_brain(hemi ~ side),
             aes(fill = centile_abs.diff)) +
  scale_fill_gradient2() +
  facet_grid(dataset~pheno_cat) + 
  theme_void() +
  theme_bw(base_size=20) +
  theme(axis.title = element_blank(), axis.text = element_blank(), axis.ticks = element_blank(), panel.grid.major = element_blank(), panel.grid.minor = element_blank(), panel.border = element_rect(color= "gray", fill = NA)) +
  theme(axis.title = element_blank(), axis.text = element_blank(), axis.ticks = element_blank()) +
  ggtitle("Mean difference in centiles across ComBat configurations")
```

Look at max difference per phenotype
```{r, fig.height=6, fig.width=7}
max.diffs.long <- max.cent.diffs %>%
  pivot_longer(cols=starts_with("abs.diff_"), values_to="max_centile_abs.diff", names_to = "pheno", names_prefix = "abs.diff_") %>%
  mutate(pheno_cat = as.factor(case_when(
      pheno %in% vol_list_global ~ "Global Volume",
      pheno %in% vol_list_regions ~ "Regional Volume",
      pheno %in% sa_list ~ "Regional SA",
      pheno %in% ct_list ~ "Regional CT",
      TRUE ~ NA_character_)))

max.diffs.long %>%
  dplyr::filter(dataset != "ukb_CN") %>%
  mutate(max_centile_abs.diff = max_centile_abs.diff*100) %>% #for visuals
  ggplot() +
  geom_histogram(aes(x=max_centile_abs.diff, fill=dataset), color="black", alpha=0.6, position="identity") +
  theme_linedraw(base_size = 12) +
  scale_x_continuous(labels = label_comma()) + facet_wrap(~pheno_cat) +
  paletteer::scale_fill_paletteer_d("Polychrome::light") +
  ggtitle("Maximum absolute differences in centile scores across \nComBat processing pipelines for each phenotype")
```

```{r}
max.diffs.long %>%
  mutate(max_centile_abs.diff = max_centile_abs.diff*100) %>% #for visuals
  dplyr::filter(dataset != "ukb_CN") %>%
  ggplot() +
  geom_histogram(aes(x=max_centile_abs.diff, fill=dataset), color="black", alpha=0.5, position="identity", binwidth = 0.5) +
  theme_linedraw(base_size = 12) +
  scale_x_continuous(labels = label_comma())+
  ggtitle("Maximum absolute differences in centile scores across \nComBat processing pipelines for each phenotype")
```

Test for differences in distribution
```{r}
summary(aov(centile_abs.diff ~ dataset, data=mean.diffs.long))

attach(mean.diffs.long)
pairwise.t.test(x=centile_abs.diff, g=dataset, p.adj = "fdr")
detach()
```

Seems like gamlss methods are significantly better than the other combat methods and not significantly different from raw centile scores (BH corrected)

Do the different combat methods introduce any kind of bias in centile scores?

```{r}
mean.cent.diffs.by.sex <- cent_diffs %>%
  group_by(dataset, sex) %>%
  summarize(across(starts_with("abs.diff_"), mean)) %>%
  ungroup() %>%
  pivot_longer(cols=starts_with("abs.diff_"), values_to="centile_diff", names_to = "pheno", names_prefix = "abs.diff_") %>%
  mutate(pheno_cat = as.factor(case_when(
      pheno %in% vol_list_global ~ "Global Volume",
      pheno %in% vol_list_regions ~ "Regional Volume",
      pheno %in% sa_list ~ "Regional SA",
      pheno %in% ct_list ~ "Regional CT",
      TRUE ~ NA_character_)))

mean.cent.diffs.by.sex %>%
  dplyr::filter(dataset != "ukb_CN") %>%
  mutate(centile_diff = centile_diff *100) %>%
ggplot() +
  geom_density(aes(x=centile_diff, fill=sex), alpha=0.5) +
  facet_wrap(~dataset, scales = "free_y")

max.cent.diffs.by.sex <- cent_diffs %>%
  group_by(dataset, sex) %>%
  summarize(across(starts_with("abs.diff_"), max)) %>%
  ungroup() %>%
  pivot_longer(cols=starts_with("abs.diff_"), values_to="max_centile_diff", names_to = "pheno", names_prefix = "abs.diff_") %>%
  mutate(pheno_cat = as.factor(case_when(
      pheno %in% vol_list_global ~ "Global Volume",
      pheno %in% vol_list_regions ~ "Regional Volume",
      pheno %in% sa_list ~ "Regional SA",
      pheno %in% ct_list ~ "Regional CT",
      TRUE ~ NA_character_)))

max.cent.diffs.by.sex %>%
  dplyr::filter(dataset != "ukb_CN") %>%
  mutate(max_centile_diff = max_centile_diff *100) %>%
ggplot() +
  geom_density(aes(x=max_centile_diff, fill=sex), alpha=0.5) +
  facet_wrap(~dataset, scales = "free_y")
```

No shift in abs. differences, looking at overall diff.
```{r, fig.height=7, fig.width=8}
cent.diffs2 <- combined_preds %>%
  mutate(dataset = factor(dataset, levels = unique(dataset), ordered = FALSE)) %>%
  mutate(dataset = relevel(dataset, ref= "ukb_CN")) %>%
  arrange(dataset) %>%
  group_by(participant) %>%
  mutate(across(all_of(pheno_list), ~ . - first(.), .names = "diff_{.col}")) %>%
  ungroup()
```

```{r}
max.cent.diffs.by.sex2 <- cent.diffs2 %>%
  group_by(dataset, sex) %>%
  summarize_at(vars(starts_with("diff")), list(max_value = ~.[which.max(abs(.))])) %>%
  ungroup() %>%
  pivot_longer(cols=starts_with("diff_"), values_to="max_centile_diff", names_to = "pheno", names_prefix = "diff_") %>%
  mutate(pheno_cat = as.factor(case_when(
      pheno %in% vol_list_global ~ "Global Volume",
      pheno %in% vol_list_regions ~ "Regional Volume",
      pheno %in% sa_list ~ "Regional SA",
      pheno %in% ct_list ~ "Regional CT",
      TRUE ~ NA_character_)))

max.cent.diffs.by.sex2 %>%
  dplyr::filter(dataset != "ukb_CN") %>%
  mutate(max_centile_diff = max_centile_diff *100) %>%
ggplot() +
  geom_density(aes(x=max_centile_diff, fill=sex), alpha=0.5) +
  facet_wrap(~dataset, scales="free")
```
Other methods are fairly consistently underestimating male centiles and overestimating female centiles.

```{r}
mean.cent.diffs.by.sex2 <- cent.diffs2 %>%
  group_by(dataset, sex) %>%
  summarize(across(starts_with("diff_"), mean)) %>%
  ungroup() %>%
  pivot_longer(cols=starts_with("diff_"), values_to="centile_diff", names_to = "pheno", names_prefix = "diff_") %>%
  mutate(pheno_cat = as.factor(case_when(
      pheno %in% vol_list_global ~ "Global Volume",
      pheno %in% vol_list_regions ~ "Regional Volume",
      pheno %in% sa_list ~ "Regional SA",
      pheno %in% ct_list ~ "Regional CT",
      TRUE ~ NA_character_)))

mean.cent.diffs.by.sex2 %>%
  dplyr::filter(dataset != "ukb_CN") %>%
  mutate(centile_diff = centile_diff *100) %>%
ggplot() +
  geom_density(aes(x=centile_diff, fill=sex), alpha=0.5) +
  facet_wrap(~dataset, scales="free")
```
T test of mean diff. between sex w/in each dataset
```{r}
t.test.by.sex <- mean.cent.diffs.by.sex2 %>%
  dplyr::filter(dataset != "ukb_CN") %>%
  group_by(dataset) %>%
  summarise(t_statistic = t.test(centile_diff ~ sex)$statistic,
            p_value = t.test(centile_diff ~ sex)$p.value) %>%
  mutate(p_val_fdr.corr = p.adjust(p_value, method="fdr", n = length(dataset)),
         fdr_sig = case_when(p_val_fdr.corr < 0.05 ~ TRUE,
                             p_val_fdr.corr >= 0.05 ~ FALSE))

t.test.by.sex
```

Both gam methods and lm combat with ref both end up inducing a bias in how males and females centiles are shifted.

#### Z-Score
calc how much z-scores change

```{r}
z_diffs <- combined_preds %>%
  mutate(dataset = factor(dataset, levels = unique(dataset), ordered = FALSE)) %>%
  mutate(dataset = relevel(dataset, ref = "ukb_CN")) %>%
  arrange(dataset) %>%
  group_by(participant) %>%
  mutate(across(ends_with(".z"), ~ abs(. - first(.)), .names = "abs.diff_{.col}")) %>%
  ungroup()
```

```{r}
#mean
mean.z.diffs <- z_diffs %>%
  group_by(dataset) %>%
  summarize(across(starts_with("abs.diff_"), mean)) %>%
  ungroup()
mean.z.long <- mean.z.diffs %>%
  pivot_longer(cols=starts_with("abs.diff_"), values_to="z_abs.diff", names_to = "pheno", names_prefix = "abs.diff_") %>%
  mutate(pheno = sub(".z", "", pheno)) %>%
  mutate(pheno_cat = as.factor(case_when(
      pheno %in% vol_list_global ~ "Global Volume",
      pheno %in% vol_list_regions ~ "Regional Volume",
      pheno %in% sa_list ~ "Regional SA",
      pheno %in% ct_list ~ "Regional CT",
      TRUE ~ NA_character_)),
      label = sub("_[^_]*_", "_", pheno))
mean.z.long <- left_join(mean.z.long, dk.parc, by="label")

#max
max.z.diffs <- z_diffs %>%
  group_by(dataset) %>%
  summarize(across(starts_with("abs.diff_"), max)) %>%
  ungroup()
max.z.long <- max.z.diffs %>%
  pivot_longer(cols=starts_with("abs.diff_"), values_to="max_z_abs.diff", names_to = "pheno", names_prefix = "abs.diff_") %>%
  mutate(pheno = sub(".z", "", pheno)) %>%
  mutate(pheno_cat = as.factor(case_when(
      pheno %in% vol_list_global ~ "Global Volume",
      pheno %in% vol_list_regions ~ "Regional Volume",
      pheno %in% sa_list ~ "Regional SA",
      pheno %in% ct_list ~ "Regional CT",
      TRUE ~ NA_character_)))
```

```{r}
mean.z.long %>%
  dplyr::filter(dataset != "ukb_CN") %>%
  ggplot() +
  geom_histogram(aes(x=z_abs.diff, fill=dataset), color="black", alpha=0.5, position="identity") +
  theme_linedraw() +
  #theme(axis.text.x = element_text(angle = 90, vjust = 0.5, hjust=1)) +
  scale_x_continuous(labels = label_comma()) +
  ggtitle("Mean absolute differences in Z-scores across ComBat \nprocessing pipelines for each phenotype")
```

```{r, fig.height=15, fig.width=12}
mean.z.long %>%
  dplyr::filter(pheno_cat != "Global Volume" & dataset != "ukb_CN") %>%
  group_by(dataset, pheno_cat) %>%
  ggplot() +
  geom_brain(atlas = dk, 
             position = position_brain(hemi ~ side),
             aes(fill = z_abs.diff)) +
  scale_fill_gradient2() +
  facet_grid(dataset~pheno_cat) + 
  theme_void() +
  theme_bw(base_size=20) +
  theme(axis.title = element_blank(), axis.text = element_blank(), axis.ticks = element_blank(), panel.grid.major = element_blank(), panel.grid.minor = element_blank(), panel.border = element_rect(color= "gray", fill = NA)) +
  theme(axis.title = element_blank(), axis.text = element_blank(), axis.ticks = element_blank()) +
  ggtitle("Mean difference in z scores across ComBat configurations")
```

```{r}
max.z.long %>%
  dplyr::filter(dataset != "ukb_CN") %>%
  ggplot() +
  geom_histogram(aes(x=max_z_abs.diff, fill=dataset), color="black", binwidth = 0.025, alpha=0.5, position="identity") +
  theme_linedraw() +
  #theme(axis.text.x = element_text(angle = 90, vjust = 0.5, hjust=1)) +
  scale_x_continuous(labels = label_comma()) +
  ggtitle("Mean absolute differences in Z-scores across ComBat \nprocessing pipelines for each phenotype")
```

```{r, fig.height=15, fig.width=12}
max.z.long %>%
  dplyr::filter(pheno_cat != "Global Volume" & dataset != "ukb_CN") %>%
  group_by(dataset, pheno_cat) %>%
  ggplot() +
  geom_brain(atlas = dk, 
             position = position_brain(hemi ~ side),
             aes(fill = max_z_abs.diff)) +
  scale_fill_gradient2() +
  facet_grid(dataset~pheno_cat) + 
  theme_void() +
  theme_bw(base_size=20) +
  theme(axis.title = element_blank(), axis.text = element_blank(), axis.ticks = element_blank(), panel.grid.major = element_blank(), panel.grid.minor = element_blank(), panel.border = element_rect(color= "gray", fill = NA)) +
  theme(axis.title = element_blank(), axis.text = element_blank(), axis.ticks = element_blank()) +
  ggtitle("Maximum difference in z scores across ComBat configurations")
```

#### GAMs
scp'd csvs from cubic (`scp -r 'gardnerm@cubic-login.uphs.upenn.edu:/cbica/home/gardnerm/combat-biovar/ukb_basic/*.csv' ./ukb_basic`)

```{r}
anova.gam <- fread("ukb_basic/anova_gam_tests.csv", stringsAsFactors = TRUE) %>%
  dplyr::select(-V1)
gam.sex.df <- fread("ukb_basic/sex_summary_gam.csv", stringsAsFactors = TRUE) %>%
  dplyr::select(-V1)
gam.sum.df <- fread("ukb_basic/gam_summary.csv", stringsAsFactors = TRUE)%>%
  dplyr::select(-V1)
```

Does the number of regions w significant age effects on mean change depending on combat version?

```{r}
anova.gam %>%
  dplyr::filter(term=="poly(age_days, 3)") %>%
  group_by(dataset) %>%
  summarize(n_sig_regions = sum(sig.drop_bf.corr),
            n_regions = n())
```

No.

```{r, fig.height=10, fig.width=15}
anova.gam %>%
  dplyr::filter(term=="poly(age_days, 3)") %>%
  rename(f_val = "F") %>%
  ggplot() +
  theme(axis.text.x = element_text(angle = 90, vjust = 0.5, hjust=1)) +
  geom_col(aes(x=phenotype, y=f_val, fill=dataset), position = "dodge")
```

```{r}
anova.gam_wide.f <- anova.gam %>%
  dplyr::filter(term=="poly(age_days, 3)") %>%
  rename(f_val = "F") %>%
  pivot_wider(id_cols= phenotype, names_from=dataset, values_from=f_val)

anova.gam.f_diff <- anova.gam_wide.f %>%
  rename_all(~ gsub("-", "_", .)) %>%
  mutate(across(starts_with("cf"), ~ . - ukb_CN_data_simsites)) %>%
  dplyr::select(!ukb_CN_data_simsites)%>%
  pivot_longer(!phenotype, names_to = "dataset", values_to = "f_diff") %>%
  mutate(dataset = as.factor(dataset))

summary(aov(f_diff ~ dataset, data=anova.gam.f_diff))

attach(anova.gam.f_diff)
pairwise.t.test(x=f_diff, g=dataset, p.adj = "bonf")
detach()

ggplot(anova.gam.f_diff) + 
  geom_density(aes(x=f_diff, fill=dataset), alpha=0.5)

anova.gam.f_diff %>%
  dplyr::filter(dataset == "cf.gamlss_data" | dataset == "cf.gam_data" | dataset == "cf.gam_refA_data" | dataset == "cf.gamlss_refA_data") %>%
ggplot() + 
  geom_density(aes(x=f_diff, fill=dataset), alpha=0.5)
```

So the effect sizes (if this F is the correct F to use as effect size) are closer to the "ground truth" for gamlss than non-gamlss combat methods. However, similar to other results, only combat w/o covariates and combat with linear covariates are significantly worse.

Does the number of regions w significant sex effects on mean change depending on combat version?

```{r}
anova.gam %>%
  dplyr::filter(term=="sexMale") %>%
  group_by(dataset) %>%
  summarize(n_sig_regions = sum(sig.drop_bf.corr),
            n_regions = n())
```

Not much, only for combat w/out covariates or linear covariates

```{r, fig.heigt=15}
gam.sum.df %>%
  dplyr::filter(sig.sum_bf.corr == TRUE & term == "sexMale") %>%
  dplyr::filter(dataset == "cf.lm-data" | dataset == "cf.gamlss-data" | dataset == "ukb-CN-data-simsites"| dataset == "cf.lm-refA-data") %>%
  ggplot() +
  geom_col(aes(x=label, y=t_stat, fill=dataset), position="dodge") +
  facet_wrap(~pheno_cat, scales = "free_x") + 
  #theme_cowplot()+
  theme(axis.text.x = element_text(angle = 90, vjust = 0.5, hjust=1)) +
  #theme_void() +
  #theme(axis.title = element_blank(), axis.text = element_blank(), axis.ticks = element_blank()) +
  ggtitle("t-statistic for effect of male sex on mean")

gam.sum.df %>%
  dplyr::filter(sig.sum_bf.corr == TRUE & term == "sexMale") %>%
  ggplot() +
    geom_density(aes(x=t_stat, fill=dataset), alpha=0.4) +
  facet_wrap(~dataset)
```