### 2023-12-03 lifespan_eda.Rmd
eda of multisite data

---
title: "lifespan_eda"
output: html_document
date: '2023-11-30'
---

## Setup

```{r setup, include=FALSE, warning = FALSE, message = FALSE}
knitr::opts_chunk$set(echo = TRUE)
if(!require('pacman')) {
  install.packages('pacman')
}
pacman::p_load(gamlss, gamlss.cens, gamlss.mx, gamlss.tr, ggplot2, tidyverse, ggpubr, skimr, gamlss.data, data.table, nlme, devtools, slider, see)

# #pull plotting functions from github
# plot.func.url <- "https://raw.githubusercontent.com/BGDlab/GAMLSS/main/plotting_functions.R?token=GHSAT0AAAAAACGAYP4HZWCTHACGUM6KG2PSZIQUPBA"
# Sys.setenv(GITHUB_PAT = "") #if needed to access private BGDlab GAMLSS repo
# source_url(plot.func.url) #sometimes doesn't work....?

source("/Users/megardn/Desktop/BGD_Repos/GAMLSS/plotting_functions.R")
source("/Users/megardn/Desktop/BGD_Repos/combat_biovar/R_scripts/gamlss_helper_funs.R")
```

Load data from Jakob
```{r}
cn.df <- fread("data/lifespan_dataset_dkatlas_CNonly.csv", stringsAsFactors = TRUE)

#drop repeat scans
cn.df.unique <- cn.df %>%
  drop_na(c(age_days, participant, study)) %>%
  arrange(age_days) %>%
  group_by(participant) %>%
  slice(1) %>%
  ungroup()

nrow(cn.df.unique) == n_unique(cn.df.unique$participant)
```

## Site-Level Data
look just at data w site info

```{r}
cn.df.site <- cn.df.unique %>% 
  dplyr::filter(site != "NA") %>%
  mutate(study = droplevels(study),
         site = gsub(" ", "_", trimws(site)))
n_unique(cn.df.site$participant)
n_unique(cn.df.site$participant) == nrow(cn.df.site)

max(cn.df.site$age_days)/365.25

grep("scan", colnames(cn.df.site), value = TRUE) #no scanner info
```

```{r}
# number studies
length(levels(cn.df.site$study))

#number sites per study
cn.df.site %>%
  group_by(study) %>% 
  summarise(Result = length(unique(site)))
```

```{r}
ggplot(cn.df.site) +
  geom_violin(aes(x=study, y=age_days), scale="width")+
  theme(axis.text.x = element_text(angle = 90, vjust = 0.5, hjust=1))

ggplot(cn.df.site) +
  geom_violin(aes(x=site, y=age_days, fill=study), scale="width") +
  theme(axis.text.x = element_text(angle = 90, vjust = 0.5, hjust=1), legend.position = "none")
```

check for sex differences in sites

```{r}
ggplot(cn.df.site) +
  geom_bar(aes(x=site, fill=sex))+
  theme(axis.text.x = element_text(angle = 90, vjust = 0.5, hjust=1))
```

```{r}
ggplot(cn.df.site) +
  geom_bar(aes( y=..count../tapply(..count.., ..x.. ,sum)[..x..], x=site, fill=sex)) + 
  scale_y_continuous(labels = scales::percent) +
  theme(axis.text.x = element_text(angle = 90, vjust = 0.5, hjust=1))
```

Write out

```{r}
fwrite(cn.df.site, file="data/lifespan_CN_site-level.csv")
```

## Keep Study Level
Keep studies w missing site info and just treat as single-site

```{r}
#clean up site NAs
cn.df.imp <- cn.df.unique %>%
  mutate(site = trimws(as.character(site)),
         study = trimws(study)) %>%
  mutate(site = gsub(" ", "_", ifelse(is.na(site), study, site)))

unique(cn.df.imp$site)
length(is.na(cn.df.imp$site))
max(cn.df.imp$age_days)/365.25
mean(cn.df.imp$age_days)/365.25
```

```{r}
ggplot(cn.df.imp) +
  geom_violin(aes(x=study, y=age_days), scale="width") +
  theme(axis.text.x = element_text(angle = 90, vjust = 0.5, hjust=1))

ggplot(cn.df.imp) +
  geom_violin(aes(x=site, y=age_days, fill=study), scale="width") +
  theme(axis.text.x = element_text(angle = 90, vjust = 0.5, hjust=1), legend.position = "none")
```

check for sex differences in sites

```{r}
ggplot(cn.df.imp) +
  geom_bar(aes(x=site, fill=sex))+
  theme(axis.text.x = element_text(angle = 90, vjust = 0.5, hjust=1))
```

```{r}
ggplot(cn.df.imp) +
  geom_bar(aes( y=..count../tapply(..count.., ..x.. ,sum)[..x..], x=site, fill=sex)) + 
  scale_y_continuous(labels = scales::percent) +
  theme(axis.text.x = element_text(angle = 90, vjust = 0.5, hjust=1))
```

Write out

```{r}
fwrite(cn.df.imp, file="data/lifespan_CN_imp-sites.csv")
```