### 2023-12-05 eda/abide_eda.Rmd
reorg

---
title: "ABIDE EDA"
output: html_document
date: '2023-11-30'
---

## Setup

```{r setup, include=FALSE, warning = FALSE, message = FALSE}
knitr::opts_chunk$set(echo = TRUE)
if(!require('pacman')) {
  install.packages('pacman')
}
pacman::p_load(gamlss, gamlss.cens, gamlss.mx, gamlss.tr, ggplot2, tidyverse, ggpubr, skimr, gamlss.data, data.table, nlme, devtools, slider, see)

# #pull plotting functions from github
# plot.func.url <- "https://raw.githubusercontent.com/BGDlab/GAMLSS/main/plotting_functions.R?token=GHSAT0AAAAAACGAYP4HZWCTHACGUM6KG2PSZIQUPBA"
# Sys.setenv(GITHUB_PAT = "") #if needed to access private BGDlab GAMLSS repo
# source_url(plot.func.url) #sometimes doesn't work....?

source("/Users/megardn/Desktop/BGD_Repos/GAMLSS/plotting_functions.R")
source("/Users/megardn/Desktop/BGD_Repos/combat_biovar/R_scripts/gamlss_helper_funs.R")
```

Load data from Lena
```{r}
abide.df <- fread("data/ABIDE.csv", stringsAsFactors = TRUE) %>%
  dplyr::select(-X) %>%
  janitor::remove_empty(which = "cols") #drop totally empty cols

skim(abide.df)

#check for duplicate subjects
nrow(abide.df) == n_unique(abide.df$participant)
```

restrict to one scan per subject for visualizing demographics

```{r}
abide.df.no_dup <- abide.df %>%
  arrange(age_days) %>%
  group_by(participant) %>%
  slice(1)

nrow(abide.df.no_dup) == n_unique(abide.df.no_dup$participant)

min(abide.df.no_dup$age_days)/365
max(abide.df.no_dup$age_days)/365
```

## Site-wise EDA

look for variance-effecting features that are confounded by site

### Age

```{r}
ggplot(abide.df.no_dup) +
  geom_violin(aes(x=site, y=age_days), scale="width", trim = FALSE, adjust=0.75) +
  theme(axis.text.x = element_text(angle = 90, vjust = 0.5, hjust=1))
```
```{r}
ggplot(abide.df.no_dup) +
  geom_boxplot(aes(x=site, y=age_days)) +
  theme(axis.text.x = element_text(angle = 90, vjust = 0.5, hjust=1))
```

```{r}
ggplot(abide.df.no_dup) +
  geom_violin(aes(x=site, y=age_days, fill=sex), scale="width", trim = FALSE, adjust=0.75) +
  theme(axis.text.x = element_text(angle = 90, vjust = 0.5, hjust=1))
```

### Sex

```{r}
ggplot(abide.df.no_dup) +
  geom_bar(aes( y=..count../tapply(..count.., ..x.. ,sum)[..x..], x=site, fill=sex)) + 
  scale_y_continuous(labels = scales::percent) +
  theme(axis.text.x = element_text(angle = 90, vjust = 0.5, hjust=1))
```

### Diagnosis

```{r}
ggplot(abide.df.no_dup) +
  geom_bar(aes( y=..count../tapply(..count.., ..x.. ,sum)[..x..], x=site, fill=dx)) + 
  scale_y_continuous(labels = scales::percent) +
  theme(axis.text.x = element_text(angle = 90, vjust = 0.5, hjust=1))
```

```{r}
ggplot(abide.df.no_dup) + 
  geom_density(aes(x=lh.CT.frontalpole, fill=dx))
```