### 2023-12-06 paper_figs.Rmd
adding gams for extremeness bias tests

---
title: "ohbm_results"
output: html_document
date: '2023-11-12'
---

```{r setup, include=FALSE}
knitr::opts_chunk$set(echo = FALSE, warning = FALSE)
if(!require('pacman')) {
  install.packages('pacman')
}
pacman::p_load(data.table, ggplot2, tidyverse, ggseg, ggsegExtra, paletteer, ggpubr, gridExtra, lemon, parallel)

source("./cubic_scripts/R_scripts/gamlss_helper_funs.R")
```

Figures for paper

# UKB Basic Models

load info
```{r}
data_path <- "./ukb_basic"

### LOAD ###
pheno_list <- readRDS(file="cubic_scripts/R_scripts/pheno_list.rds")
vol_list_global <- readRDS(file="cubic_scripts/R_scripts/vol_list_global.rds")
vol_list_regions <- readRDS(file="cubic_scripts/R_scripts/Vol_list_regions.rds")
sa_list <- readRDS(file="cubic_scripts/R_scripts/SA_list.rds")
ct_list <- readRDS(file="cubic_scripts/R_scripts/CT_list.rds")

aseg_cerebrum <- readRDS(file="/Users/megardn/Desktop/BGD_Repos/ggseg_atlas/aseg_cerebrum_atlas.RDS")

dataset_names <- c(ukb_CN = "Raw Data", cf = "ComBat w/o Covar", cf.lm = "ComBat", cf.gam ="ComBat-GAM", cf.gamlss ="ComBatLS")
dataset_colors <- c(cf = "#3D5983FF", cf.lm = "#5A8A54FF", cf.gam ="#262626FF", cf.gamlss ="#CFB023FF")

#ukb data from ohbm_results.Rmd
cent_err.p <- readRDS("data/ukb_perm_csvs/centile_errors.RDS")
z_diffs2 <- readRDS("ukb_basic/z_diffs.RDS")
cent.diffs3 <- readRDS("ukb_basic/cent_diffs.RDS")
ukb_data <- fread("data/ukb_to_model/ukb_CN_data_simsites.csv")
#cent_abs.diffs <- readRDS("data/centile_abs_diffs.rds")
```

Big plot - too big to run locally :(

```{r, eval=FALSE}
#test big plot
plot.cents <- cent_abs.diffs %>%
  #dplyr::filter(dataset == "cf.gamlss") %>%
  pivot_longer(starts_with("abs.diff_"), names_to="pheno", values_to="abs.diff", names_prefix = "abs.diff_") %>%
  mutate(pheno_cat = as.factor(case_when(
    pheno %in% vol_list_global ~ "Global Volume",
    pheno %in% vol_list_regions ~ "Regional Volume",
    pheno %in% sa_list ~ "Regional SA",
    pheno %in% ct_list ~ "Regional CT",
    TRUE ~ NA_character_))) %>%
  ggplot() +
  geom_density(aes(x=diff, fill=dataset), alpha=0.4) +
  facet_wrap(~pheno_cat) +
  theme(legend.position="none")
ggsave("figs/cent_plot_test.jpeg", var.plot, width=10, height=10)
```

## Sex Effects on Scale
fix age, get diff between male and female pred. variance
```{r}
pred.csvs <- list.files(path = data_path, pattern = "_variance.csv", full.names = TRUE)
combined_var <- data.frame()

for (file in pred.csvs) {
  # Read each CSV file
  data <- fread(file)

  # Add a "Source_File" column with the file name
  data <- data %>%
    mutate(Source_File = as.factor(basename(file))) %>%
    mutate(dataset = sub("_data_variance.csv", "", Source_File),
           sex_standard = sex_effect/f.var)

  # Bind the data to the combined dataframe
  combined_var <- bind_rows(combined_var, data, .id = "File_ID")
}

#get summary df for p vals, plotting
sigma.sex.df <- fread("ukb_basic/gamlss_summary.csv", stringsAsFactors = TRUE) %>%
  dplyr::select(-V1) %>%
  mutate(dataset = gsub("-data", "", dataset),
         dataset = gsub("-", "_", dataset),
         dataset = gsub("_simsites", "", dataset)) %>% #clean up dataset naming
  mutate(dataset = factor(dataset, levels = unique(dataset), ordered = FALSE)) %>%
  mutate(dataset = relevel(dataset, ref= "ukb_CN")) %>%
  dplyr::filter(parameter == "sigma" & term == "sexMale") %>%
  dplyr::select(mod_name, pheno, dataset, pheno_cat, label, sig.sum_bf.corr)

#add more info from parcellations
dk.parc <- dk$data %>%
  as.data.frame() %>%
  na.omit() %>%
  dplyr::select(c(hemi, region, label)) %>%
  distinct()

sigma.sex.df <- left_join(sigma.sex.df, dk.parc, by="label")

plot.info <- sigma.sex.df %>%
  dplyr::select(mod_name, pheno, dataset, pheno_cat, label, hemi, region, sig.sum_bf.corr) %>%
  unique() %>%
  mutate(dataset = gsub("-", "_", dataset),
         dataset = gsub("_simsites", "", dataset))

combined_var <- inner_join(combined_var, plot.info)
```

```{r}
#var plot
var.plot.df <- combined_var %>%
  dplyr::filter(sig.sum_bf.corr == TRUE & pheno_cat != "Global Volume") %>%
  mutate(dataset = factor(dataset, levels = c("ukb_CN", "cf", "cf.lm", "cf.gam", "cf.gamlss"), ordered = TRUE)) %>%
  group_by(dataset, pheno_cat)

tp <- unique(var.plot.df[,c('dataset','pheno_cat')]) %>%
  dplyr::filter(dataset=="ukb_CN")
tp$sex_standard <- 1

var.plot <- ggplot(var.plot.df) +
  geom_rect(data = tp, fill="gray", xmin = -Inf,xmax = Inf,
            ymin = -Inf, ymax = Inf, alpha = 0.4) +
  ggseg::geom_brain(atlas = dk, 
             position = position_brain(hemi ~ side),
             aes(fill = sex_standard)) +
  #scale_fill_gradient2() +
  facet_grid(dataset ~ pheno_cat, labeller = labeller(dataset = as_labeller(dataset_names))) + 
  theme_linedraw(base_size=12) +
  theme(axis.title = element_blank(), axis.text = element_blank(), axis.ticks = element_blank(), panel.grid.major = element_blank(), panel.grid.minor = element_blank(), panel.border = element_rect(color= "gray", fill = NA), aspect.ratio=2/3, plot.margin = margin(1,1,1,1, 'pt')) +
  paletteer::scale_fill_paletteer_c("grDevices::Cork", name = "Sex Effect", limits = c(-0.12, 0.12), breaks = c(-0.1, 0,0.1))

#ggsave("figs/male_variance_ohbm.jpeg", var.plot, width=10, height=10)
```

```{r}
#plot global
var.plot.global.df <- combined_var %>%
  dplyr::filter(pheno_cat == "Global Volume") %>%
  mutate(dataset = factor(dataset, levels = c("ukb_CN", "cf", "cf.lm", "cf.gam", "cf.gamlss"), ordered = TRUE)) %>%
  dplyr::select(!c(label, region, hemi)) %>% #drop cols corresponding to dk regions
  mutate(tissue_class = as.factor(case_when(pheno == "GMV" ~ "cGM",
                                            pheno == "Ventricles" ~ "CSF",
                                            pheno == "WMV" ~ "WM",
                                            pheno == "sGMV" ~ "sGM",
                                            TRUE ~ NA)),
         sex_standard = ifelse(sig.sum_bf.corr == TRUE, sex_standard, NA)) %>% #drop NS values
  group_by(dataset)

tp <- unique(var.plot.global.df[,c('dataset','pheno_cat')]) %>%
  dplyr::filter(dataset=="ukb_CN")
tp$sex_standard <- 1

var.plot.global <- ggplot(var.plot.global.df) +
  geom_rect(data = tp, fill="gray", xmin = -Inf,xmax = Inf,
            ymin = -Inf, ymax = Inf, alpha = 0.4) +
  ggseg::geom_brain(atlas = aseg_cerebrum, 
             aes(fill = sex_standard),
             side = c("coronal")) +
  #scale_fill_gradient2() +
  facet_grid(dataset~pheno_cat) + 
  theme_linedraw(base_size=12) +
  theme(axis.title = element_blank(), axis.text = element_blank(), axis.ticks = element_blank(), panel.grid.major = element_blank(), panel.grid.minor = element_blank(), panel.border = element_rect(color= "gray", fill = NA), aspect.ratio=, plot.margin = margin(1,1,1,1, 'pt'), strip.text.y = element_blank()) +
  paletteer::scale_fill_paletteer_c("grDevices::Cork", name = "Sex Effect", limits = c(-0.12, 0.12), breaks = c(-0.1, 0,0.1))
```

Combine plots
```{r}
p <- ggarrange(
  var.plot.global, var.plot,
  ncol=2, nrow=1, 
  widths = c(1, 3.1),
  #labels=c("A", "B"),
  #hjust=0.01,
  align = "hv",
  common.legend = TRUE, legend = "bottom"
  )

p_annotate <- annotate_figure(p, top = text_grob("Figure 1. Standardized effect of male sex on feature variance", 
               color = "black", size = 14, hjust=.65))
p_annotate
ggsave("figs/male_variance_ohbm_big.jpeg", p_annotate, width=7, height=8)
```

## Z Errors

```{r}
ukb_sites <- ukb_data %>%
  dplyr::select(participant, sim.site)

z_diffs <- left_join(z_diffs2, ukb_sites) %>%
  dplyr::filter(dataset != "ukb_CN")

ggplot(z_diffs, aes(x=GMV.z, y=diff_GMV.z, color=sex)) +
  geom_point(alpha=0.1, aes(shape=sim.site)) +
  geom_smooth() +
  facet_wrap(~dataset)

ggplot(z_diffs, aes(x=GMV.z, y=diff_GMV.z, color=sex)) +
  geom_point(alpha=0.1, aes(shape=sim.site)) +
  geom_smooth(method="lm") +
  facet_wrap(~dataset)

ggplot(z_diffs, aes(x=WMV.z, y=diff_WMV.z, color=sex)) +
  geom_point(alpha=0.1, aes(shape=sim.site)) +
  geom_smooth(method="lm") +
  facet_wrap(~dataset)

ggplot(z_diffs) +
  geom_jitter(aes(x=sex, y=WMV.z, color=diff_WMV.z), width = 0.2, height = 0, alpha=0.2) +
  #geom_smooth(method="lm") +
  facet_grid(sim.site~dataset) +
  paletteer::scale_color_paletteer_c("grDevices::Cork") +
  theme_bw()

ggplot(z_diffs) +
  geom_jitter(aes(x=sex, y=diff_WMV.z, color=WMV.z), width = 0.2, height = 0, alpha=0.2) +
  #geom_smooth(method="lm") +
  facet_grid(sim.site~dataset) +
  paletteer::scale_color_paletteer_c("grDevices::Cork") +
  theme_bw()

ggplot(z_diffs, aes(x=dataset, y=diff_WMV.z, color=sex)) +
  geom_boxplot() +
  theme_bw() +
  stat_compare_means(aes(group = sex))

ggplot(z_diffs, aes(x=dataset, y=diff_WMV.z, color=sim.site)) +
  geom_boxplot() +
  theme_bw() +
  stat_compare_means(aes(group = sim.site, label = paste0("p = ", after_stat(p.format))))
```

```{r}
ggplot(z_diffs) +
  geom_boxplot(aes(y=diff_WMV.z, x=sim.site, color=sex)) +
  #geom_jitter(aes(y=diff_WMV.z, x=sim.site, color=sex), alpha=0.3) +
  #geom_smooth(method="lm", aes(linetype=sim.site)) +
  facet_wrap(~dataset)
```

## Centile Errors

```{r}
cent_diffs <- left_join(cent.diffs3, ukb_sites) %>%
  dplyr::filter(dataset != "ukb_CN")

ggplot(cent_diffs, aes(x=GMV, y=diff_GMV, color=sex)) +
  geom_point(alpha=0.1, aes(shape=sim.site)) +
  #geom_smooth(method="lm") +
  facet_wrap(~dataset)

ggplot(cent_diffs, aes(x=WMV, y=diff_WMV, color=sex)) +
  geom_point(alpha=0.1, aes(shape=sim.site)) +
  geom_smooth(method="lm") +
  facet_wrap(~dataset)
```
```{r}
ggplot(cent_diffs, aes(x=dataset, y=diff_WMV, color=sex)) +
  geom_boxplot() +
  theme_bw() +
  stat_compare_means(aes(group = sex))
```

## Replication Centiles
W/in subj average plots

```{r}
cent.diffs.v.plot <- cent_err.p %>%
  mutate(mean = mean*100) %>% #for visuals
  ggplot() +
  geom_violin(aes(x=dataset, y=mean, fill=permutation)) +
  labs(y= "Magnitude Centile Error", x = "ComBat Configurations") +
  ggtitle("Within-subject mean magnitude of error in centile scores across ComBat \nconfigurations replicated 10x")

ggsave(filename="figs/centile_violin.png", cent.diffs.v.plot)
```