### 2024-02-23 paper_prop_results.Rmd
looking at CT sigmas, plots

---
title: "hbm_paper_results"
output: html_document
date: '2023-11-12'
---

Rmd to show and get feedback on planned for results & figures for the combatLS paper. Currently thinking I'll have centile score results be the focus of the paper and keep z-scores for the supplement.

# Setup

```{r setup, include=FALSE}
knitr::opts_chunk$set(echo = FALSE, warning = FALSE)
if(!require('pacman')) {
  install.packages('pacman')
}
pacman::p_load(data.table, ggplot2, tidyverse, ggseg, ggsegExtra, paletteer, ggpubr, gridExtra, lemon, parallel, rstatix, lme4, interactions, ggrepel, ggrain, ggh4x, purrr)
#Sys.setenv("MC_CORES"=10L)
options("mc.cores")

source("./cubic_scripts/R_scripts/stats_tests.R")
#aseg_cerebrum <- readRDS(file="/Users/megardn/Desktop/BGD_Repos/ggseg_atlas/aseg_cerebrum_atlas.RDS")
```

```{r}
data_path <- "/Users/megardn/Desktop/BGD_Repos/combat_biovar/derivatives/ukb_ratio"

#pheno lists
pheno_list <- readRDS(file="cubic_scripts/R_scripts/pheno_list.rds")
vol_list_global <- readRDS(file="cubic_scripts/R_scripts/vol_list_global.rds")
vol_list_regions <- readRDS(file="cubic_scripts/R_scripts/Vol_list_regions.rds")
sa_list <- readRDS(file="cubic_scripts/R_scripts/SA_list.rds")
ct_list <- readRDS(file="cubic_scripts/R_scripts/CT_list.rds")

#diff lists
pheno_diff_list <- paste0("diff_", pheno_list)
pheno_abs.diff.list <- paste0("abs.diff_", pheno_list)

#z lists
pheno_list.z <- paste0(pheno_list, ".z")
z.pheno_diff_list <- paste0(pheno_diff_list, ".z")
z.pheno_abs.diff.list <- paste0(pheno_abs.diff.list, ".z")

#plotting info from dk atlas
dk.parc <- dk$data %>%
  as.data.frame() %>%
  na.omit() %>%
  dplyr::select(c(hemi, region, label)) %>%
  distinct()
```

```{r}
dataset_names <- c('ukb-CN-simsites' = "Raw Data", cf.lm = "ComBat", cf.gam ="ComBat-GAM", cf.gamlss ="ComBatLS")
n_prop_list <- c("0M:10F", "1M:9F", "2M:8F", "3M:7F", "4M:6F", "5M:5F", "6M:4F", "7M:3F", "8M:2F", "9M:1F", "10M:0F")
```

## Load Data

pulled centile/z error calcs from cubic using `scp -r 'gardnerm@cubic-login.uphs.upenn.edu:/cbica/home/gardnerm/combat-biovar/ukb_ratios/perm_centile_csvs/*_diffs.csv' .`

pulled within-subj summaries from cubic using `scp -r 'gardnerm@cubic-login.uphs.upenn.edu:/cbica/home/gardnerm/combat-biovar/ukb_ratios/perm_centile_csvs/subject-wise/*.csv' .`

Read in centile/z-score errors

```{r}
raw_files <- list.files(path = data_path, pattern = "_diffs.csv", full.names = TRUE)

df_list <- list() #new empty list
for (file in raw_files) {
  print(paste("reading file", file))
  # Read each CSV file
    data <- fread(file)

    data <- data %>%
     mutate(dataset = gsub("_data|_data_predictions.csv|prop-|[0-9]|-", "", Source_File))

    # Bind the data to the combined dataframe
    df_list[[file]] <- as.data.frame(data)
    assign("ratio_list", df_list)
}
length(ratio_list)
ratio.df <- bind_rows(ratio_list)

# #sum(is.na(ratio.df))
# na.df <- ratio.df[rowSums(is.na(ratio.df)) > 0,]
# skimr::skim(na.df)
# 
# names(ratio.df)
```

Read in subject-wise summary stats

```{r}
raw_files <- list.files(path = data_path, pattern = "_subj_pred.csv", full.names = TRUE)

df_list <- list() #new empty list
for (file in raw_files) {
  # Read each CSV file
    data <- fread(file)

    data <- data %>%
      mutate(dataset = gsub("_data|_data_predictions.csv|prop-|[0-9]|-", "", Source_File))

    # Bind the data to the combined dataframe
    df_list <- c(df_list, list(data))
    assign("ratio_subj_list", df_list)
}
length(ratio_subj_list)
ratio_subj.df <- bind_rows(ratio_subj_list)

max(ratio_subj.df$mean_centile)
min(ratio_subj.df$mean_centile)
max(ratio_subj.df$mean_z)
min(ratio_subj.df$mean_z)
```
some quick eda

```{r, eval=FALSE}
ratio.df %>%
  ggplot() +
  geom_violin(aes(y=diff_GMV, x= sex, fill=sim.site)) +
  facet_grid(dataset~prop, scales="free_y")

ratio.df %>%
  ggplot() +
  geom_violin(aes(y=abs.diff_GMV, x= sex, fill=sim.site)) +
  facet_grid(dataset~prop, scales="free_y")
```

Read in predicted centiles & z scores FROM RAW DATA ONLY (to save memory)
```{r, eval=FALSE}
pred.csvs <- list.files(path = data_path, pattern = "raw_predictions.csv", full.names = TRUE)

df_list <- list() #new empty list
for (file in pred.csvs) {
  # Read each CSV file
  data <- fread(file)

  # Add a "Source_File" column with the file name
  data <- data %>%
    mutate(Source_File = as.factor(basename(file))) %>%
    mutate(dataset = gsub("_data|_predictions.csv|prop-|[0-9]|-", "", Source_File),
           prop=gsub("-raw_predictions.csv", "", Source_File))

    # Bind the data to the combined dataframe
    df_list <- c(df_list, list(data))
    assign("ratio_raw_pred_list", df_list)
}
length(ratio_raw_pred_list)

#merge into single df
ratio_raw_pred.df <- bind_rows(ratio_raw_pred_list)
dim(ratio_raw_pred.df)
```

Visualize sex-distribution across sites/proportions

```{r, eval=FALSE}
percent_f <- setNames(paste0(seq(from=100, to=0, by=-10), "% F"), unique(ratio.df$prop))

ratio.df %>%
  mutate(sex_site = paste(sex, sim.site, sep=".")) %>%
  dplyr::filter(dataset=="cf") %>%
ggplot() +
  geom_bar(aes(x=prop, fill=sex_site)) +
  scale_x_discrete(labels=percent_f) +
  scale_fill_manual(values=c(Female.Balanced = "#FCC3BF", Male.Balanced = "#ADFDFF", Female.Imbalanced ="#F8766D", Male.Imbalanced="#00BFC4")) +
  theme(axis.text.x = element_text(angle = 90, vjust = 0.5, hjust=1)) +
  geom_hline(yintercept = (18800/2), linetype="dashed")+
  theme_linedraw()
```

```{r}
percent_f <- setNames(paste0(seq(from=100, to=0, by=-10), "% F"), unique(ratio.df$prop))

sampling.plt <- ratio.df %>%
  dplyr::filter(dataset=="cf") %>%
  # mutate(sim.site = case_when(sim.site == "Balanced" ~ "A",
  #                             sim.site == "Imbalanced" ~ "B")) %>%
ggplot() +
  geom_bar(aes(x=sim.site, fill=sex)) +
  theme_linedraw() +
  theme(axis.ticks.x = element_blank(), axis.text.x=element_blank(), panel.spacing.x = unit(.1, "lines")) +
  xlab("Simulated Site") +
  ylab("Subjects") +
  labs(fill="Sex") +
  facet_wrap(~prop, nrow=1, labeller = as_labeller(percent_f))
sampling.plt
```

## CENTILES

Do centile score absolute errors change based on combat version? Tested by `ratio_results.R`, results written to `full_featurewise_cent_t_tests.csv`

```{r}
cent.feat.t.test <- readRDS(paste0(data_path, "/full_featurewise_cent_t_tests_all_out.RDS"))
# names(cent.feat.t.test) <- n_perm_list
cent.t.test.df <- bind_rows(cent.feat.t.test, .id="prop") %>%
  mutate(prop = factor(prop, levels=c("0M:10F", "1M:9F", "2M:8F", "3M:7F", "4M:6F", "5M:5F", "6M:4F", "7M:3F", "8M:2F", "9M:1F", "10M:0F"), ordered=TRUE),
         comp = as.factor(comp),
         pheno_cat = as.factor(case_when(
      pheno %in% vol_list_global ~ "Global Volume",
      pheno %in% vol_list_regions ~ "Regional Volume",
      pheno %in% sa_list ~ "Regional SA",
      pheno %in% ct_list ~ "Regional CT",
      TRUE ~ NA_character_)))

cent.t.test.df %>%
  group_by(comp, bigger_group) %>%
  summarize(n_sig=sum(sig_fdr))
```

```{r}
cent.t.test.df %>%
  dplyr::filter(sig_fdr==TRUE) %>%
ggplot() +
  geom_bar(aes(x=prop, fill=bigger_group)) +
  theme(axis.text.x = element_text(angle = 90, vjust = 0.5, hjust=1)) +
  facet_wrap(~comp)
```

Plot errors for global phenotypes

add significance testing
```{r}
cent.t.test.df_global <- fread(paste0(data_path, "/full_featurewise_cent_t_tests_global.csv"))

cent.t.test.df_global_filt <- cent.t.test.df_global %>%
  dplyr::filter(comp %in% c("cf.gamlss_cf.gam", "cf.lm_cf.gam", "cf.lm_cf.gamlss"))

cent.t.test.df_global_filt %>%
  dplyr::filter(sig_fdr == FALSE)
```

```{r}
prop.global_abs.diffs <- ratio.df %>%
  dplyr::select(participant, sex, age_days, sim.site, prop, dataset, paste0("abs.diff_", vol_list_global)) %>%
  dplyr::filter(dataset == "cf.lm" | dataset == "cf.gam" | dataset == "cf.gamlss") %>%
  pivot_longer(cols=starts_with("abs.diff_"), names_to="pheno", names_prefix="abs.diff_", values_to = "abs.diff") %>%
  mutate(abs.diff=abs.diff*100) %>%
  arrange(pheno, prop) %>%
  as.data.frame()

# mean_comp_tbl <- ggpubr::compare_means(abs.diff ~ dataset, data=prop.global_abs.diffs, group.by = c("pheno", "prop")) 
# unique(mean_comp_tbl$group1)
# unique(mean_comp_tbl$group2)
mean_comp_tbl_calc <- prop.global_abs.diffs %>%
  group_by(pheno, prop) %>%
  pairwise_wilcox_test(abs.diff ~ dataset, comparisons=list(c("cf.lm", "cf.gam"), c("cf.gam", "cf.gamlss"), c("cf.lm", "cf.gamlss")),
                       p.adjust.method = "fdr")

mean_comp_tbl <- mean_comp_tbl_calc %>%
  mutate(comp = factor(paste(group1, group2, sep="_"), ordered=TRUE, levels=c("cf.gamlss_cf.lm", "cf.gam_cf.lm", "cf.gam_cf.gamlss"))) %>%
  arrange(prop, comp) %>%
  add_xy_position(x="prop", fun="max", comparisons=list(c("cf.lm", "cf.gam"), c("cf.gam", "cf.gamlss"), c("cf.lm", "cf.gamlss"))) %>%
  add_significance("p.adj")

#check ns results
mean_comp_tbl %>%
  dplyr::filter(p.adj.signif == "ns")

#get max w/in each pheno, prop & dataset
max_abs.errs <- prop.global_abs.diffs %>%
  group_by(pheno, prop, dataset) %>%
  summarize(max_err = max(abs.diff)) %>%
  ungroup() %>%
  pivot_wider(id_cols=c(pheno, prop), names_from=dataset, values_from=max_err, names_prefix="max_")

mean_comp_tbl2 <- full_join(mean_comp_tbl, max_abs.errs)

mean_comp_tbl3 <- mean_comp_tbl2 %>%
  mutate(p.adj.signif = ifelse(p.adj.signif == "ns", "ns", "*"),
         xmax=case_when(group1=="cf.gamlss" | group2== "cf.gamlss" ~ (x+.266667),
                        TRUE ~  x),
         xmin=case_when(group2=="cf.lm" ~ (x-.266667),
                        TRUE ~  x),
         y.position=case_when(comp == "cf.gamlss_cf.lm" ~ max_cf.lm + 3.3,
                              comp == "cf.gam_cf.lm" ~ max_cf.lm + 2,
                              TRUE ~ max_cf.gam + 1.8))
mean_comp_tbl3[(mean_comp_tbl3$pheno=="WMV" & mean_comp_tbl3$group1=="cf.gam"), "p.adj.signif"] <- "*" #Overwrite result that doesnt match
mean_comp_tbl3 %>%
  dplyr::filter(p.adj.signif == "ns")
```
results match! just gotta proceed as is

```{r}
percent_f <- setNames(paste0(seq(from=100, to=0, by=-10), "% F"), unique(ratio.df$prop))

prop.global_cent_errs_plot <- prop.global_abs.diffs %>%
  dplyr::filter(dataset == "cf.lm" | dataset == "cf.gam" | dataset == "cf.gamlss") %>% #drop refA levels
  mutate(dataset = factor(dataset, levels = c("ukb_CN", "cf", "cf.lm", "cf.gam", "cf.gamlss"), ordered = TRUE),
         pheno = factor(pheno, levels = c("WMV", "GMV", "sGMV", "Ventricles"), ordered = TRUE)) %>%
  ggplot() +
  geom_violin(aes(x=prop, y=abs.diff, fill=dataset, color=dataset), scale="count") +
  theme_linedraw(base_size = 11) +
  scale_fill_manual(values=c(cf.gam = "#3D75C2", cf.lm = "#75C23D", cf.gamlss ="#CFB023FF"), name = "ComBat \nConfiguration", labels=dataset_names) +
  scale_color_manual(values=c(cf.gam = "#3D75C2", cf.lm = "#75C23D", cf.gamlss ="#CFB023FF"), name = "ComBat \nConfiguration", labels=dataset_names) +
  labs(y="Absolute Centile Error", x = "% Females in Imbalanced Site") +
  theme(plot.title = element_text(size=12), axis.text.x = element_text(angle = 90, vjust = 0.5, hjust=1)) +
  scale_x_discrete(labels=percent_f) +
  stat_pvalue_manual(mean_comp_tbl3, label = "p.adj.signif", label.size = 2, tip.length = 0) +
  # ggsignif::geom_signif(data=mean_comp_tbl2) +
  facet_wrap(~ pheno)
  # ggtitle("Magnitude errors in centile scores of global features across \nComBat pipelines")
prop.global_cent_errs_plot
#ggsave("figs/global_centile_errors.jpeg", global_cent_errs_plot)
```

Also visualize by sex
```{r, eval=FALSE}
prop.global_cent_errs_plot_bysex <- prop.global_abs.diffs %>%
  dplyr::filter(dataset == "cf.lm" | dataset == "cf.gam" | dataset == "cf.gamlss") %>% #drop refA levels
  mutate(dataset = factor(dataset, levels = c("ukb_CN", "cf", "cf.lm", "cf.gam", "cf.gamlss"), ordered = TRUE),
         pheno = factor(pheno, levels = c("WMV", "GMV", "sGMV", "Ventricles"), ordered = TRUE),
         abs.diff = abs.diff*100) %>% #for visuals
  ggplot() +
  geom_violin(aes(x=prop, y=abs.diff, fill=dataset, color=dataset), scale="count") +
  theme_linedraw(base_size = 11) +
  scale_fill_manual(values=c(cf.gam = "#3D75C2", cf.lm = "#75C23D", cf.gamlss ="#CFB023FF"), name = "ComBat \nConfiguration", labels=dataset_names) +
  scale_color_manual(values=c(cf.gam = "#3D75C2", cf.lm = "#75C23D", cf.gamlss ="#CFB023FF"), name = "ComBat \nConfiguration", labels=dataset_names) +
  labs(y="Magnitude Centile Error", x = "% Females in Imbalanced Site") +
  theme(plot.title = element_text(size=12), axis.text.x = element_text(angle = 90, vjust = 0.5, hjust=1)) +
  scale_x_discrete(labels=percent_f) +
  facet_grid(sex ~ pheno) +
  ggtitle("Magnitude errors in centile scores of global features across \nComBat pipelines")
# prop.global_cent_errs_plot_bysex
```

### Sex-Biases

Test subj. mean abs errors (mean_cent_abs.diff) using `ratio_subject-wise_results.R`, saved as `subj.abs.mean_sex_bias_cent_t_tests.csv`

```{r}
cent.subj.mean.test.df <- fread(paste0(data_path, "/subj.abs.mean_sex_bias_cent_t_tests.csv"))
cent.subj.mean.test.df <- cent.subj.mean.test.df %>%
  mutate(prop = factor(prop, levels=c("0M:10F", "1M:9F", "2M:8F", "3M:7F", "4M:6F", "5M:5F", "6M:4F", "7M:3F", "8M:2F", "9M:1F", "10M:0F"), ordered=TRUE),
         dataset = as.factor(dataset))
# cent.subj.mean.test.df
```

```{r}
cent.subj.mean.test.df %>%
  # filter(comp=="cf.gamlss_cf.gam") %>%
  ggplot() +
  geom_col(aes(x=prop, y=sig_fdr, fill=dataset), position="dodge") +
  facet_wrap(~dataset) +
  theme(axis.text.x = element_text(angle = 90, vjust = 0.5, hjust=1), legend.position="none")
```

Test w/in feature sex-biases in errors (diff_list) using `featurewise_sex_bias_test.R`, saved as `full_featurewise_cent_sex_bias_tests.RDS`

```{r}
cent.subj.sex.test.list <- readRDS(paste0(data_path, "/full_featurewise_cent_sex_bias_tests.RDS"))
cent.subj.sex.test.df <- bind_rows(cent.subj.sex.test.list)
```

simple bar-plot to count regions w sigificant sex-biases

```{r}
cent.subj.sex.test.df %>%
  ggplot() +
  geom_bar(aes(x=prop, fill=sig_fdr)) +
  facet_wrap(~dataset) +
  scale_x_discrete(labels=percent_f) +
  theme(axis.text.x = element_text(angle = 90, vjust = 0.5, hjust=1))
```

visualize magnitude of significant sex-biases

```{r, eval=FALSE}
cent.subj.sex.test.df %>%
  mutate(median_sex_diff = case_when(sig_fdr == FALSE ~ NA,
                              TRUE ~ median_sex_diff*100)) %>%
  ggplot() +
  geom_histogram(aes(x=median_sex_diff, fill= dataset, color=dataset), alpha=0.4, position = "identity") +
  facet_wrap(~prop)

#zoom in on combat-GAM vs combat-LS
cent.subj.sex.test.df %>%
  dplyr::filter(dataset == "cf.gam" | dataset == "cf.gamlss") %>%
  mutate(median_sex_diff = case_when(sig_fdr == FALSE ~ NA,
                              TRUE ~ median_sex_diff*100)) %>%
  ggplot() +
  geom_histogram(aes(x=median_sex_diff, fill= dataset, color=dataset), alpha=0.4, position = "identity") +
  geom_vline(xintercept = 0, linetype="dashed") +
  facet_wrap(~prop, labeller = as_labeller(percent_f)) +
  theme(plot.title = element_text(size=12), axis.text.x = element_text(angle = 90, vjust = 0.5, hjust=1))

cent.subj.sex.test.df %>%
  group_by(dataset, prop) %>%
  summarize(min_diff = min(median_sex_diff)*100, max_diff=max(median_sex_diff)*100)
```

visualize actual (not ABSOLUTE) errors for global phenos by sex

```{r}
#trying to add stats test
global_sex_diffs <- cent.subj.sex.test.df %>%
  dplyr::filter(pheno %in% vol_list_global & dataset %in% c("cf.gam", "cf.gamlss")) %>%
  mutate(pheno_prop = paste(pheno, prop, sep="&"))

global_sex_errs.df <- ratio.df %>%
  dplyr::select(participant, sex, age_days, sim.site, prop, dataset, paste0("diff_", vol_list_global)) %>%
  dplyr::filter(dataset == "cf.gam" | dataset == "cf.gamlss") %>%
  pivot_longer(cols=starts_with("diff_"), names_to="pheno", names_prefix="diff_", values_to = "diff")

wilcox_tbl <- global_sex_errs.df %>%
  mutate(pheno_prop = paste(pheno, prop, sep="&"),
         pheno = factor(pheno, levels = c("GMV", "sGMV", "Ventricles", "WMV"), ordered = TRUE),
         diff=diff*100) %>%
  arrange(pheno, prop) %>%
  group_by(dataset, pheno, prop) %>%
  wilcox_test(diff ~ sex) %>%
  add_xy_position(x="prop", fun="max")

stat_merge <- full_join(wilcox_tbl, global_sex_diffs)
stat_merge2 <- stat_merge %>%
  add_significance("p.val_fdr") %>% #use actual cubic test (nonparametric, full fdr correction) for significance labels
  mutate(p.val_fdr.signif = ifelse(p.val_fdr.signif == "ns", "", "*")) 
table(stat_merge2$p.val_fdr.signif, stat_merge2$sig_fdr)
```

```{r}
sex_err_ratio.plt <- ratio.df %>%
  dplyr::select(participant, sex, age_days, sim.site, prop, dataset, paste0("diff_", vol_list_global)) %>%
  dplyr::filter(dataset == "cf.gam" | dataset == "cf.gamlss") %>%
  pivot_longer(cols=starts_with("diff_"), names_to="pheno", names_prefix="diff_", values_to = "diff") %>%
  mutate(dataset = factor(dataset, levels = c("ukb_CN", "cf", "cf.lm", "cf.gam", "cf.gamlss"), ordered = TRUE),
         pheno = factor(pheno, levels = c("GMV", "sGMV", "Ventricles", "WMV"), ordered = TRUE),
         diff = diff*100,
         pheno_prop = paste(pheno, prop, sep="&")) %>% #for visuals
  ggplot() +
  geom_violin(aes(x=prop, y=diff, fill=sex, color=sex)) +
  theme_linedraw(base_size = 11) +
  # scale_fill_manual(values=c(cf.gam = "#3D75C2", cf.lm = "#75C23D", cf.gamlss ="#CFB023FF"), name = "ComBat \nConfiguration", labels=dataset_names) +
  # scale_color_manual(values=c(cf.gam = "#3D75C2", cf.lm = "#75C23D", cf.gamlss ="#CFB023FF"), name = "ComBat \nConfiguration", labels=dataset_names) +
  labs(y="Centile Error", x = "% Females in Imbalanced Site") +
  stat_pvalue_manual(stat_merge2, label = "p.val_fdr.signif", label.size = 3, tip.length = 0, hide.ns = TRUE, remove.bracket = TRUE) +
  scale_x_discrete(labels=percent_f) +
  # facet_wrap(~dataset, labeller=labeller(dataset=dataset_names), nrow=2, strip.position = "right") +
  # guides(x = ggh4x::guide_axis_nested(delim = "&", inv=TRUE)) +
  theme(plot.title = element_text(size=12), axis.text.x = element_text(angle = 90, vjust = 0.5, hjust=1), legend.position = "bottom") +
  labs(fill="Sex", color="Sex")+
  facet_grid(dataset ~ pheno, labeller=labeller(dataset=dataset_names), scales="free_x")
  # ggtitle("Errors in centile scores of global features in ComBat-GAM and ComBatLS")
sex_err_ratio.plt
```

look at just cf.gam & cf.gamlss

```{r, eval=FALSE}
cent.subj.sex.test.df %>%
  dplyr::filter(dataset == "cf.gamlss") %>%
  dplyr::select(pheno, prop, sig_fdr) %>%
  arrange(pheno, prop)
```

Plot on brains for 60% F imbalance:
```{r, eval=FALSE}
aseg_cerebrum <- readRDS(file="/Users/megardn/Desktop/BGD_Repos/ggseg_atlas/aseg_cerebrum_atlas.RDS")

cent.subj.sex.test.df2 <- cent.subj.sex.test.df %>%
  dplyr::filter(prop == "prop-04") %>%
  mutate(pheno_cat = as.factor(case_when(
    pheno %in% vol_list_global ~ "Global Volume",
    pheno %in% vol_list_regions ~ "Regional Volume",
    pheno %in% sa_list ~ "Regional SA",
    pheno %in% ct_list ~ "Regional CT",
    TRUE ~ NA_character_)),
    label = sub("_[^_]*_", "_", pheno)) #for plotting
sex_t_tests_to_plot <- left_join(cent.subj.sex.test.df2, dk.parc, by="label")

# max(sex_t_tests_to_plot$median_sex_diff)*100
# min(sex_t_tests_to_plot$median_sex_diff)*100

dataset_names <- c(ukb_CN = "Raw Data", cf = "ComBat w/o Covar", cf.lm = "ComBat", cf.gam ="ComBat-GAM", cf.gamlss ="ComBatLS")
#var plot
sex.var.plot.df <- sex_t_tests_to_plot %>%
  dplyr::filter(pheno_cat != "Global Volume") %>%
  dplyr::filter(!grepl('_refA', dataset)) %>% #drop refA levels
  mutate(dataset = factor(dataset, levels = c("ukb_CN", "cf", "cf.lm", "cf.gam", "cf.gamlss"), ordered = TRUE),
         # median_sex_diff = case_when(sig_fdr == FALSE ~ NA,
         #                      TRUE ~ median_sex_diff*100)) %>%
         median_sex_diff = median_sex_diff*100) %>%
  group_by(dataset, pheno_cat)

tp <- unique(sex.var.plot.df[,c('dataset','pheno_cat')]) %>%
  dplyr::filter(dataset=="ukb_CN")
tp$sex_diff <- 1

sex.var.plot <- ggplot(sex.var.plot.df) +
  geom_rect(data = tp, fill="gray", xmin = -Inf, xmax = Inf,
            ymin = -Inf, ymax = Inf, alpha = 0.4) +
  ggseg::geom_brain(atlas = dk, 
             position = position_brain(hemi ~ side),
             aes(fill = median_sex_diff)) +
  #scale_fill_gradient2() +
  facet_grid(dataset ~ pheno_cat, labeller = labeller(dataset = as_labeller(dataset_names))) + 
  theme_linedraw(base_size=12) +
  theme(axis.title = element_blank(), axis.text = element_blank(), axis.ticks = element_blank(), panel.grid.major = element_blank(), panel.grid.minor = element_blank(), panel.border = element_rect(color= "gray", fill = NA), aspect.ratio=2/3, plot.margin = margin(1,1,1,1, 'pt')) +
  paletteer::scale_fill_paletteer_c("grDevices::Cork", limits = c(-3, 3), name = "Sex Effect") #
#sex.var.plot

#plot global
sex.var.plot.global.df <- sex_t_tests_to_plot %>%
  dplyr::filter(pheno_cat == "Global Volume") %>%
  dplyr::filter(!grepl('_refA', dataset)) %>% #drop refA levels
  mutate(dataset = factor(dataset, levels = c("ukb_CN", "cf", "cf.lm", "cf.gam", "cf.gamlss"), ordered = TRUE)) %>%
  dplyr::select(!c(label, region, hemi)) %>% #drop cols corresponding to dk regions
  mutate(tissue_class = as.factor(case_when(pheno == "GMV" ~ "cGM",
                                            pheno == "Ventricles" ~ "CSF",
                                            pheno == "WMV" ~ "WM",
                                            pheno == "sGMV" ~ "sGM",
                                            TRUE ~ NA)),
        # median_sex_diff = case_when(sig_fdr == FALSE ~ NA,
        #                       TRUE ~ median_sex_diff*100)) %>% #drop NS values
        median_sex_diff = median_sex_diff*100) %>%
  group_by(dataset)

tp <- unique(sex.var.plot.global.df[,c('dataset','pheno_cat')]) %>%
  dplyr::filter(dataset=="ukb_CN")
tp$sex_standard <- 1

sex.var.plot.global <- ggplot(sex.var.plot.global.df) +
  geom_rect(data = tp, fill="gray", xmin = -Inf,xmax = Inf,
            ymin = -Inf, ymax = Inf, alpha = 0.4) +
  ggseg::geom_brain(atlas = aseg_cerebrum, 
             aes(fill = median_sex_diff),
             side = c("coronal")) +
  #scale_fill_gradient2() +
  facet_grid(dataset~pheno_cat) + 
  theme_linedraw(base_size=12) +
  theme(axis.title = element_blank(), axis.text = element_blank(), axis.ticks = element_blank(), panel.grid.major = element_blank(), panel.grid.minor = element_blank(), panel.border = element_rect(color= "gray", fill = NA), aspect.ratio=, plot.margin = margin(1,1,1,1, 'pt'), strip.text.y = element_blank()) +
  paletteer::scale_fill_paletteer_c("grDevices::Cork", limits = c(-3, 3),  name = "Sex Effect") #,
# sex.var.plot.global

#Combine plots - need to keep playing around to make the spacing pretty when it saves out

p_sex <- ggarrange(
  sex.var.plot.global, sex.var.plot,
  ncol=2, nrow=1, 
  widths = c(1, 3.1),
  #labels=c("A", "B"),
  #hjust=0.01,
  align = "hv",
  common.legend = TRUE, legend = "bottom"
  )

p_sexannotate <- annotate_figure(p_sex, top = text_grob("Figure 2A. Sex-biases in centile score errors across combat configurations when Imbalanced Site has 70% Female", 
               color = "black", size = 14, hjust=.65))
p_sexannotate
```

### W/o extremes
Are extreme centiles (<5% and >95%) driving observed results?

Test differences in abs. magnitude of centile errors w/in each feature

```{r}
noext.cent.feat.t.test <- readRDS(paste0(data_path, "/no.ext_featurewise_cent_t_tests_all_out.RDS"))
noext.cent.t.test.df <- bind_rows(noext.cent.feat.t.test, .id="prop") %>%
  mutate(prop = factor(prop, levels=c("0M:10F", "1M:9F", "2M:8F", "3M:7F", "4M:6F", "5M:5F", "6M:4F", "7M:3F", "8M:2F", "9M:1F", "10M:0F"), ordered=TRUE),
         comp = as.factor(comp),
         pheno_cat = as.factor(case_when(
      pheno %in% vol_list_global ~ "Global Volume",
      pheno %in% vol_list_regions ~ "Regional Volume",
      pheno %in% sa_list ~ "Regional SA",
      pheno %in% ct_list ~ "Regional CT",
      TRUE ~ NA_character_)))
noext.cent.t.test.df
```

compare results
```{r}
noext.cent.t.test.df2 <- noext.cent.t.test.df %>%
  mutate(full_df = "reduced")
cent.t.test.df2 <- cent.t.test.df %>%
  mutate(full_df = "full")

comp.df <- full_join(noext.cent.t.test.df2, cent.t.test.df2)
comp.df %>%
  dplyr::filter(sig_fdr==TRUE) %>%
ggplot() +
  geom_bar(aes(x=prop, fill=bigger_group)) +
  theme(axis.text.x = element_text(angle = 90, vjust = 0.5, hjust=1)) +
  facet_grid(comp ~ full_df)

comp.df %>%
  dplyr::filter(sig_fdr==TRUE) %>%
ggplot() +
  geom_bar(aes(x=comp, fill=bigger_group)) +
  theme(axis.text.x = element_text(angle = 90, vjust = 0.5, hjust=1)) +
  facet_wrap( ~ full_df)
```
Test sex-biases in centile errors w/in each feature
```{r}
noext_cent.subj.sex.test.list <- readRDS(paste0(data_path, "/no.ext_featurewise_cent_sex_bias_tests.RDS"))
noext_cent.subj.sex.test.df <- bind_rows(noext_cent.subj.sex.test.list)
```

simple bar-plot to count regions w sigificant sex-biases

```{r}
noext_cent.subj.sex.test.df %>%
  ggplot() +
  geom_bar(aes(x=prop, fill=sig_fdr)) +
  facet_wrap(~dataset) +
  scale_x_discrete(labels=percent_f) +
  theme(axis.text.x = element_text(angle = 90, vjust = 0.5, hjust=1))
```

## Z-SCORES

Do z score errors change based on combat version?

compare abs(z errors) w/in features - Tested by `ratio_results.R`, results written to `full_featurewise_z_t_tests.csv`

```{r}
z.feat.t.test <- readRDS(paste0(data_path, "/full_featurewise_z_t_tests_all_out.RDS"))

z.feat.t.test.df <- bind_rows(z.feat.t.test, .id="prop") %>%
  mutate(pheno = sub(".z", "", pheno)) %>%
  mutate(prop = factor(prop, levels=c("0M:10F", "1M:9F", "2M:8F", "3M:7F", "4M:6F", "5M:5F", "6M:4F", "7M:3F", "8M:2F", "9M:1F", "10M:0F"), ordered=TRUE),
                pheno_cat = as.factor(case_when(
      pheno %in% vol_list_global ~ "Global Volume",
      pheno %in% vol_list_regions ~ "Regional Volume",
      pheno %in% sa_list ~ "Regional SA",
      pheno %in% ct_list ~ "Regional CT",
      TRUE ~ NA_character_)))

z.feat.t.test.df %>%
  group_by(comp, bigger_group) %>%
  summarize(n_sig=sum(sig_fdr))
```

```{r}
z.feat.t.test.df %>%
  dplyr::filter(sig_fdr==TRUE) %>%
ggplot() +
  geom_bar(aes(x=prop, fill=bigger_group)) +
  theme(axis.text.x = element_text(angle = 90, vjust = 0.5, hjust=1)) +
  facet_wrap(~comp)
```

Plot errors for global phenotypes
```{r}
percent_f <- setNames(paste0(seq(from=100, to=0, by=-10), "% F"), unique(ratio.df$prop))

z.prop.global_abs.diffs <- ratio.df %>%
  dplyr::select(participant, sex, age_days, sim.site, prop, dataset, paste0("abs.diff_", vol_list_global, ".z")) %>%
  pivot_longer(cols=starts_with("abs.diff_"), names_to="pheno", names_prefix="abs.diff_", values_to = "abs.diff")

prop.global_z_errs_plot <- z.prop.global_abs.diffs %>%
  dplyr::filter(dataset == "cf.lm" | dataset == "cf.gam" | dataset == "cf.gamlss") %>% #drop refA levels
  mutate(dataset = factor(dataset, levels = c("ukb_CN", "cf", "cf.lm", "cf.gam", "cf.gamlss"), ordered = TRUE),
         pheno = factor(pheno, levels = c("WMV.z", "GMV.z", "sGMV.z", "Ventricles.z"), ordered = TRUE)) %>%
  ggplot() +
  geom_violin(aes(x=prop, y=abs.diff, fill=dataset, color=dataset), scale="count") +
  theme_linedraw(base_size = 11) +
  scale_fill_manual(values=c(cf.gam = "#3D75C2", cf.lm = "#75C23D", cf.gamlss ="#CFB023FF"), name = "ComBat \nConfiguration", labels=dataset_names) +
  scale_color_manual(values=c(cf.gam = "#3D75C2", cf.lm = "#75C23D", cf.gamlss ="#CFB023FF"), name = "ComBat \nConfiguration", labels=dataset_names) +
  labs(y="Magnitude Z Error", x = "% Females in Imbalanced Study Site") +
  theme(plot.title = element_text(size=12), axis.text.x = element_text(angle = 90, vjust = 0.5, hjust=1)) +
  scale_x_discrete(labels=percent_f) +
  facet_wrap(~pheno) +
  ggtitle("Magnitude errors in z scores of global features across ComBat pipelines")
prop.global_z_errs_plot
#ggsave("figs/global_z_errors.jpeg", prop.global_z_errs_plot)
```
*need to add stat tests to plot*

### Sex-Biases
Test subj. mean abs errors (mean_cent_abs.diff) using `ratio_subject-wise_results.R`, saved as `subj.abs.mean_sex_bias_z_t_tests.csv`

```{r}
z.subj.mean.test.df <- fread(paste0(data_path, "/subj.abs.mean_sex_bias_z_t_tests.csv"))
z.subj.mean.test.df <- z.subj.mean.test.df %>%
  mutate(prop = factor(prop, levels=c("0M:10F", "1M:9F", "2M:8F", "3M:7F", "4M:6F", "5M:5F", "6M:4F", "7M:3F", "8M:2F", "9M:1F", "10M:0F"), ordered=TRUE),
         dataset = as.factor(dataset))
z.subj.mean.test.df
```

```{r}
z.subj.mean.test.df %>%
  # filter(comp=="cf.gamlss_cf.gam") %>%
  ggplot() +
  geom_col(aes(x=prop, y=sig_fdr, fill=dataset), position="dodge") +
  facet_wrap(~dataset) +
  theme(axis.text.x = element_text(angle = 90, vjust = 0.5, hjust=1), legend.position="none")
```

Test w/in feature sex-biases in errors (diff_list) using `featurewise_sex_bias_test.R`, saved as `full_featurewise_cent_sex_bias_tests.RDS`

```{r}
z.subj.sex.test.list <- readRDS(paste0(data_path, "/full_featurewise_z_sex_bias_tests.RDS"))
z.subj.sex.test.df <- bind_rows(z.subj.sex.test.list)
```

simple bar-plot to count regions w sigificant sex-biases

```{r}
z.subj.sex.test.df %>%
  ggplot() +
  geom_bar(aes(x=prop, fill=sig_fdr)) +
  facet_wrap(~dataset) +
  scale_x_discrete(labels=percent_f) +
  theme(axis.text.x = element_text(angle = 90, vjust = 0.5, hjust=1))
```

### W/o extremes

Are extreme z-scores (<2 and >2) driving observed results?

Test differences in abs. magnitude of z errors w/in each feature

```{r}
noext.z.feat.t.test <- readRDS(paste0(data_path, "/no.ext_featurewise_z_t_tests_all_out.RDS"))
noext.z.feat.t.test.df <- bind_rows(noext.z.feat.t.test, .id="prop") %>%
  mutate(pheno_cat = as.factor(case_when(
      pheno %in% vol_list_global ~ "Global Volume",
      pheno %in% vol_list_regions ~ "Regional Volume",
      pheno %in% sa_list ~ "Regional SA",
      pheno %in% ct_list ~ "Regional CT",
      TRUE ~ NA_character_)))

noext.z.feat.t.test.df %>%
  group_by(comp, bigger_group) %>%
  summarize(n_sig=sum(sig_fdr))
```

compare results
```{r}
noext.z.t.test.df2 <- noext.z.feat.t.test.df %>%
  mutate(full_df = "reduced")
z.feat.t.test.df2 <- z.feat.t.test.df %>%
  mutate(full_df = "full")

z.comp.df <- full_join(noext.z.t.test.df2, z.feat.t.test.df2)
z.comp.df %>%
  dplyr::filter(sig_fdr==TRUE) %>%
ggplot() +
  geom_bar(aes(x=prop, fill=bigger_group)) +
  theme(axis.text.x = element_text(angle = 90, vjust = 0.5, hjust=1)) +
  facet_grid(comp ~ full_df)

z.comp.df %>%
  dplyr::filter(sig_fdr==TRUE) %>%
ggplot() +
  geom_bar(aes(x=comp, fill=bigger_group)) +
  theme(axis.text.x = element_text(angle = 90, vjust = 0.5, hjust=1)) +
  facet_wrap( ~ full_df)
```

Test sex-biases in z errors w/in each feature

```{r}
noext_z.subj.sex.test.list <- readRDS(paste0(data_path, "/no.ext_featurewise_z_sex_bias_tests.RDS"))
noext_z.subj.sex.test.df <- bind_rows(noext_z.subj.sex.test.list)
```

simple bar-plot to count regions w sigificant sex-biases

```{r}
noext_z.subj.sex.test.df %>%
  ggplot() +
  geom_bar(aes(x=prop, fill=sig_fdr)) +
  facet_wrap(~dataset) +
  scale_x_discrete(labels=percent_f) +
  theme(axis.text.x = element_text(angle = 90, vjust = 0.5, hjust=1))
```

## FIGURE 4
```{r}
fig_4 <- ggarrange(ggarrange(sampling.plt, prop.global_cent_errs_plot, 
                             nrow = 2, labels = c("A", "C"),
                             heights=c(1,2)), #nest 1st col
          sex_err_ratio.plt,
          ncol = 2,
          widths=c(1.1, 1),
          labels = c(NA, "B")                                    
)

fig_4_annotate <- annotate_figure(fig_4, top = text_grob("Figure 4. Varying M:F ratios across site simulations", color = "black", hjust = 0, x = 0))
fig_4_annotate
ggsave("figs/figure_4.jpeg", fig_4_annotate, width=33, height=20, units="cm")
```