### 2024-05-17 eda/lifespan_eda.Rmd
cleaning up

---
title: "lifespan_eda"
output: html_document
date: '2023-11-30'
---

## Setup

```{r setup, include=FALSE, warning = FALSE, message = FALSE}
knitr::opts_chunk$set(echo = TRUE)
if(!require('pacman')) {
  install.packages('pacman')
}
pacman::p_load(gamlss, gamlss.cens, gamlss.mx, gamlss.tr, ggplot2, tidyverse, ggpubr, skimr, gamlss.data, data.table, nlme, devtools, slider, see, broom.mixed)

base_path <- "/Users/megardn/Desktop/BGD_Repos/combat_biovar"
vol_list_global <- readRDS(file=paste0(base_path, "/cubic_scripts/R_scripts/vol_list_global.rds"))

source("/Users/megardn/Desktop/BGD_Repos/GAMLSS/plotting_functions.R")
source(paste0(base_path, "/cubic_scripts/R_scripts/gamlss_helper_funs.R"))
```

Load data from Jakob - updated with euler numbers!
```{r}
cn.df <- fread(paste0(base_path,"/data/lifespan_dataset_dkatlas_CNonly_euler.csv"), stringsAsFactors = TRUE)

#drop repeat scans
cn.df.unique <- cn.df %>%
  drop_na(c(age_days, participant, study)) %>%
  arrange(age_days) %>%
  group_by(participant) %>%
  slice(1) %>%
  ungroup() %>%
  mutate(sexMale = as.numeric(case_when(sex == "Male" ~ 1,
                                        sex == "Female" ~ 0,
                                        TRUE ~ NA)),
         sex.age = sexMale*age_days,
         log_age = log(age_days, base=10))

nrow(cn.df.unique) == n_unique(cn.df.unique$participant)
```

```{r}
skim(cn.df.unique)
levels(cn.df.unique$study)
```

euler # noted as surfholes

## Visualize Features
plot global brain features for basic qc - check again post-combat

```{r}
ggplot(cn.df.unique) +
  geom_point(aes(x=age_days, y=GMV, color=sex))

ggplot(cn.df.unique) +
  geom_point(aes(x=age_days, y=sGMV, color=sex))

ggplot(cn.df.unique) +
  geom_point(aes(x=age_days, y=WMV, color=sex))

ggplot(cn.df.unique) +
  geom_point(aes(x=age_days, y=Ventricles, color=sex))
```

log-scaled
```{r}
ggplot(cn.df.unique) +
  geom_point(aes(x=log_age, y=GMV, color=sex))

ggplot(cn.df.unique) +
  geom_point(aes(x=log_age, y=sGMV, color=sex))

ggplot(cn.df.unique) +
  geom_point(aes(x=log_age, y=WMV, color=sex))

ggplot(cn.df.unique) +
  geom_point(aes(x=log_age, y=Ventricles, color=sex))
```

## Curate

Remove infants (anyone younger than 3 yrs)

```{r}
cn.df.noinf <- cn.df.unique %>%
  dplyr::filter(age_days >= (365.25*3)) %>%
  mutate(study = droplevels(study),
         site = droplevels(site))
nrow(cn.df.unique) - nrow(cn.df.noinf)
n_unique(cn.df.unique$study) - n_unique(cn.df.noinf$study)
min(cn.df.noinf$age_days)
```

Visualizing euler distributions
```{r}
cn.df.noinf %>%
  ggplot() +
  geom_boxplot(aes(x=study, y=surfholes))
```

Removing subjects w/o surfholes, subjects with surfholes values > 2 MADs above study median (same as sensitivity analyses in LBCC)

```{r}
cn.df.eulerfilt <- cn.df.noinf %>%
  dplyr::filter(!is.na(surfholes)) #drops 2 studies w missing euler data

study_mad <- cn.df.eulerfilt %>%
  group_by(study) %>%
  summarize(study_euler_mad = mad(surfholes),
            # study_euler_madconst1 = mad(surfholes, constant=1),
            study_euler_median = median(surfholes)) %>%
  mutate(euler_threshold = study_euler_median+(2*study_euler_mad))

left_join(cn.df.eulerfilt, study_mad) %>%
  ggplot() +
  geom_boxplot(aes(x=study, y=surfholes)) +
  geom_point(aes(x=study, y=(study_euler_median+2*study_euler_mad)), color="red")

left_join(cn.df.eulerfilt, study_mad) %>%
  mutate(scan_pass = ifelse(surfholes <= euler_threshold, "yes", "no")) %>%
  group_by(study) %>%
  summarise(fail_percent = (sum(scan_pass == "no") / n())*100) %>%
  summarise(max_fail = max(fail_percent),
            min_fail=min(fail_percent),
            mean_fail=mean(fail_percent))
```

using default mad() constant, looks better

```{r}
cn.df.eulerfilt2 <- left_join(cn.df.eulerfilt, study_mad) %>%
  dplyr::filter(surfholes <= euler_threshold)
```

Removed `r nrow(cn.df.unique) - nrow(cn.df.eulerfilt2)` subjects, `r nrow(cn.df.eulerfilt2)` remain

Keep studies w missing site info and just treat as single-site

```{r}
#clean up site NAs
cn.df.imp1 <- cn.df.eulerfilt2 %>%
  mutate(site = trimws(as.character(site)),
         study = trimws(study)) %>%
  mutate(site = gsub(" ", "_", ifelse(is.na(site), study, paste(study, site, sep="_")))) 
skim(cn.df.imp1)
```

```{r}
#drop NAs listwise
cn.df.imp <- cn.df.imp1%>%
  dplyr::select(!c(INDEX.ID, country)) %>%#drop index ID for filtering
  na.omit()

skim(cn.df.imp)

nrow(cn.df.imp1) - nrow(cn.df.imp)
n_unique(cn.df.imp$site)
n_unique(cn.df.imp$study)
# table(cn.df.imp$site, cn.df.imp$study)
# sum(is.na(cn.df.imp$site))
```

Demographics

```{r}
#sample size, sex
length(unique(cn.df.imp$participant))
nrow(cn.df.imp) == length(unique(cn.df.imp$participant))

table(cn.df.imp$sex)

table(cn.df.imp$sex)["Female"]/nrow(cn.df.imp)

#studies, sites
length(unique(cn.df.imp$study))
length(unique(cn.df.imp$site))

#age
min(cn.df.imp$age_days)/365.25
max(cn.df.imp$age_days)/365.25
mean(cn.df.imp$age_days)/365.25

abcd <- cn.df.imp %>%
  dplyr::filter(study=="ABCD")
length(unique(abcd$site))
```

```{r}
ggplot(cn.df.imp) +
  geom_violin(aes(x=reorder(study, age_days), y=age_days), scale="width") +
  theme(axis.text.x = element_text(angle = 90, vjust = 0.5, hjust=1))

ggplot(cn.df.imp) +
  geom_violin(aes(x=reorder(site, age_days), y=age_days, fill=study), scale="width") +
  theme(axis.text.x = element_text(angle = 90, vjust = 0.5, hjust=1), legend.position = "none")
```

check for sex differences in sites

```{r}
ggplot(cn.df.imp) +
  geom_bar(aes( y=..count../tapply(..count.., ..x.. ,sum)[..x..], x=site, fill=sex)) + 
  scale_y_continuous(labels = scales::percent) +
  theme(axis.text.x = element_text(angle = 90, vjust = 0.5, hjust=1))
```

Write out

```{r}
fwrite(cn.df.imp, file=paste0(base_path, "/data/lifespan_CN_imp-sites_euler.csv"))

#save names with site (since not saved correctly from centile outputs)
study_site_key <- cn.df.imp %>%
  dplyr::select(study, site) %>%
  unique()
n_unique(study_site_key$study) == n_unique(cn.df.imp$study)
n_unique(study_site_key$site) == n_unique(cn.df.imp$site)
fwrite(study_site_key, file=paste0(base_path, "/data/lifespan_study_site_key.csv"))
```