### 2024-05-25 paper_lbcc_results.Rmd
cleaning up

---
title: "hbm_paper_results"
output: html_document
date: '2023-11-12'
---

Rmd to show and get feedback on planned for results & figures for the combatLS paper. Currently thinking I'll have centile score results be the focus of the paper and keep z-scores for the supplement.

# Setup

```{r setup, include=FALSE}
knitr::opts_chunk$set(echo = FALSE, warning = FALSE)
if(!require('pacman')) {
  install.packages('pacman')
}
pacman::p_load(data.table, ggplot2, tidyverse, ggseg, ggsegExtra, paletteer, ggpubr, gridExtra, lemon, parallel, rstatix, lme4, interactions, ggrepel, ggrain, skimr, gam, ggeffects, effectsize, gratia)
#Sys.setenv("MC_CORES"=10L)
options("mc.cores")

source("./cubic_scripts/R_scripts/gamlss_helper_funs.R")
#source("./cubic_scripts/R_scripts/ranked_welch_tests.R")

#ATLAS
aseg_cerebrum <- readRDS(file="/Users/megardn/Desktop/BGD_Repos/ggseg_atlas/aseg_cerebrum_atlas.RDS")
#write out new dk atlas w/o medial wall
dk_nomed <- dk
dk_nomed$data <- dk_nomed$data %>%
  dplyr::filter(!is.na(region) & region !="corpus callosum")
#plotting info from dk atlas
dk.parc <- dk_nomed$data %>%
  as.data.frame() %>%
  na.omit() %>%
  dplyr::select(c(hemi, region, label)) %>%
  distinct()

#PHENO LISTS
pheno_list <- readRDS(file="cubic_scripts/R_scripts/pheno_list.rds")
pheno_list.z <- paste0(pheno_list, ".z")
vol_list_global <- readRDS(file="cubic_scripts/R_scripts/vol_list_global.rds")
vol_list_regions <- readRDS(file="cubic_scripts/R_scripts/Vol_list_regions.rds")
sa_list <- readRDS(file="cubic_scripts/R_scripts/SA_list.rds")
ct_list <- readRDS(file="cubic_scripts/R_scripts/CT_list.rds")

#key to get study names
study_site_key <- fread("/Users/megardn/Desktop/BGD_Repos/combat_biovar/data/lifespan_study_site_key.csv")

#PATH
der_path <- "/Users/megardn/Desktop/BGD_Repos/combat_biovar/derivatives/lifespan"
```

#Load

Combat-GAM Centiles - study-level batch correction w/o infants

```{r}
gam_centile_df <- fread(file=paste0(der_path, "/lifespan-CN-imp-sites-euler-log-cf-cf.gam-batch.study_predictions.csv"))

gam_centile_df <- full_join(gam_centile_df, study_site_key) %>% 
  mutate(age_years=age_days/365.25)
n_unique(gam_centile_df$study)

dim(gam_centile_df)
```

Combat-LS Centiles
```{r}
gamlss_centile_df <- fread(file=paste0(der_path, "/lifespan-CN-imp-sites-euler-log-cf-cf.gamlss-batch.study_predictions.csv"))
gamlss_centile_df <- full_join(gamlss_centile_df, study_site_key) %>% 
  mutate(age_years=age_days/365.25)

dim(gamlss_centile_df)
```

```{r}
min(gam_centile_df$age_years) == min(gamlss_centile_df$age_years)
min(gamlss_centile_df$age_years)
```

Unharmonized Centiles
```{r}
raw_centile_df <- fread(file=paste0(der_path, "/lifespan-CN-imp-sites-euler-raw_predictions.csv"))
raw_centile_df <- full_join(raw_centile_df, study_site_key) %>% 
  mutate(age_years=age_days/365.25)

dim(raw_centile_df)
min(raw_centile_df$age_years) == min(gamlss_centile_df$age_years)
```

Deltas

```{r}
deltas_df <- fread(file=paste0(der_path, "/batch.study_pred_deltas.csv"))
deltas_df <- full_join(deltas_df, study_site_key) %>% 
  mutate(age_years=age_days/365.25)

dim(deltas_df)
min(deltas_df$age_years) == min(gamlss_centile_df$age_years)
```

Nature paper models to look at age effects in variance
from [https://github.com/brainchart/Lifespan/tree/main/Share](https://github.com/brainchart/Lifespan/tree/main/Share)

```{r}
paper_fit_sgmv <- readRDS("derivatives/nature_2020_mods/FIT_sGMV.rds")
paper_fit_gmv <- readRDS("derivatives/nature_2020_mods/FIT_GMV.rds")
paper_fit_wmv <- readRDS("derivatives/nature_2020_mods/FIT_WMV.rds")
paper_fit_ven <- readRDS("derivatives/nature_2020_mods/FIT_Ventricles.rds")
```

#Clean & EDA

Check numbers
```{r}
nrow(gamlss_centile_df) #n
n_unique(gamlss_centile_df$study) 
n_unique(gamlss_centile_df$site) #correct
```

Sample plot
```{r}
#reorder studies by age
age.df <- gamlss_centile_df %>%
  group_by(study) %>%
  summarize(mean_age = mean(age_years)) %>%
  arrange(-mean_age)

sample_boxplot <- gamlss_centile_df %>%
  mutate(study = factor(study, levels=c(age.df$study), ordered=TRUE)) %>%
ggplot(aes(y=study, x=age_years)) +
  geom_boxplot(outlier.size = 1, outlier.colour=NULL) +
  theme_linedraw(base_size = 11) +
  #guides(y = ggh4x::guide_axis_nested(delim = "&")) +
  theme(axis.title.x=element_blank(), axis.ticks.y = element_blank(), axis.text.x=element_blank()) +
  labs(color="Sex", y = "Study")

#overall sample
full_boxplot <- 
  gamlss_centile_df %>%
  mutate(ylab = "   Full Sample") %>%
  ggplot(aes(y=ylab, x=age_years, color=sex)) +
  geom_boxplot(outlier.size = 1, outlier.colour=NULL) +
  theme_linedraw(base_size = 11) +
  # theme(plot.margin = margin(0,0,0,0, 'cm')) +
  # scale_y_discrete(labels=c("1" = "Full Sample")) +
  labs(color="Sex", x = "Age (Years)", y="")

sample_barplot <- gamlss_centile_df %>%
  mutate(study = factor(study, levels=c(age.df$study), ordered=TRUE)) %>%
  ggplot() +
  geom_bar(aes( x=..count../tapply(..count.., ..y.. ,sum)[..y..], y=study, fill=sex)) +
  scale_x_continuous(labels = scales::percent) +
  geom_vline(xintercept = .5, color="lightgray", linetype="longdash") +
  theme_linedraw(base_size = 11) +
  theme(axis.ticks.y = element_blank(), axis.text=element_blank(), axis.title=element_blank()) +
  labs(fill="Sex")

full_barplot <- 
  gamlss_centile_df %>%
  mutate(ylab = "Full Sample") %>%
  ggplot() +
  geom_bar(aes(x=..count../tapply(..count.., ..y.. ,sum)[..y..], y=ylab, fill=sex)) + 
  scale_x_continuous(labels = scales::percent) +
  geom_vline(xintercept = .5, color="lightgray", linetype="longdash") +
  theme_linedraw(base_size = 11) +
  theme(axis.ticks.y = element_blank(), axis.text.y=element_blank(), axis.title.y=element_blank()) +
  labs(fill="Sex", x = "Sex")

ggarrange(sample_boxplot, sample_barplot,
          full_boxplot, full_barplot,
          nrow=2, ncol=2, 
          align = "hv", 
          common.legend = TRUE,
          widths=c(2, 1),
          heights=c(10, 1))
```

#Batch Effects

Test to see if centile score is associated with study in either ComBatLS, ComBat-GAM, or non-harmonized data.

## Unharmonized
```{r}
#clean predictions
raw_centile_df.clean <- full_join(raw_centile_df, study_site_key) %>% 
  mutate(age_years=age_days/365.25) %>%
  dplyr::select(!contains(".z") & !ends_with("_total") & !contains("TBV")) #drop z-scores and non-centiles

skim(raw_centile_df.clean)
```

w/in feature ANOVAs

```{r}
raw.feat_to_test <- intersect(pheno_list, names(raw_centile_df.clean))
length(raw.feat_to_test)

df_raw_feat_aov <- data.frame()
for(pheno in raw.feat_to_test){
  # print(pheno)
  df <- tidy(aov(raw_centile_df.clean[[pheno]] ~ raw_centile_df.clean[["study"]]))
  df$pheno <- pheno
  df_raw_feat_aov <- rbind(df_raw_feat_aov, df)
}

df_raw_feat_aov.study <- df_raw_feat_aov %>%
  group_by(pheno) %>%
  mutate(etasq = sumsq / sum(sumsq)) %>%
  ungroup() %>%
  dplyr::filter(term != "Residuals") %>%
  dplyr::mutate(p.val_fdr = p.adjust(p.value, method="fdr", n = length(raw.feat_to_test)), #apply fdr correction across all phenos tested
                  sig_fdr = case_when(p.val_fdr < 0.05 ~ TRUE,
                                      p.val_fdr >= 0.05 ~ FALSE))
df_raw_feat_aov.study %>%
  ggplot() +
  geom_bar(aes(y=sig_fdr))

table(df_raw_feat_aov.study$sig_fdr)
```
```{r}
raw_centile_df.mean <- raw_centile_df.clean %>%
  pivot_longer(cols=any_of(pheno_list), names_to="pheno", values_to = "centile") %>%
  group_by(participant) %>%
  summarize(study = unique(study),
            site = unique(site),
            age_days=unique(age_days),
            sex=unique(sex),
            mean_cent=mean(centile))
nrow(raw_centile_df.mean) == n_unique(raw_centile_df.mean$participant)
```

```{r}
raw_centile_df.mean %>%
  ggplot() +
  geom_boxplot(aes(x=reorder(study, age_days), y=mean_cent)) + 
  #geom_boxplot(aes(x=study, y=mean_cent)) +
  theme(axis.text.x = element_text(angle=90, hjust=0.5))

raw_centile_df.mean %>%
  ggplot() +
  geom_point(aes(x=age_days/365.25, y=mean_cent, color=site), alpha=0.5) +
  geom_smooth(aes(x=age_days/365.25, y=mean_cent), method="lm") +
  theme(legend.position = "none") +
  facet_wrap(~study)
```

ANOVA of subjects' mean centiles across studies

```{r}
summary(aov(mean_cent ~ study, data=raw_centile_df.mean))
```

###Z-Scores

```{r}
#clean predictions
raw_z_df.clean <- full_join(raw_centile_df, study_site_key) %>% 
  mutate(age_years=age_days/365.25) %>%
  dplyr::select(!any_of(pheno_list) & !ends_with("_total") & !contains("TBV")) #drop z-scores and non-centiles
```

w/in feature ANOVAs

```{r}
raw.feat_to_test.z <- paste0(raw.feat_to_test, ".z")
df_raw_feat_aov.z <- data.frame()
for(pheno in raw.feat_to_test.z){
  # print(pheno)
  df <- tidy(aov(raw_z_df.clean[[pheno]] ~ raw_z_df.clean[["study"]]))
  df$pheno <- pheno
  df_raw_feat_aov.z <- rbind(df_raw_feat_aov.z, df)
}

df_raw_feat_aov.z.study <- df_raw_feat_aov.z %>%
  dplyr::filter(term != "Residuals") %>%
  dplyr::mutate(p.val_fdr = p.adjust(p.value, method="fdr", n = length(raw.feat_to_test.z)), #apply fdr correction across all phenos tested
                  sig_fdr = case_when(p.val_fdr < 0.05 ~ TRUE,
                                      p.val_fdr >= 0.05 ~ FALSE))
df_raw_feat_aov.z.study %>%
  ggplot() +
  geom_bar(aes(y=sig_fdr))

table(df_raw_feat_aov.z.study$sig_fdr)
```

average w/in subject

```{r}
raw_z_df.mean <- raw_z_df.clean %>%
  pivot_longer(cols=any_of(pheno_list.z), names_to="pheno", values_to = "z_score") %>%
  group_by(participant) %>%
  summarize(study = unique(study),
            site = unique(site),
            age_days=unique(age_days),
            sex=unique(sex),
            mean_z=mean(z_score))
nrow(raw_z_df.mean) == n_unique(raw_z_df.clean$participant)
```

ANOVA of subject means

```{r}
summary(aov(mean_z ~ study, data=raw_z_df.mean))
```

## ComBat-GAM

```{r}
#clean predictions
gam_centile_df.clean <- full_join(gam_centile_df, study_site_key) %>% 
  mutate(age_years=age_days/365.25)%>%
  dplyr::select(!contains(".z") & !ends_with("_total") & !contains("TBV")) #drop z-scores and non-centiles

skim(gam_centile_df.clean)
```

w/in feature ANOVAs

```{r}
gam.feat_to_test <- intersect(pheno_list, names(gam_centile_df.clean))
length(gam.feat_to_test)

df_gam_feat_aov <- data.frame()
for(pheno in gam.feat_to_test){
  # print(pheno)
  df <- tidy(aov(gam_centile_df.clean[[pheno]] ~ gam_centile_df.clean[["study"]]))
  df$pheno <- pheno
  df_gam_feat_aov <- rbind(df_gam_feat_aov, df)
}

df_gam_feat_aov.study <- df_gam_feat_aov %>%
  group_by(pheno) %>%
  mutate(etasq = sumsq / sum(sumsq)) %>%
  ungroup() %>%
  dplyr::filter(term != "Residuals") %>%
  dplyr::mutate(p.val_fdr = p.adjust(p.value, method="fdr", n = length(gam.feat_to_test)), #apply fdr correction across all phenos tested
                  sig_fdr = case_when(p.val_fdr < 0.05 ~ TRUE,
                                      p.val_fdr >= 0.05 ~ FALSE))
df_gam_feat_aov.study %>%
  ggplot() +
  geom_bar(aes(y=sig_fdr))

table(df_gam_feat_aov.study$sig_fdr)
```

average w/in subject

```{r}
gam_centile_df.mean <- gam_centile_df.clean %>%
  pivot_longer(cols=any_of(pheno_list), names_to="pheno", values_to = "centile") %>%
  group_by(participant) %>%
  summarize(study = unique(study),
            site = unique(site),
            age_days=unique(age_days),
            sex=unique(sex),
            mean_cent=mean(centile))
nrow(gam_centile_df.mean) == n_unique(gam_centile_df.mean$participant)
```

```{r}
gam_centile_df.mean %>%
  ggplot() +
  geom_boxplot(aes(x=reorder(study, age_days), y=mean_cent)) + 
  #geom_boxplot(aes(x=study, y=mean_cent)) +
  theme(axis.text.x = element_text(angle=90, hjust=0.5))

gam_centile_df.mean %>%
  ggplot() +
  geom_point(aes(x=age_days/365.25, y=mean_cent, color=site), alpha=0.5) +
  geom_smooth(aes(x=age_days/365.25, y=mean_cent), method="lm") +
  theme(legend.position = "none") +
  facet_wrap(~study)
```

ANOVA

```{r}
summary(aov(mean_cent ~ study, data=gam_centile_df.mean))
```

###Z-Scores

```{r}
#clean predictions
gam_z_df.clean <- full_join(gam_centile_df, study_site_key) %>% 
  mutate(age_years=age_days/365.25) %>%
  dplyr::select(!any_of(pheno_list) & !ends_with("_total") & !contains("TBV")) #drop z-scores and non-centiles
```

w/in feature ANOVAs

```{r}
gam.feat_to_test.z <- paste0(gam.feat_to_test, ".z")
df_gam_feat_aov.z <- data.frame()
for(pheno in gam.feat_to_test.z){
  # print(pheno)
  df <- tidy(aov(gam_z_df.clean[[pheno]] ~ gam_z_df.clean[["study"]]))
  df$pheno <- pheno
  df_gam_feat_aov.z <- rbind(df_gam_feat_aov.z, df)
}

df_gam_feat_aov.z.study <- df_gam_feat_aov.z %>%
  dplyr::filter(term != "Residuals") %>%
  dplyr::mutate(p.val_fdr = p.adjust(p.value, method="fdr", n = length(gam.feat_to_test.z)), #apply fdr correction across all phenos tested
                  sig_fdr = case_when(p.val_fdr < 0.05 ~ TRUE,
                                      p.val_fdr >= 0.05 ~ FALSE))
df_gam_feat_aov.z.study %>%
  ggplot() +
  geom_bar(aes(y=sig_fdr))

table(df_gam_feat_aov.z.study$sig_fdr)
```

average w/in subject

```{r}
gam_z_df.mean <- gam_z_df.clean %>%
  pivot_longer(cols=any_of(pheno_list.z), names_to="pheno", values_to = "z_score") %>%
  group_by(participant) %>%
  summarize(study = unique(study),
            site = unique(site),
            age_days=unique(age_days),
            sex=unique(sex),
            mean_z=mean(z_score))
nrow(gam_z_df.mean) == n_unique(gam_z_df.clean$participant)
```

```{r}
gam_z_df.mean %>%
  ggplot() +
  geom_point(aes(x=age_days/365.25, y=mean_z), alpha=0.5) +
  geom_smooth(aes(x=age_days/365.25, y=mean_z), method="lm") 
```

Subject mean ANOVA

```{r}
summary(aov(mean_z ~ study, data=gam_z_df.mean))
```

##ComBatLS

```{r}
#clean predictions
gamlss_centile_df.clean <- full_join(gamlss_centile_df, study_site_key) %>% 
  mutate(age_years=age_days/365.25) %>%
  dplyr::select(!contains(".z")) #drop z-scores and non-centiles & !ends_with("_total") & !contains("TBV")
```

w/in feature ANOVAs

```{r}
gamlss.feat_to_test <- intersect(pheno_list, names(gamlss_centile_df.clean))

df_gamlss_feat_aov <- data.frame()
for(pheno in gamlss.feat_to_test){
  # print(pheno)
  df <- tidy(aov(gamlss_centile_df.clean[[pheno]] ~ gamlss_centile_df.clean[["study"]]))
  df$pheno <- pheno
  df_gamlss_feat_aov <- rbind(df_gamlss_feat_aov, df)
}

df_gamlss_feat_aov.study <- df_gamlss_feat_aov %>%
  group_by(pheno) %>%
  mutate(etasq = sumsq / sum(sumsq)) %>%
  ungroup() %>%
  dplyr::filter(term != "Residuals") %>%
  dplyr::mutate(p.val_fdr = p.adjust(p.value, method="fdr", n = length(gamlss.feat_to_test)), #apply fdr correction across all phenos tested
                  sig_fdr = case_when(p.val_fdr < 0.05 ~ TRUE,
                                      p.val_fdr >= 0.05 ~ FALSE))
df_gamlss_feat_aov.study %>%
  ggplot() +
  geom_bar(aes(y=sig_fdr))

table(df_gamlss_feat_aov.study$sig_fdr)
```

average w/in subject

```{r}
gamlss_centile_df.mean <- gamlss_centile_df.clean %>%
  pivot_longer(cols=any_of(pheno_list), names_to="pheno", values_to = "centile") %>%
  group_by(participant) %>%
  summarize(study = unique(study),
            site = unique(site),
            age_days=unique(age_days),
            sex=unique(sex),
            log_age=unique(log_age),
            sexMale=unique(sexMale),
            fs_version=unique(fs_version),
            mean_cent=mean(centile))
nrow(gamlss_centile_df.mean) == n_unique(gamlss_centile_df.mean$participant)
```

```{r}
gamlss_centile_df.mean %>%
  ggplot() +
  geom_boxplot(aes(x=reorder(study, age_days), y=mean_cent)) + 
  #geom_boxplot(aes(x=study, y=mean_cent)) +
  theme(axis.text.x = element_text(angle=90, hjust=0.5))
```

subject mean ANOVA

```{r}
summary(aov(mean_cent ~ study, data=gamlss_centile_df.mean))
```

###Z-Scores

```{r}
#clean predictions
gamlss_z_df.clean <- full_join(gamlss_centile_df, study_site_key) %>% 
  mutate(age_years=age_days/365.25) %>%
  dplyr::select(!any_of(pheno_list) & !ends_with("_total") & !contains("TBV")) #drop z-scores and non-centiles
```

w/in feature ANOVAs

```{r}
gamlss.feat_to_test.z <- paste0(gamlss.feat_to_test, ".z")
df_gamlss_feat_aov.z <- data.frame()
for(pheno in gamlss.feat_to_test.z){
  # print(pheno)
  df <- tidy(aov(gamlss_z_df.clean[[pheno]] ~ gamlss_z_df.clean[["study"]]))
  df$pheno <- pheno
  df_gamlss_feat_aov.z <- rbind(df_gamlss_feat_aov.z, df)
}

df_gamlss_feat_aov.z.study <- df_gamlss_feat_aov.z %>%
  dplyr::filter(term != "Residuals") %>%
  dplyr::mutate(p.val_fdr = p.adjust(p.value, method="fdr", n = length(gamlss.feat_to_test.z)), #apply fdr correction across all phenos tested
                  sig_fdr = case_when(p.val_fdr < 0.05 ~ TRUE,
                                      p.val_fdr >= 0.05 ~ FALSE))
df_gamlss_feat_aov.z.study %>%
  ggplot() +
  geom_bar(aes(y=sig_fdr))

table(df_gamlss_feat_aov.z.study$sig_fdr)
```

average w/in subject

```{r}
gamlss_z_df.mean <- gamlss_z_df.clean %>%
  pivot_longer(cols=any_of(pheno_list.z), names_to="pheno", values_to = "z_score") %>%
  group_by(participant) %>%
  summarize(study = unique(study),
            site=unique(site),
            age_days=unique(age_days),
            sex=unique(sex),
            mean_z=mean(z_score))
nrow(gamlss_z_df.mean) == n_unique(gamlss_z_df.clean$participant)
```

```{r}
gamlss_z_df.mean %>%
  ggplot() +
  geom_point(aes(x=age_days/365.25, y=mean_z, color=study), alpha=0.5) +
  geom_smooth(aes(x=age_days/365.25, y=mean_z), method="lm") +
  facet_wrap(~study) +
  theme(legend.position="none")
```

subject mean ANOVA

```{r}
summary(aov(mean_z ~ study, data=gamlss_z_df.mean))
```

##Comparison Plots

Number of models converged in each dataset
```{r}
length(raw.feat_to_test)
length(gam.feat_to_test)
length(gamlss.feat_to_test)

#missing features
c(setdiff(pheno_list, gamlss.feat_to_test),
setdiff(pheno_list, raw.feat_to_test),
setdiff(pheno_list, gam.feat_to_test)) %>% unique()
```

```{r}
gamlss_centile_df.mean$combat <- "gamlss"
gam_centile_df.mean$combat <- "gam"
raw_centile_df.mean$combat <- "raw"

#put all data frames into list
df_list.tmp <- list(gamlss_centile_df.mean, gam_centile_df.mean, raw_centile_df.mean)

#merge all data frames in list
mean_cents.df <- df_list.tmp %>% reduce(full_join)

dataset_names <- c(gam = "ComBat-GAM", gamlss = "ComBatLS", raw = "Unharmonized")

mean_cents_bxplt <- mean_cents.df %>%
  ggplot() +
  geom_boxplot(aes(x=reorder(study, age_days), y=mean_cent, fill=combat)) + 
  theme_linedraw() +
  theme(axis.text.x = element_text(angle=90, hjust=0.5), legend.position="none", strip.text.x = element_text(size = 12)) +
  scale_fill_manual(values = c(cf= "#440D54", gam = "#3D75C2", cf.lm = "#75C23D", gamlss ="#CFB023FF", raw="darkgray")) +
  facet_wrap(~combat, labeller = labeller(combat=dataset_names)) +
  labs(y="Mean Centile", x = "Study")

mean_cents_bxplt_nolabel <- mean_cents.df %>%
  ggplot() +
  geom_boxplot(aes(x=reorder(study, age_days), y=mean_cent, fill=combat)) + 
  theme_linedraw() +
  theme(axis.text.x = element_blank(), axis.ticks.x = element_blank(), legend.position="none", strip.text.x = element_text(size = 12)) +
  scale_fill_manual(values = c(cf= "#440D54", gam = "#3D75C2", cf.lm = "#75C23D", gamlss ="#CFB023FF", raw="darkgray")) +
  facet_wrap(~combat, labeller = labeller(combat=dataset_names)) +
  labs(y="Mean Centile", x = "Study")
```

Cohen's F squared for study effects

```{r}
gam_centile_f2 <- fread(file=paste0(der_path, "/lifespan-CN-imp-sites-euler-log-cf-cf.gam-batch.study-data_cohenfsq.csv"))
gam_centile_f2$combat <- "ComBat-GAM"

gamlss_centile_f2 <- fread(file=paste0(der_path, "/lifespan-CN-imp-sites-euler-log-cf-cf.gamlss-batch.study-data_cohenfsq.csv"))
gamlss_centile_f2$combat <- "ComBatLS"

f2.df <- rbind(gam_centile_f2, gamlss_centile_f2)
f2.df <- f2.df %>%
  mutate(pheno_cat = as.factor(case_when(
      pheno %in% vol_list_global ~ "Global Volume",
      pheno %in% vol_list_regions ~ "Regional Volume",
      pheno %in% sa_list ~ "Regional SA",
      pheno %in% ct_list ~ "Regional CT",
      TRUE ~ NA_character_)))

fsq.df <- ggplot(f2.df) +
  geom_col(aes(x=pheno, y=fsq, fill=combat), position="dodge") + 
  facet_wrap(~pheno_cat, scales="free") +
  theme_linedraw() +
  theme(axis.text.x = element_blank(), axis.ticks.x = element_blank()) +
  labs(x = "Feature", y = "Effect Size (Cohen's F-Squared)", fill="ComBat")

ggsave("figs/supp_fsq.jpeg", fsq.df, width=8, height=6, units="in")
```

```{r}
f2.df.sum <- f2.df %>%
  group_by(pheno) %>%
  arrange(combat) %>%
  mutate(F2_diff = fsq - fsq[1]) #gamlss - gam value, + number means site effect is larger in combatLS

ggplot(f2.df.sum) +
  geom_col(aes(x=pheno, y=F2_diff), position="dodge") + 
  facet_wrap(~pheno_cat, scales="free") +
  theme(axis.text.x = element_blank())

f2.df.sum %>%
  group_by(pheno_cat) %>%
  summarise(mean_fsq_diff = mean(F2_diff)) %>%
  ggplot() + 
  geom_col(aes(x=pheno_cat, y=mean_fsq_diff), position="dodge")
```

#ComBat-GAM vs ComBatLS: Delta Centile

Get study info
```{r}
nrow(raw_centile_df.clean) == n_unique(raw_centile_df.clean$participant)

study_info <- raw_centile_df.clean %>%
  group_by(study) %>%
  summarize(study_mean_age=mean(age_years),
            study_median_age=median(age_years),
            study_n = n()) %>%
  dplyr::select(study, study_mean_age, study_median_age, study_n)
```

clean delta df

```{r}
cent_deltas_df <- deltas_df %>%
  dplyr::select(!contains(".z") & !ends_with("_total") & !contains("TBV")) #drop z-scores and non-centiles

#get subject-level means
mean_cent_deltas_df <- cent_deltas_df %>%
  pivot_longer(cols=any_of(paste0("delta.", pheno_list)), names_to="pheno", values_to = "centile") %>%
  group_by(participant) %>%
  summarize(study = unique(study),
            site = unique(site),
            age_days=unique(age_days),
            sex=unique(sex),
            mean_delta_cent=(mean(centile))*100, #convert units to centiles space
            abs_mean_delta_cent=(mean(abs(centile)))*100)

mean_cent_deltas_df_study <- full_join(mean_cent_deltas_df, study_info)
```

control for N when looking at deltas b/c UKB will always have low deltas since it's having an outsized effect on setting the distribution that both ComBatGAM and ComBatLS are targeting.

check mean subj delta ~ study mean age

```{r}
delta_lm <- lm(mean_delta_cent ~ study_mean_age + study_n, data=mean_cent_deltas_df_study)
summary(delta_lm)

mean_cent_deltas_df_study %>%
  ggplot(aes(x=study_mean_age, y=mean_delta_cent)) +
  geom_point(aes(color=study), alpha=0.2) +
  geom_abline(slope = coef(delta_lm)[["study_mean_age"]], 
              intercept = coef(delta_lm)[["(Intercept)"]])+
  theme(legend.position="none")
```

mean abs delta ~ study mean age

```{r}
abs_lm <- lm(abs_mean_delta_cent ~ study_mean_age + study_n, data=mean_cent_deltas_df_study)
summary(abs_lm)

mean_cent_deltas_df_study %>%
  ggplot(aes(x=study_mean_age, y=abs_mean_delta_cent)) +
  geom_point(aes(color=study), alpha=0.2) +
  geom_abline(slope = coef(abs_lm)[["study_mean_age"]], 
              intercept = coef(abs_lm)[["(Intercept)"]])+
  theme(legend.position="none")
```

See if a subject's distance from their study's mean age is assoc with how much their centile changes

```{r}
mean_cent_deltas_df_study <- mean_cent_deltas_df_study %>%
  mutate(age_diff = (age_days/365.25) - study_mean_age,
         abs_age_diff = abs(age_diff))

abs_lm2 <- lm(abs_mean_delta_cent ~ study_n + abs_age_diff, data=mean_cent_deltas_df_study)
summary(abs_lm2)

mean_cent_deltas_df_study %>%
  ggplot(aes(x=abs(age_diff), y=abs_mean_delta_cent)) +
  geom_point(alpha=0.05) +
  geom_smooth(method="lm")
```

## Subject-Level Mean Centiles ~ Mean Study Age

Make sure that results of mean centile delta ~ mean study age are related to the difference itself, not because ComBatLS and/or ComBat-GAM centiles are associated with mean study age.

```{r}
raw_centile_df.mean2 <- full_join(raw_centile_df.mean, study_info)
summary(lm(mean_cent ~ study_mean_age, data=raw_centile_df.mean2))

raw_centile_df.mean2 %>%
  ggplot(aes(x=study_mean_age, y=mean_cent)) +
  geom_point(aes(color=study), alpha=0.7) +
  geom_smooth(method="lm") +
  theme(legend.position="none")
```

NS

```{r}
gam_centile_df.mean2 <- full_join(gam_centile_df.mean, study_info)
summary(lm(mean_cent ~ study_mean_age, data=gam_centile_df.mean2))

gam_centile_df.mean2 %>%
  ggplot(aes(x=study_mean_age, y=mean_cent)) +
  geom_point(aes(color=study), alpha=0.7) +
  geom_smooth(method="lm") +
  theme(legend.position="none")

gam_centile_df.mean2 <- gam_centile_df.mean2 %>%
  mutate(age_diff = (age_days/365.25) - study_mean_age)

summary(lm(mean_cent ~ abs(age_diff), data=gam_centile_df.mean2))
gam_centile_df.mean2 %>%
  ggplot(aes(x=abs(age_diff), y=mean_cent)) +
  geom_point(aes(color=study), alpha=0.7) +
  geom_smooth(method="lm") +
  theme(legend.position="none")
```

NS

```{r}
gamlss_centile_df.mean2 <- full_join(gamlss_centile_df.mean, study_info)
summary(lm(mean_cent ~ study_mean_age, data=gamlss_centile_df.mean2))

gamlss_centile_df.mean2 %>%
  ggplot(aes(x=study_mean_age, y=mean_cent)) +
  geom_point(aes(color=study), alpha=0.7) +
  geom_smooth(method="lm") +
  theme(legend.position="none")

gamlss_centile_df.mean2 <- gamlss_centile_df.mean2 %>%
  mutate(age_diff = (age_days/365.25) - study_mean_age)

summary(lm(mean_cent ~ abs(age_diff), data=gamlss_centile_df.mean2))
gamlss_centile_df.mean2 %>%
  ggplot(aes(x=abs(age_diff), y=mean_cent)) +
  geom_point(aes(color=study), alpha=0.7) +
  geom_smooth(method="lm") +
  theme(legend.position="none")
```

NS

#FIGURE 5

Reformatting plots for prettier composite fig

```{r}
blank <- ggplot() + theme_void()

sample_boxplot2 <- gamlss_centile_df %>%
  mutate(study = factor(study, levels=c(age.df$study), ordered=TRUE)) %>%
ggplot(aes(y=age_years, x=study, color=sex)) +
  geom_boxplot(outlier.size = 1, outlier.colour=NULL) +
  theme_linedraw(base_size = 11) +
  #guides(y = ggh4x::guide_axis_nested(delim = "&")) +
  theme(axis.title.x=element_blank(), axis.text.x = element_blank()) +
  labs(color="Sex", y = "Age (Years)")

#overall sample
full_boxplot2 <- 
  gamlss_centile_df %>%
  mutate(xlab = "Full Sample ") %>%
  ggplot(aes(x=xlab, y=age_years, color=sex)) +
  geom_boxplot(outlier.size = 1, outlier.colour=NULL) +
  theme_linedraw(base_size = 11) +
  theme(axis.title=element_blank(), axis.text=element_blank(), axis.ticks.y=element_blank()) +
  # scale_y_discrete(labels=c("1" = "Full Sample")) +
  labs(color="Sex", x=" ")

sample_barplot2 <- gamlss_centile_df %>%
  mutate(study = factor(study, levels=c(age.df$study), ordered=TRUE)) %>%
  ggplot() +
  geom_bar(aes( y=..count../tapply(..count.., ..x.. ,sum)[..x..], x=study, fill=sex)) +
  #scale_y_continuous(labels = scales::percent) +
  geom_hline(yintercept = .5, color="lightgray", linetype="longdash") +
  theme_linedraw(base_size = 11) +
  scale_y_continuous(labels = scales::percent) +
  #theme(axis.ticks.x = element_blank(), axis.text=element_blank(), axis.title=element_blank()) +
  theme(axis.ticks.x = element_blank(), axis.text.x = element_text(angle=90, hjust=0.5), axis.text.y = element_text(size=6)) +
  labs(fill="Sex", x = "Study", y="Percent Sex")

full_barplot2 <- 
  gamlss_centile_df %>%
  mutate(xlab = "  Full Sample  ") %>%
  ggplot() +
  geom_bar(aes(y=..count../tapply(..count.., ..x.. ,sum)[..x..], x=xlab, fill=sex)) + 
  geom_hline(yintercept = .5, color="lightgray", linetype="longdash") +
  theme_linedraw(base_size = 11) +
  theme(axis.ticks.y = element_blank(), axis.title=element_blank(), axis.text.y=element_blank(), axis.text.x = element_text(angle=90, hjust=0.5, size=12)) +
  labs(color="Sex", y = "Age (Years)", y="")
```

```{r}
# study demographics
fig5_vert <- ggarrange(
  #ggarrange(blank,
ggarrange(sample_boxplot2, full_boxplot2,
          sample_barplot2, full_barplot2,
          nrow=2, ncol=2, 
          #align = "v", 
          common.legend = TRUE,
          legend = "right",
          widths=c(18, 1),
          heights=c(1, 1)
          ),
# blank,
# nrow=1, ncol=3,
# common.legend = FALSE,
# widths= c(1, 20, 1)),
# boxplots
mean_cents_bxplt_nolabel,
common.legend=FALSE,
nrow=2, ncol=1,
heights=c(1, 1.5),
labels="AUTO"
)

ggsave("figs/figure_5_vertical.jpeg", fig5_vert, width=15, height=9, units="in")
```

##Supplemental Figs

mean abs delta ~ study mean age

```{r}
summary(abs_lm)

#predict w CIs
pdat <- with(mean_cent_deltas_df_study,
             tibble(study_mean_age = seq(min(study_mean_age),
                            max(study_mean_age),
                            length = 200),
             study_n = rep(median(study_n), length=200)))

pred_abs_lm <- predict(abs_lm, pdat, se.fit=TRUE)

pred_df <- pdat %>% 
  bind_cols(data.frame(pred_abs_lm))

pred_df <- pred_df %>%
  mutate(upper = fit + (1.96 * se.fit), #95% CIs
         lower = fit - (1.96 * se.fit))

mean_delta_fig <- mean_cent_deltas_df_study %>%
  ggplot(aes(x=study_mean_age, y=abs_mean_delta_cent)) +
  geom_point(aes(color=study), alpha=0.2) +
  geom_ribbon(aes(ymin = lower, ymax = upper,
                    x = study_mean_age), data = pred_df,
                inherit.aes = FALSE, alpha = 0.2) +
    geom_line(data = pred_df, aes(y = fit, x=study_mean_age))+
  theme_linedraw() +
  theme(legend.position="none") +
  labs(x="Mean Age of Primary Study Sample (Years)", y="Mean Absolute Difference in Centile Scores (Centiles)", title="Subject-wise Mean Distance Between Centiles Calculated From ComBatLS and ComBat-GAM \n-harmonized Data")
ggsave("figs/supp_mean_delta_fig.jpeg", mean_delta_fig, width=9, height=6, units="in")
```

lines show marginal effects of mean study age

```{r}
summary(abs_lm2)

#predict w CIs
pdat2 <- with(mean_cent_deltas_df_study,
             tibble(abs_age_diff = seq(min(abs_age_diff),
                            max(abs_age_diff),
                            length = 200),
             study_n = rep(median(study_n), length=200),
             study_mean_age = rep(median(study_mean_age), length=200)))

pred2 <- predict(abs_lm2, pdat2, se.fit = TRUE)

pdat2 <- pdat2 %>%
  bind_cols(data.frame(pred2)) %>%
  mutate(upper = fit + (1.96 * se.fit),
         lower = fit - (1.96 * se.fit))

mean_delta_fig2 <- mean_cent_deltas_df_study %>%
  ggplot(aes(x=abs_age_diff, y=abs_mean_delta_cent)) +
  geom_point(alpha=0.05, aes(color=study)) +
  geom_ribbon(aes(ymin = lower, ymax = upper,
                    x = abs_age_diff), data = pdat2,
                inherit.aes = FALSE, alpha = 0.2) +
    geom_line(data = pdat2, aes(y = fit, x=abs_age_diff)) +
  theme_linedraw() +
  theme(legend.position="none") +
  labs(x="Subject's Offset from Mean Age of Primary Study (Years)", y="Mean Absolute Difference in Centile Scores (Centiles)", title="Subject-wise Mean Distance Between Centiles Calculated From ComBatLS and ComBat-GAM \n-harmonized Data")
ggsave("figs/supp_mean_delta_fig2.jpeg", mean_delta_fig2, width=9, height=6, units="in")
```