### 2025-02-02 paper_centilebrain_supp_results.Rmd
look at centilebrain models and plot

---
title: "ComBatLS LBCC Results"
output: html_document
date: '2023-11-12'
---

Rmd to compile, test, and visualize ComBatLS vs ComBat-GAM harmonization in the LBCC with centilebrain models for reviewers

# Setup

```{r setup, include=FALSE}
knitr::opts_chunk$set(echo = FALSE, warning = FALSE)
if(!require('pacman')) {
  install.packages('pacman')
}
pacman::p_load(data.table, ggplot2, tidyverse, ggseg, ggsegExtra, paletteer, ggpubr, gridExtra, lemon, parallel, rstatix, lme4, interactions, ggrepel, ggrain, skimr, gam, ggeffects, effectsize, gratia)

#PHENO LISTS
pheno_list <- readRDS(file="cubic_scripts/R_scripts/pheno_list.rds")
pheno_list.z <- paste0(pheno_list, ".z")
vol_list_global <- readRDS(file="cubic_scripts/R_scripts/vol_list_global.rds")
vol_list_regions <- readRDS(file="cubic_scripts/R_scripts/Vol_list_regions.rds")
sa_list <- readRDS(file="cubic_scripts/R_scripts/SA_list.rds")
ct_list <- readRDS(file="cubic_scripts/R_scripts/CT_list.rds")

#PATH
der_path <- "/Users/megardn/Desktop/BGD_Repos/combat_biovar/derivatives/cb"
```

#Load

Combat-GAM Centiles - study-level batch correction

```{r}
gam_centile_female <- fread(file=paste0(der_path, "/lifespan-CN-imp-sites-euler-log-cf-cf.gam-batch.study-data-cb-female-data_cb_mod_predictions.csv"))

gam_centile_male <- fread(file=paste0(der_path, "/lifespan-CN-imp-sites-euler-log-cf-cf.gam-batch.study-data-cb-male-data_cb_mod_predictions.csv"))

gam_centile_df <- rbind(gam_centile_female, gam_centile_male)
stopifnot(nrow(gam_centile_df) == (nrow(gam_centile_male) + nrow(gam_centile_female)))

gam_centile_df <- gam_centile_df %>% 
  mutate(age_years=(age_days-280)/365.25) %>%
  unite(INDEX.ID, c(study, participant, sex), sep = "|", remove=FALSE)

n_unique(gam_centile_df$study)

dim(gam_centile_df)
```

Combat-LS Centiles
```{r}
gamlss_centile_female <- fread(file=paste0(der_path, "/lifespan-CN-imp-sites-euler-log-cf-cf.gamlss-batch.study-data-cb-female-data_cb_mod_predictions.csv"))

gamlss_centile_male <- fread(file=paste0(der_path, "/lifespan-CN-imp-sites-euler-log-cf-cf.gamlss-batch.study-data-cb-male-data_cb_mod_predictions.csv"))

gamlss_centile_df <- rbind(gamlss_centile_female, gamlss_centile_male)
stopifnot(nrow(gamlss_centile_df) == (nrow(gamlss_centile_male) + nrow(gamlss_centile_female)))

gamlss_centile_df <- gamlss_centile_df %>% 
  mutate(age_years=(age_days-280)/365.25) %>%
  unite(INDEX.ID, c(study, participant, sex), sep = "|", remove=FALSE)

dim(gamlss_centile_df)
```

```{r}
stopifnot(min(gam_centile_df$age_years) == min(gamlss_centile_df$age_years))
stopifnot(n_unique(gam_centile_df$INDEX.ID) == n_unique(gamlss_centile_df$INDEX.ID))
```

Deltas

```{r}
deltas_df_m <- fread(file=paste0(der_path, "/batch.study-data-cb-male-data_cb_mod_predictions_pred_deltas.csv"))
deltas_df_f <- fread(file=paste0(der_path, "/batch.study-data-cb-female-data_cb_mod_predictions_pred_deltas.csv"))
deltas_df <- rbind(deltas_df_m, deltas_df_f) %>% 
  unite(INDEX.ID, c(study, participant, sex), sep = "|", remove=FALSE) %>%
  mutate(age_years=(age_days-280)/365.25)

dim(deltas_df)
stopifnot(min(deltas_df$age_years) == min(gamlss_centile_df$age_years))
```

#Clean & EDA

Check numbers
```{r}
nrow(gamlss_centile_df) #n
n_unique(gamlss_centile_df$study) 
n_unique(gamlss_centile_df$site) #correct
```

plot mean subject centiles

```{r}
gam_centile_df.mean <- gam_centile_df %>%
  dplyr::select(!contains(".z") & !ends_with("_total") & !contains("TBV")) %>%
  pivot_longer(cols=any_of(pheno_list), names_to="pheno", values_to = "centile") %>%
  group_by(INDEX.ID) %>%
  summarize(study = unique(study),
            age_days=unique(age_days),
            sex=unique(sex),
            mean_cent=mean(centile))
stopifnot(nrow(gam_centile_df.mean) == n_unique(gam_centile_df.mean$INDEX.ID))

gamlss_centile_df.mean <- gamlss_centile_df %>%
  dplyr::select(!contains(".z") & !ends_with("_total") & !contains("TBV")) %>%
  pivot_longer(cols=any_of(pheno_list), names_to="pheno", values_to = "centile") %>%
  group_by(INDEX.ID) %>%
  summarize(study = unique(study),
            age_days=unique(age_days),
            sex=unique(sex),
            mean_cent=mean(centile))
stopifnot(nrow(gamlss_centile_df.mean) == n_unique(gamlss_centile_df.mean$INDEX.ID))
```

```{r}
gam_centile_df.mean$combat <- "ComBat-GAM"
gamlss_centile_df.mean$combat <- "ComBatLS"

mean_cents.df <- full_join(gamlss_centile_df.mean, gam_centile_df.mean)

mean_cents.df %>%
  ggplot() +
  geom_boxplot(aes(x=reorder(study, age_days), y=mean_cent, fill=combat)) + 
  theme_linedraw() +
  theme(axis.text.x = , legend.position="bottom") +
  facet_wrap(~sex) +
  scale_fill_manual(values = c('ComBat-GAM' = "#3D75C2", ComBatLS ="#CFB023FF", Unharmonized = "darkgray"), name = "Harmonization") +
  labs(y="Mean Centile", x = "Study")

mean_cents_bxplt_nolabel <- mean_cents.df %>%
  ggplot() +
  geom_boxplot(aes(x=reorder(study, age_days), y=mean_cent, fill=combat)) + 
  theme_linedraw() +
  theme(axis.text.x = element_blank(), axis.ticks.x = element_blank(), legend.position="bottom", strip.text.x = element_text(size = 12)) +
  facet_wrap(~sex) +
  scale_fill_manual(values = c('ComBat-GAM' = "#3D75C2", ComBatLS ="#CFB023FF", Unharmonized = "darkgray"), name = "Harmonization") +
  labs(y="Mean Centile", x = "Study")
ggsave("figs/cb_mean_centiles.jpeg", mean_cents_bxplt_nolabel, width=12, height=7, units="in")
```

#Cohen's F-squared for study effects

```{r}
raw_centile_f2_m <- fread(file=paste0(der_path, "/lifespan-CN-imp-sites-euler-raw-cb-male-data_cb_cohenfsq.csv"))
raw_centile_f2_m$combat <- "Unharmonized"
raw_centile_f2_m$sex <- "Male"
raw_centile_f2_f <- fread(file=paste0(der_path, "/lifespan-CN-imp-sites-euler-raw-cb-female-data_cb_cohenfsq.csv"))
raw_centile_f2_f$combat <- "Unharmonized"
raw_centile_f2_f$sex <- "Female"

gam_centile_f2_f <- fread(file=paste0(der_path, "/lifespan-CN-imp-sites-euler-log-cf-cf.gam-batch.study-data-cb-female-data_cb_cohenfsq.csv"))
gam_centile_f2_f$combat <- "ComBat-GAM"
gam_centile_f2_f$sex <- "Female"
gam_centile_f2_m <- fread(file=paste0(der_path, "/lifespan-CN-imp-sites-euler-log-cf-cf.gam-batch.study-data-cb-male-data_cb_cohenfsq.csv"))
gam_centile_f2_m$combat <- "ComBat-GAM"
gam_centile_f2_m$sex <- "Male"

gamlss_centile_f2_f <- fread(file=paste0(der_path, "/lifespan-CN-imp-sites-euler-log-cf-cf.gamlss-batch.study-data-cb-female-data_cb_cohenfsq.csv"))
gamlss_centile_f2_f$combat <- "ComBatLS"
gamlss_centile_f2_f$sex <- "Female"
gamlss_centile_f2_m <- fread(file=paste0(der_path, "/lifespan-CN-imp-sites-euler-log-cf-cf.gamlss-batch.study-data-cb-male-data_cb_cohenfsq.csv"))
gamlss_centile_f2_m$combat <- "ComBatLS"
gamlss_centile_f2_m$sex <- "Male"

f2.df <- rbind(raw_centile_f2_m, raw_centile_f2_f, 
               gam_centile_f2_m, gam_centile_f2_f,
               gamlss_centile_f2_m, gamlss_centile_f2_f)
f2.df <- f2.df %>%
  mutate(pheno_cat = as.factor(case_when(
      pheno %in% vol_list_global ~ "Global Volume",
      pheno %in% vol_list_regions ~ "Regional Volume",
      pheno %in% sa_list ~ "Regional SA",
      pheno %in% ct_list ~ "Regional CT",
      TRUE ~ NA_character_)))

unique(f2.df$combat)

f2.df %>%
  group_by(combat) %>%
  summarize(median = median(fsq),
            IQR = IQR(fsq))

length(unique(f2.df$pheno))
setdiff(pheno_list, unique(f2.df$pheno)) #pheno not loaded
```

pretty plot

```{r}
fsq_plt <- ggplot(f2.df) +
  geom_col(aes(x=pheno, y=fsq, fill=combat), position="dodge") + 
  facet_wrap(~pheno_cat, scales="free") +
  theme_linedraw() +
  theme(axis.text.x = element_blank(), legend.position="bottom", axis.ticks.x = element_blank()) + 
  scale_fill_manual(values = c('ComBat-GAM' = "#3D75C2", ComBatLS ="#CFB023FF", Unharmonized = "darkgray"), name = "Harmonization") +
  labs(y="Effect Size (Cohen's F-Squared)", x = "Brain Feature")
ggsave("figs/cohens_fsq_centilebrain.jpeg", fsq_plt)
```

#ComBat-GAM vs ComBatLS: Delta Centile

Get study info
```{r}
study_info <- gam_centile_df %>%
  group_by(study) %>%
  summarize(study_mean_age=mean(age_years),
            study_median_age=median(age_years),
            study_n= n(),
            age_range = max(age_years) - min(age_years)) %>%
  dplyr::select(study, study_mean_age, study_median_age, study_n, age_range)
```

clean delta df

```{r}
cent_deltas_df <- deltas_df %>%
  dplyr::select(!contains(".z") & !ends_with("_total") & !contains("TBV")) #drop z-scores and non-centiles

#get subject-level means
mean_cent_deltas_df <- cent_deltas_df %>%
  pivot_longer(cols=any_of(paste0("delta.", pheno_list)), names_to="pheno", values_to = "centile") %>%
  group_by(INDEX.ID) %>%
  summarize(study = unique(study),
            age_days=unique(age_days),
            sex=unique(sex),
            mean_delta_cent=(mean(centile))*100, #convert units to centiles space
            abs_mean_delta_cent=(mean(abs(centile)))*100)

mean_cent_deltas_df_study <- full_join(mean_cent_deltas_df, study_info)
mean_cent_deltas_df_study <- mean_cent_deltas_df_study %>%
  mutate(age_diff = (age_days/365.25) - study_mean_age,
         abs_age_diff = abs(age_diff))

summary(mean_cent_deltas_df_study$abs_mean_delta_cent)
```

see how often features move between extreme (>95% or <5%) to not-extreme
```{r}
gam_centile_df$combat <- "cf.gam"
gamlss_centile_df$combat <- "cf.gamlss"

centiles_df <- full_join(gam_centile_df, gamlss_centile_df)

centiles_cat <- centiles_df %>%
  dplyr::select(!contains(".z") & !ends_with("_total") & !contains("TBV")) %>%
  mutate(combat = factor(combat, ordered=TRUE, levels=c("cf.gam", "cf.gamlss"))) %>%
  mutate(across(pheno_list, ~ as.factor(case_when(. > 0.95 ~ "high", 
                                               . < 0.05 ~ "low",
                                               TRUE ~ "mid")))) %>%
  arrange(combat) %>%
  group_by(INDEX.ID) %>%
  mutate(across(pheno_list, 
                ~ . == first(.), .names = "is_equal_{col}")) %>%
  ungroup()
saveRDS(centiles_cat, file="centiles_cat.rds")
```

```{r}
centiles_cat_filt <- centiles_cat %>%
  dplyr::select(INDEX.ID, combat, all_of(pheno_list))

ext_count_df <- centiles_cat_filt %>%
  rowwise() %>%
  mutate(count_high = sum(c_across(pheno_list) == "high"),
         count_mid = sum(c_across(pheno_list) == "mid"),
         count_low = sum(c_across(pheno_list) == "low")) %>%
  dplyr::select(INDEX.ID, combat, starts_with("count"))

ext_count_df %>%
  ggplot() +
  geom_bar(aes(x=count_high, fill=combat, color=combat), position="identity", alpha=.4)
ext_count_df %>%
  ggplot() +
  geom_bar(aes(x=count_low, fill=combat, color=combat), position="identity", alpha=.4)
```

```{r}
centiles_cat_filt2 <- centiles_cat %>%
  dplyr::filter(combat =="cf.gamlss") %>%
  dplyr::select(INDEX.ID, starts_with("is_equal_")) %>%
  mutate(sum = rowSums(across(starts_with("is_equal_")))) %>%
  mutate(changed = ifelse(sum == length(pheno_list), "no_change", "change")) #if not all features are true -> changed

stopifnot(max(centiles_cat_filt2$sum) == length(pheno_list)) #correct

length(pheno_list) - mean(centiles_cat_filt2$sum) #mean per subject
length(pheno_list) - min(centiles_cat_filt2$sum) #range
length(pheno_list) - max(centiles_cat_filt2$sum)

#percentage subjects
table(centiles_cat_filt2$changed)["change"]/nrow(centiles_cat_filt2)

#percentage features
#n changed features = # total features - sum
n_feat <- n_unique(centiles_cat_filt2$INDEX.ID)*length(pheno_list)
n_unchanged <- sum(centiles_cat_filt2$sum) #true = 1 = no change

n_feat-n_unchanged
((n_feat-n_unchanged)/n_feat)*100 # % changed
```

plot
```{r}
centiles_cat_filt2_long <- centiles_cat_filt2 %>%
  pivot_longer(cols=starts_with("is_equal_"), names_to="pheno", names_prefix="is_equal_", values_to="is_equal")

n_unique(centiles_cat_filt2_long$INDEX.ID)*208 == nrow(centiles_cat_filt2_long)

cent_cat_plt <- centiles_cat_filt2_long %>%
  dplyr::filter(is_equal==FALSE) %>%
ggplot() +
  geom_bar(aes(x=pheno), fill="#1D213B") +
  theme_linedraw()+
  theme(axis.text.x=element_blank(), axis.ticks.x = element_blank()) +
  labs(x="Brain Feature", y= "Instances of Centile Re-categorization")
ggsave("figs/cent_cat_plt.jpeg", cent_cat_plt, width=10, height=6, units="in")
```

check mean subj delta ~ study mean age

```{r}
delta_lm <- lm(mean_delta_cent ~ study_mean_age + study_n, data=mean_cent_deltas_df_study)
summary(delta_lm)

mean_cent_deltas_df_study %>%
  ggplot(aes(x=study_mean_age, y=mean_delta_cent)) +
  geom_point(aes(color=study), alpha=0.2) +
  geom_abline(slope = coef(delta_lm)[["study_mean_age"]], 
              intercept = coef(delta_lm)[["(Intercept)"]])+
  theme(legend.position="none")
```

mean abs delta ~ study mean age

```{r}
abs_lm <- lm(abs_mean_delta_cent ~ study_mean_age + study_n, data=mean_cent_deltas_df_study)
summary(abs_lm)

mean_cent_deltas_df_study %>%
  ggplot(aes(x=study_mean_age, y=abs_mean_delta_cent)) +
  geom_point(aes(color=study), alpha=0.2) +
  geom_abline(slope = coef(abs_lm)[["study_mean_age"]], 
              intercept = coef(abs_lm)[["(Intercept)"]])+
  theme(legend.position="none")

mean_cent_deltas_df_study <- mean_cent_deltas_df_study %>%
  mutate(age_diff = (age_days/365.25) - study_mean_age,
         abs_age_diff = abs(age_diff))
```

mean abs delta ~ age_diff

```{r}
abs_lm2 <- lm(abs_mean_delta_cent ~ study_n + abs_age_diff, data=mean_cent_deltas_df_study)
summary(abs_lm2)

mean_cent_deltas_df_study %>%
  ggplot(aes(x=abs(age_diff), y=abs_mean_delta_cent)) +
  geom_point(alpha=0.05) +
  geom_smooth(method="lm")
```

similarly, studies with larger age ranges change more
```{r}
summary(lm(abs_mean_delta_cent ~ age_range + study_n, data=mean_cent_deltas_df_study))

mean_cent_deltas_df_study %>%
  ggplot(aes(x=age_range, y=abs_mean_delta_cent, color=study)) +
  geom_point(alpha=0.05) +
  geom_smooth(method="lm") +
  theme(legend.position = "none")
```

control for N when looking at deltas b/c UKB will always have low deltas since it's having an outsized effect on setting the distribution that both ComBatGAM and ComBatLS are targeting.

## Subject-Level Mean Centiles ~ Mean Study Age

Make sure that delta centile results are not driven by either ComBatLS or ComBat-GAM centiles being assoc. with study age

```{r}
gam_centile_df.mean2 <- full_join(gam_centile_df.mean, study_info)
summary(lm(mean_cent ~ study_mean_age, data=gam_centile_df.mean2))

gam_centile_df.mean2 %>%
  ggplot(aes(x=study_mean_age, y=mean_cent)) +
  geom_point(aes(color=study), alpha=0.7) +
  geom_smooth(method="lm") +
  theme(legend.position="none")

gam_centile_df.mean2 <- gam_centile_df.mean2 %>%
  mutate(age_diff = (age_days/365.25) - study_mean_age)

summary(lm(mean_cent ~ abs(age_diff), data=gam_centile_df.mean2))
gam_centile_df.mean2 %>%
  ggplot(aes(x=abs(age_diff), y=mean_cent)) +
  geom_point(aes(color=study), alpha=0.7) +
  geom_smooth(method="lm") +
  theme(legend.position="none")
```

```{r}
gamlss_centile_df.mean2 <- full_join(gamlss_centile_df.mean, study_info)
summary(lm(mean_cent ~ study_mean_age, data=gamlss_centile_df.mean2))

gamlss_centile_df.mean2 %>%
  ggplot(aes(x=study_mean_age, y=mean_cent)) +
  geom_point(aes(color=study), alpha=0.7) +
  geom_smooth(method="lm") +
  theme(legend.position="none")

gamlss_centile_df.mean2 <- gamlss_centile_df.mean2 %>%
  mutate(age_diff = (age_days/365.25) - study_mean_age)

summary(lm(mean_cent ~ abs(age_diff), data=gamlss_centile_df.mean2))
gamlss_centile_df.mean2 %>%
  ggplot(aes(x=abs(age_diff), y=mean_cent)) +
  geom_point(aes(color=study), alpha=0.7) +
  geom_smooth(method="lm") +
  theme(legend.position="none")
```

#Supplemental Figs

mean abs delta ~ study mean age

```{r}
summary(abs_lm)

#predict w CIs
pdat1 <- with(mean_cent_deltas_df_study,
             tibble(study_mean_age = seq(min(study_mean_age),
                            max(study_mean_age),
                            length = 200),
             study_n = rep(median(study_n), length=200)))

pred_abs_lm1 <- predict(abs_lm, pdat1, se.fit=TRUE)

pred_df1 <- pdat1 %>% 
  bind_cols(data.frame(pred_abs_lm1))

pred_df1 <- pred_df1 %>%
  mutate(upper = fit + (1.96 * se.fit), #95% CIs
         lower = fit - (1.96 * se.fit))

mean_delta_fig <- mean_cent_deltas_df_study %>%
  mutate(study = gsub("_", " ", study)) %>%
  ggplot(aes(x=study_mean_age, y=abs_mean_delta_cent)) +
  geom_point(aes(color=study), alpha=0.2) +
  geom_ribbon(aes(ymin = lower, ymax = upper,
                    x = study_mean_age), data = pred_df1,
                inherit.aes = FALSE, alpha = 0.2) +
    geom_line(data = pred_df1, aes(y = fit, x=study_mean_age))+
  theme_linedraw()+
  theme(legend.position="bottom") +
  labs(x="Mean Age of Primary Study Sample (Years)", y="Mean Absolute\nDifference in Centiles",
       color="Study")
#ggsave("figs/supp_mean_delta_fig.jpeg", mean_delta_fig, width=10, height=7, units="in")
```

mean abs delta ~ study range

```{r}
range_mod <- lm(abs_mean_delta_cent ~ age_range + study_n, data=mean_cent_deltas_df_study)
summary(range_mod)

#predict w CIs
pdat2 <- with(mean_cent_deltas_df_study,
             tibble(age_range = seq(min(age_range),
                            max(age_range),
                            length = 200),
             study_n = rep(median(study_n), length=200)))

pred_abs_lm2 <- predict(range_mod, pdat2, se.fit=TRUE)

pred_df2 <- pdat2 %>% 
  bind_cols(data.frame(pred_abs_lm2))

pred_df2 <- pred_df2 %>%
  mutate(upper = fit + (1.96 * se.fit), #95% CIs
         lower = fit - (1.96 * se.fit))

range_delta_fig <- mean_cent_deltas_df_study %>%
  mutate(study = gsub("_", " ", study)) %>%
  ggplot(aes(x=age_range, y=abs_mean_delta_cent)) +
  geom_point(aes(color=study), alpha=0.2) +
  geom_ribbon(aes(ymin = lower, ymax = upper,
                    x = age_range), data = pred_df2,
                inherit.aes = FALSE, alpha = 0.2) +
    geom_line(data = pred_df2, aes(y = fit, x=age_range))+
  theme_linedraw()+
  theme(legend.position="bottom") +
  labs(x="Age Range of Primary Study Sample (Years)", y="Mean Absolute\nDifference in Centiles")
```

```{r}
summary(abs_lm2)

#predict w CIs
pdat3 <- with(mean_cent_deltas_df_study,
             tibble(abs_age_diff = seq(min(abs_age_diff),
                            max(abs_age_diff),
                            length = 200),
             study_n = rep(median(study_n), length=200),
             study_mean_age = rep(median(study_mean_age), length=200)))

pred3 <- predict(abs_lm2, pdat3, se.fit = TRUE)

pdat3 <- pdat3 %>%
  bind_cols(data.frame(pred3)) %>%
  mutate(upper = fit + (1.96 * se.fit),
         lower = fit - (1.96 * se.fit))

mean_delta_fig2 <- mean_cent_deltas_df_study %>%
  mutate(study = gsub("_", " ", study)) %>%
  ggplot(aes(x=abs_age_diff, y=abs_mean_delta_cent)) +
  geom_point(alpha=0.05, aes(color=study)) +
  geom_ribbon(aes(ymin = lower, ymax = upper,
                    x = abs_age_diff), data = pdat3,
                inherit.aes = FALSE, alpha = 0.4) +
    geom_line(data = pdat3, aes(y = fit, x=abs_age_diff)) +
  theme_linedraw()+
  theme(legend.position="bottom") +
  labs(x="Subject's Offset from Mean Age of Primary Study (Years)", y="Mean Absolute\nDifference in Centiles",
       color="Study")

```

```{r}
sfig_cb <- ggarrange(
mean_delta_fig,
range_delta_fig,
mean_delta_fig2,
nrow=3, ncol=1,
labels=c("B", "C", "D"),
legend = "right",
common.legend=TRUE
)

sfig_cb2 <- ggarrange(
fsq_plt,
sfig_cb,
nrow=2, ncol=1,
labels=c("A", ""),
heights=c(1.4, 2),
legend = "bottom",
common.legend=FALSE
)

ggsave("figs/supp_cb.jpeg", sfig_cb2, width=10, height=12, units="in")
```